### 2023-03-11 manuscript.tex
Initial Overleaf Import

\documentclass[11pt]{article}
\usepackage[margin=0.6in]{geometry}
\usepackage[utf8]{inputenc}
\usepackage{authblk}
\usepackage{doi}
\usepackage{tcolorbox}
\usepackage{enumitem}
\usepackage{graphicx,pdflscape,multirow}
\usepackage{array}
\usepackage{xcolor}
\usepackage{multicol}
\usepackage{wrapfig,lipsum,booktabs}
\usepackage{fancybox}
\usepackage{amsmath}

\newcommand{\fixme}[1]{{\color{red} (#1)}}
\newcommand{\coloc}{\texttt{spCOCOON}}

% \renewcommand\Authfont{\fontsize{8}{14.4}\selectfont} % change author fontsize
% \renewcommand\Affilfont{\fontsize{6}{10.8}\itshape}   % change auth affil fontsize
\makeatletter % make affiliations on one line
\renewcommand\AB@affilsepx{, \protect\Affilfont}
\makeatother

\usepackage[sort&compress,square,numbers]{natbib}
\bibliographystyle{unsrtnat}

%\setmainfont{Helvetica}
\title{\coloc: Spatial Visualization Using Color Contrast Optical Illusion}
\author[1]{Boyi Guo\thanks{Corresponding author}}
\author[1]{Stephanie C. Hicks}
\affil[1]{Department of Biostatistics, Johns Hopkins Bloomberg School of Public Health, MD, USA}
\date{\today}

\begin{document}
\maketitle

\vspace{-.6in}

%% ABSTRACT ============================================================================================
\section*{Abstract}
\subsection*{Summary}
Colocalization is highly interested research questions in the analysis of spatial omics data. While many sophisticated computational models are proposed to infer the colocalization relationship between genes or proteins in the spatial context, there completely lack of intuitive spatial visualization, making interpretation extremely challenging. The biggest challenge for colocalization visualization is two sources of orthogonal information, spot colocalization and spot membership (such as spatial domain) the same spatial plot. To address this issue, we developed \coloc to display the colocalization while maintaining an intuitive representation of the surrounding spatial domain. \coloc leverages color contrast optical illusion to display spatial domains to create shadows representing different spatial domain while using different spots to denote colocalization status.

\subsection*{Availability and implementation}
\coloc R package is freely available at Github (\url{https://github.com/boyiguo1/VisColoc}).
% and Zenodo (\fixme{add link}).

%% INTRODUCTION ============================================================================================
\section*{Introduction}
Data visualization plays a critical role in bioinformatics, serving many purposes such as verifying hypothesis, scientific communication and many more. Many researchers treat data visualization as imperative part in the cognitive process to understand hypothesis and models. With the increasly complex data generated in the bioinformatics field, tranmendous of the efforts are spend on developing new computational methods to interpolate knowledge from the data, with little efforts on how to visualize these complex data. "extracting insight from such massive, complex datasets requires development of highly tailored, innovative visual analysis methods" (O'Donoghue SI, 2021).

 The advent of spatial omics technologies brought new outlook to study molecular, enabling researchers to investigate omics in the spatial contexts. Nevertheless, visualizing spatial data are growingly challenging due to the necessity to display omics information while maintaining the spatial relationship. Previous data visualization of omics largely relies on projecting the omics space to an embedding space, such as t-SNE and UMAP, which is infeasible for spatial omics technologies. One of the most commonly used visualization strategy is to color code each spatial unit to represent a dimension information of interest, for example the level of expression or spatial domains. A gap exists (both literally and figuratively) to simultaneously visualization two dimension of information, e.g. comparing the level of colocalization between 2 spatial domains, creating interpretation challenges. There completely lack efficient and informative visualization tools to simultaneously visualize co-localization in the spatial context, preventing the interpretation and understanding of the multi-omics profiling. While interactive platforms or 3-D visualization can alleviate this problems, these alternative solutions are not publication firendly that are normally static. It is imperative to develop a static spatial visualization that allows the simultaneous display of more than 1 dimension of information.

\noindent To address this problem, we propose a novel data visualization for spatial omics data that allows simulatneous display of multiple information. The new data visaulization \textbf{leverage color contrast optical illusion}\cite{zhang_2018} to create another layer of information in spatial maps. We apply \coloc to the spatial transcriptomics dataset (Fig. 1G). The results show \fixme{colocalization}.

\section*{Methods and Implementation}
The main thrust of \coloc is to integrate another dimension of information to the spatial map using color constrast optical illusion. The color contrast optical illusion is a phenomenon where the appearance of a color is influenced by the colors surrounding it. This illusion occurs due to the way our brain processes visual information and perceives color. In a spatial map, we can construct each spatial unit as a hollow points using \texttt{ggplot2::geom\_points} where the boundary of each spatial unit are color-coded to represent one dimension of the information, working for both continuous measures such as gene expression or categorical measures such as spatial domains. It is possible to add another layer of information for an additional variable using sympols or contrast colors in side of each spot. Hence, wihtin each spatial map, multiple layers of informaiton can be displayed jointly.

\coloc is  an open source package implemented in the R programming environment. The package \coloc is constructed based on \texttt{ggplot2} and can be easily integrated to any spatial visualization pipeline, for example \texttt{spaitalLIBD} and \texttt{seurat}, that adapts \texttt{ggplot2} principles and would be able to generalized to users in both bioconductor community and seurat community

\section*{Application: Visualizing Colocalization}
To demonstrate the new visualization, we apply \coloc to visaulize the collocalization of two genes, (FYN) and (EFNA5) in the human brain dorsolateral prefrontal cortex. The two genes are previously identified that relates to SCZ expression via cell-cell communication analysis. The purpose of the visualization is to see if the density of colocalization are different across different cortex layers. Using \coloc, we can simultaneously display the colocalization and cortex layer in the same spatial plot, instead of having two saprate plots, one for cortex layer and one for colocalization, making the verification of the visual within the spaital context more straightforward and easier.

To create a colocalization plots, \coloc follows a similar principle as \texttt{ggplot2}'s layering strategy. In the first step, we will use the funciton \texttt{spCOCOON} to create a base plot that converts a spatial transcroptimcs data structure to a plot object. In a second step, we use the function \texttt{add\_fill} to color code each spots with colcalization status, which includes three levels, spots that expressed both genes and spots that expressed either of the genes. Then, we will apply the function \texttt{add\_COCOON} to show the spatial domain (cortex layer) information. Optionally, one can apply the funciton \texttt{add\_symbol} to display another layer of information. Here, we use  the symbol to highlight the colocalization spots, i.e. the spots that expresses both genes. Because we followed the layering strategy that \texttt{ggplot2}. One can seamsely to integrate \coloc to any \texttt{ggplot2}-based spatial visiulization by piping a \texttt{ggplot2} object to \texttt{add\_COCOON} and \texttt{add\_symbol} and customize the plots using the \texttt{ggplot2} syntex and themes, greatly reduce the techinical burden to customize these new spatial plot designs.

\section*{Summary}

While \coloc is to design to display colocalization in their spatial context, the underlying principle can be broadly applied to address visualizaiton challenges in informaiton rich visualizations. The contrast color optical illusion provides allows to add another dimension of information in any spatial visualization, which can be further used to display multi-omics presence, greatly improve the interpretability.

%% EPILOGUE ================================================================
\section*{Acknowledgements}
B.G. would like to acknowledge Dr. Leonardo Collado Torres,  and Louise Huuki-Myers for their helpful feedback.

\vspace{0.2in}
\noindent \textit{Conflict of Interest:} none declared.

\section*{Data availability}
The spatial transcriptomics dataset was obtained from \texttt{spatialLIBD}(\url{http://research.libd.org/spatialLIBD/}). The code that generates these figures are deposited at \fixme{github manuscript repo}.

\clearpage
% \vspace{-0.1in}
\begin{wrapfigure}[13]{r}{0.8\textwidth}
\vspace{-0.3in}
\begin{center}
\includegraphics[width=0.78\textwidth]{figure/new_visiual.png}
\end{center}
\vspace{-0.35in}
\caption{\footnotesize \textbf{Optimized visualization interpreting co-localization in the spatial context}. (\textbf{A}) Current visualization of co-localization requires two separate plots, one to display spatial domain information and one to display co-localization, difficulty to associate the two sources of information for interpretation. (\textbf{B}) The proposed visualization is optimized to display both spatial domain and co-localization information in one plot.}
\label{fig:visual} 
\end{wrapfigure}

%% BIBLIOGRAPHY ==========================================================

\clearpage 
% \printbibliography
\bibliography{refs}

\end{document}

### 2023-03-17 Manuscript/manuscript.tex
Updates from Overleaf

--- 2023-03-11 manuscript.tex
@@
 \newcommand{\fixme}[1]{{\color{red} (#1)}}
-\newcommand{\coloc}{\texttt{spCOCOON}}
+\newcommand{\coloc}{\texttt{escheR}}
 
@@
 %\setmainfont{Helvetica}
-\title{\coloc: Spatial Visualization Using Color Contrast Optical Illusion}
+\title{\coloc: Optimize Spatial Visualization Following Gestalt Principles}
 \author[1]{Boyi Guo\thanks{Corresponding author}}
@@
 \subsection*{Availability and implementation}
-\coloc R package is freely available at Github (\url{https://github.com/boyiguo1/VisColoc}).
+The R package \coloc is freely available at Github (\url{https://github.com/boyiguo1/VisColoc}).
 % and Zenodo (\fixme{add link}).
@@
 \section*{Introduction}
-Data visualization plays a critical role in bioinformatics, serving many purposes such as verifying hypothesis, scientific communication and many more.
-Many researchers treat data visualization as imperative part in the cognitive process to understand hypothesis and models.
-With the increasly complex data generated in the bioinformatics field, tranmendous of the efforts are spend on developing new computational methods to interpolate knowledge from the data, with little efforts on how to visualize these complex data.
-"extracting insight from such massive, complex datasets requires development of highly tailored, innovative visual analysis methods" (O'Donoghue SI, 2021).
+Visualization is an indispensable component of data analysis, providing clarity that connects quantitative evidence to key conclusions.
+\cite{dagostinomcgowan_2022} In bioinformatics research, visualization receives growing recognition as essential: many scientists rely on visualization to complete their cognitive process from analysis to insight, including analytic validation of automated pipelines and scientific communication.
+\cite{odonoghue_2021} Nevertheless, current methods and tools are inadequate to visualize the growingly complex biological data and computational models, limiting the ability to derive new insights and discoveries.
+\cite{odonoghue_2010}
 
- The advent of spatial omics technologies brought new outlook to study molecular, enabling researchers to investigate omics in the spatial contexts.
-Nevertheless, visualizing spatial data are growingly challenging due to the necessity to display omics information while maintaining the spatial relationship.
-Previous data visualization of omics largely relies on projecting the omics space to an embedding space, such as t-SNE and UMAP, which is infeasible for spatial omics technologies.
-One of the most commonly used visualization strategy is to color code each spatial unit to represent a dimension information of interest, for example the level of expression or spatial domains.
-A gap exists (both literally and figuratively) to simultaneously visualization two dimension of information, e.g.
-comparing the level of colocalization between 2 spatial domains, creating interpretation challenges.
-There completely lack efficient and informative visualization tools to simultaneously visualize co-localization in the spatial context, preventing the interpretation and understanding of the multi-omics profiling.
-While interactive platforms or 3-D visualization can alleviate this problems, these alternative solutions are not publication firendly that are normally static.
-It is imperative to develop a static spatial visualization that allows the simultaneous display of more than 1 dimension of information.
+The recent technology breakthrough in spatial profiling enables characterizing the diversity of molecules, e.g.
+DNA, RNA, and proteins, in the spatial context \cite{moffitt_2022}, generating novel spatially-resolved data that requires highly tailored \textit{in situ} visualizations \cite{dries_2021, Lewis_2021, odonoghue_2021}.
+Current \textit{in situ} visualizations, as opposed to embedding visualizations (e.g.
+t-SNE and UMAP) that project information to some mathematical space, visualize molecular information in the original cellular location, creating spatial maps that color-code a single variable (e.g., molecule expression, cell type, or spatial domain).
+This approach lacks the ability to simultaneously exhibit multiple variables, possibly from disparate data domains (such as expression domain and spatial domain) or modalities (such as transcriptomics and proteomics), creating cognitive gaps to interpret findings regarding their (micro-) environment.
+For example, when studying differential expression in different spatial domains (commonly defined by tissue architecture, e.g.
+cancerous regions, cortical layers), two spatial maps are displayed side-by-side, one for gene expression and one for spatial domains, creating significant challenges to associate these two information together.
+While software-based interactive visualizations or 3-D visualizations have the potential to mitigate this challenge, they are infeasible for scientific communications in static media, such as prints.
+Developing a static spatial visualization that enables the simultaneous display of multiple dimensions of information is crucial for spatially-resolved data research.
 
-\noindent To address this problem, we propose a novel data visualization for spatial omics data that allows simulatneous display of multiple information.
-The new data visaulization \textbf{leverage color contrast optical illusion}\cite{zhang_2018} to create another layer of information in spatial maps.
+% Previous data visualization of omics largely relies on projecting the omics space to an embedding space, such as t-SNE and UMAP, which is infeasible for spatial omics technologies.
+
+ To optimize \textit{in situ} visualization, we introduce gestalt principles to the design of spatial maps.
+Gestalt principles \cite{todorovic_2008, palmer_1999} refer to a set of rules describing how humans perceive and interpret visual information and are commonly applied in art and designs.
+Following gestalt principles, we add additional visual dimensions to spatial maps, leveraging different aesthetics, such as color, fill, and symbols, to simultaneously display disparate variables.
+Specifically, we apply figure-ground articulation to display two variables: one variable (e.g.
+expression) can be plotted as color-filled circles, serving as the \textit{figure}; one variable (e.g.
+spatial domains) can be plotted as the backgrounds of the circles, creating a \textit{ground} for the figure.
+For adjacent circles with limited space between them to display the background color, we use an economic implementation, colored outlines for these circles, inspired by watercolor effect \cite{pinna_1987, pinna_2001}.
+Watercolor effect describes the phenomenon in visual perception that surface color arises from thin boundaries and hence is applied here to perceive the background color in tight space.
+Overall, the figure-ground articulation creates two isolated layers in visual perception to display the two variables while maintaining the relative spatial relationship serving as a reference between the two.
+In addition, other fundamental principles \cite{todorovic_2008}, such as proximity, similarity, continuity, and closure, incentivize the brain to group elements and dimensions in the visualization, guaranteeing an integrative perception of the complex multi-dimensional spatial map.
+
+The main thrust of \coloc is to integrate another dimension of information to the spatial map using color contrast optical illusion.
+The color contrast optical illusion is a phenomenon where the appearance of a color is influenced by the colors surrounding it.
+This illusion occurs due to the way our brain processes visual information and perceives color.
+In a spatial map, we can construct each spatial unit as a hollow point using \texttt{ggplot2::geom\_points} where the boundary of each spatial unit are color-coded to represent one dimension of the information, working for both continuous measures such as gene expression or categorical measures such as spatial domains.
+It is possible to add another layer of information for an additional variable using symbols or contrast colors on side of each spot.
+Hence, within each spatial map, multiple layers of information can be displayed jointly.
+
+\coloc is  an open-source package implemented in the R programming environment.
+The package \coloc is constructed based on \texttt{ggplot2} and can be easily integrated into any spatial visualization pipeline, for example, \texttt{spaitalLIBD} and \texttt{seurat}, that adapts \texttt{ggplot2} principles and would be able to generalize to users in both Bioconductor community and seurat community
+
+ in an interactive process
+
+ We provide an R package called \coloc (named after the graphic artist M.C. Escher) to implement this visual innovation.
+
+ , following an iterative process that adds layers to a spatial map.
+Extending \texttt{ggplot2}, \coloc can seamlessly integrate to any established visualizations that is based on the most popular R visualization framework \texttt{ggplot2}, following the same syntax.
 We apply \coloc to the spatial transcriptomics dataset (Fig.
@@
 
-\section*{Methods and Implementation}
-The main thrust of \coloc is to integrate another dimension of information to the spatial map using color constrast optical illusion.
-The color contrast optical illusion is a phenomenon where the appearance of a color is influenced by the colors surrounding it.
-This illusion occurs due to the way our brain processes visual information and perceives color.
-In a spatial map, we can construct each spatial unit as a hollow points using \texttt{ggplot2::geom\_points} where the boundary of each spatial unit are color-coded to represent one dimension of the information, working for both continuous measures such as gene expression or categorical measures such as spatial domains.
-It is possible to add another layer of information for an additional variable using sympols or contrast colors in side of each spot.
-Hence, wihtin each spatial map, multiple layers of informaiton can be displayed jointly.
-
-\coloc is  an open source package implemented in the R programming environment.
-The package \coloc is constructed based on \texttt{ggplot2} and can be easily integrated to any spatial visualization pipeline, for example \texttt{spaitalLIBD} and \texttt{seurat}, that adapts \texttt{ggplot2} principles and would be able to generalized to users in both bioconductor community and seurat community
-
-\section*{Application: Visualizing Colocalization}
-To demonstrate the new visualization, we apply \coloc to visaulize the collocalization of two genes, (FYN) and (EFNA5) in the human brain dorsolateral prefrontal cortex.
+\section*{Use Case 1: Differential Expression in Spatial Domains}
+To demonstrate the new visualization, we apply \coloc to visualize the colocalization of two genes, (FYN) and (EFNA5) in the human dorsolateral prefrontal cortex.
 The two genes are previously identified that relates to SCZ expression via cell-cell communication analysis.
@@
 
+\section*{Use Case 2: Multi-dimensional UMAP Plot}
+
 \section*{Summary}
@@
 The contrast color optical illusion provides allows to add another dimension of information in any spatial visualization, which can be further used to display multi-omics presence, greatly improve the interpretability.
+
+In this manuscript, we demonstrate the utility of the proposed spatial visualization using spot-based spatial resolved data, nevertheless, this method can be applied to image-based spatial resolved data by applying \texttt{schex} to avoid overplotting commonly for image-based spatial resolved data, similar to our second UMAP example where the first step is to reduce done the dimensiona
+
+our programming provides flexibility by following the grammar of graphs, especially ggplot2
+
+translating the findings in vision science to dat science to address many of the current challenges in visualization.
 

### 2023-03-17 Manuscript/manuscript.tex
Finish revise introduction

@@
 
-The main thrust of \coloc is to integrate another dimension of information to the spatial map using color contrast optical illusion.
-The color contrast optical illusion is a phenomenon where the appearance of a color is influenced by the colors surrounding it.
-This illusion occurs due to the way our brain processes visual information and perceives color.
-In a spatial map, we can construct each spatial unit as a hollow point using \texttt{ggplot2::geom\_points} where the boundary of each spatial unit are color-coded to represent one dimension of the information, working for both continuous measures such as gene expression or categorical measures such as spatial domains.
-It is possible to add another layer of information for an additional variable using symbols or contrast colors on side of each spot.
-Hence, within each spatial map, multiple layers of information can be displayed jointly.
+To implement the proposed visualization, we leverage the state-of-art data visualization framework \texttt{ggplot2} \cite{ggplot2} following the grammar of graphics \cite{wilkinson_2012}.
+Specifically, we use an interactive process to layer each variable in spatial maps using different aesthetics.
+For example, we use the combination of \texttt{color} and \texttt{fill="transparent"} to create the background layer and \texttt{fill} to create the figure layer.
+When necessary to display an additional layer for a third variable, \texttt{shape} can be used to add symbols such as cross (+) and asterisk (*) to highlight in the spatial map.
+We provide an open-source package called \coloc (named after the graphic artist M.C.
+Escher) in the R programming environment \cite{R}, providing a simplified interface to navigate the implementation of the multi-dimensional \textit{in situ} visualization.
+By adapting \texttt{ggplot2} standard, \coloc can be seamlessly integrated into many popular spatial resolved toolboxes, such as \texttt{SpatialLIBD}\cite{pardo_2022}, \texttt{Seurat}\cite{hao_2021}, \texttt{Giotto} \cite{dries_2021} to name a few, and allow further theme customization with ease.
+We provide two use cases to examplify some utility of the proposed spatial visualization: 1) the spatially differential gene colocalization in the human dorsolateral prefrontal cortex (DLPFC) using spatial transcriptomics data; 2) multi-dimensional UMAP highlighting differential gene expression in data-driven cell clusters.
 
-\coloc is  an open-source package implemented in the R programming environment.
-The package \coloc is constructed based on \texttt{ggplot2} and can be easily integrated into any spatial visualization pipeline, for example, \texttt{spaitalLIBD} and \texttt{seurat}, that adapts \texttt{ggplot2} principles and would be able to generalize to users in both Bioconductor community and seurat community
-
- in an interactive process
-
- We provide an R package called \coloc (named after the graphic artist M.C. Escher) to implement this visual innovation.
-
- , following an iterative process that adds layers to a spatial map.
-Extending \texttt{ggplot2}, \coloc can seamlessly integrate to any established visualizations that is based on the most popular R visualization framework \texttt{ggplot2}, following the same syntax.
-We apply \coloc to the spatial transcriptomics dataset (Fig.
-1G).
-The results show \fixme{colocalization}.
-
-\section*{Use Case 1: Differential Expression in Spatial Domains}
+\section*{Use Case 1: Spatially Differential Colocalization in DLPFC}
 To demonstrate the new visualization, we apply \coloc to visualize the colocalization of two genes, (FYN) and (EFNA5) in the human dorsolateral prefrontal cortex.
@@
 
-\section*{Use Case 2: Multi-dimensional UMAP Plot}
+\section*{Use Case 2: Multi-dimensional UMAP}
 
@@
 \section*{Acknowledgements}
-B.G. would like to acknowledge Dr. Leonardo Collado Torres,  and Louise Huuki-Myers for their helpful feedback.
+B.G. would like to acknowledge Dr. Leonardo Collado Torres, Louise Huuki-Myers and Leon Di Stefano for their helpful feedback.
 
@@
 \section*{Data availability}
-The spatial transcriptomics dataset was obtained from \texttt{spatialLIBD}(\url{http://research.libd.org/spatialLIBD/}).
+The spatial transcriptomics dataset was obtained from \texttt{spatialLIBD} (\url{http://research.libd.org/spatialLIBD/}).
+The UMAP example follows the tutorial from \texttt{schex} (\url{https://www.bioconductor.org/packages/release/bioc/vignettes/schex/inst/doc/using_schex.html}).
 The code that generates these figures are deposited at \fixme{github manuscript repo}.

### 2023-03-17 Manuscript/manuscript.tex
add references.

@@
 \section*{Introduction}
-Visualization is an indispensable component of data analysis, providing clarity that connects quantitative evidence to key conclusions.
-\cite{dagostinomcgowan_2022} In bioinformatics research, visualization receives growing recognition as essential: many scientists rely on visualization to complete their cognitive process from analysis to insight, including analytic validation of automated pipelines and scientific communication.
-\cite{odonoghue_2021} Nevertheless, current methods and tools are inadequate to visualize the growingly complex biological data and computational models, limiting the ability to derive new insights and discoveries.
-\cite{odonoghue_2010}
+Visualization is an indispensable component of data analysis, providing clarity that connects quantitative evidence to key conclusions.\cite{dagostinomcgowan_2022} In bioinformatics research, visualization receives growing recognition as essential: many scientists rely on visualization to complete their cognitive process from analysis to insight, including analytic validation of automated pipelines and scientific communication.\cite{odonoghue_2021} Nevertheless, current methods and tools are inadequate to visualize the growingly complex biological data and computational models, limiting the ability to derive new insights and discoveries.\cite{odonoghue_2010}
 
 The recent technology breakthrough in spatial profiling enables characterizing the diversity of molecules, e.g.
-DNA, RNA, and proteins, in the spatial context \cite{moffitt_2022}, generating novel spatially-resolved data that requires highly tailored \textit{in situ} visualizations \cite{dries_2021, Lewis_2021, odonoghue_2021}.
+DNA, RNA, and proteins, in the spatial context\cite{moffitt_2022}, generating novel spatially-resolved data that requires highly tailored \textit{in situ} visualizations\cite{dries_2021, Lewis_2021, odonoghue_2021}.
 Current \textit{in situ} visualizations, as opposed to embedding visualizations (e.g.
@@
 cancerous regions, cortical layers), two spatial maps are displayed side-by-side, one for gene expression and one for spatial domains, creating significant challenges to associate these two information together.
-While software-based interactive visualizations or 3-D visualizations have the potential to mitigate this challenge, they are infeasible for scientific communications in static media, such as prints.
+While software-based interactive visualizations\cite{sriworarat_2023} or 3-D visualizations have the potential to mitigate this challenge, they are infeasible for scientific communications in static media, such as prints.
 Developing a static spatial visualization that enables the simultaneous display of multiple dimensions of information is crucial for spatially-resolved data research.
@@
  To optimize \textit{in situ} visualization, we introduce gestalt principles to the design of spatial maps.
-Gestalt principles \cite{todorovic_2008, palmer_1999} refer to a set of rules describing how humans perceive and interpret visual information and are commonly applied in art and designs.
+Gestalt principles\cite{todorovic_2008, palmer_1999} refer to a set of rules describing how humans perceive and interpret visual information and are commonly applied in art and designs.
 Following gestalt principles, we add additional visual dimensions to spatial maps, leveraging different aesthetics, such as color, fill, and symbols, to simultaneously display disparate variables.
@@
 spatial domains) can be plotted as the backgrounds of the circles, creating a \textit{ground} for the figure.
-For adjacent circles with limited space between them to display the background color, we use an economic implementation, colored outlines for these circles, inspired by watercolor effect \cite{pinna_1987, pinna_2001}.
+For adjacent circles with limited space between them to display the background color, we use an economic implementation, colored outlines for these circles, inspired by watercolor effect\cite{pinna_1987, pinna_2001}.
 Watercolor effect describes the phenomenon in visual perception that surface color arises from thin boundaries and hence is applied here to perceive the background color in tight space.
 Overall, the figure-ground articulation creates two isolated layers in visual perception to display the two variables while maintaining the relative spatial relationship serving as a reference between the two.
-In addition, other fundamental principles \cite{todorovic_2008}, such as proximity, similarity, continuity, and closure, incentivize the brain to group elements and dimensions in the visualization, guaranteeing an integrative perception of the complex multi-dimensional spatial map.
+In addition, other fundamental principles\cite{todorovic_2008}, such as proximity, similarity, continuity, and closure, incentivize the brain to group elements and dimensions in the visualization, guaranteeing an integrative perception of the complex multi-dimensional spatial map.
 
-To implement the proposed visualization, we leverage the state-of-art data visualization framework \texttt{ggplot2} \cite{ggplot2} following the grammar of graphics \cite{wilkinson_2012}.
+To implement the proposed visualization, we leverage the state-of-art data visualization framework \texttt{ggplot2}\cite{ggplot2} following the grammar of graphics\cite{wilkinson_2012}.
 Specifically, we use an interactive process to layer each variable in spatial maps using different aesthetics.
@@
 We provide an open-source package called \coloc (named after the graphic artist M.C.
-Escher) in the R programming environment \cite{R}, providing a simplified interface to navigate the implementation of the multi-dimensional \textit{in situ} visualization.
-By adapting \texttt{ggplot2} standard, \coloc can be seamlessly integrated into many popular spatial resolved toolboxes, such as \texttt{SpatialLIBD}\cite{pardo_2022}, \texttt{Seurat}\cite{hao_2021}, \texttt{Giotto} \cite{dries_2021} to name a few, and allow further theme customization with ease.
-We provide two use cases to examplify some utility of the proposed spatial visualization: 1) the spatially differential gene colocalization in the human dorsolateral prefrontal cortex (DLPFC) using spatial transcriptomics data; 2) multi-dimensional UMAP highlighting differential gene expression in data-driven cell clusters.
+Escher) in the R programming environment\cite{R}, providing a simplified interface to navigate the implementation of the multi-dimensional \textit{in situ} visualization.
+By adapting \texttt{ggplot2} standard, \coloc can be seamlessly integrated into many popular spatial resolved toolboxes, such as \texttt{SpatialLIBD}\cite{pardo_2022}, \texttt{Seurat}\cite{hao_2021}, \texttt{Giotto}\cite{dries_2021} to name a few, and allow further theme customization with ease.
+We provide two use cases to exemplify some utility of the proposed spatial visualization: 1) the spatially differential gene colocalization in the human dorsolateral prefrontal cortex (DLPFC) using spatial transcriptomics data\cite{huukimyers_2023}; 2) multi-dimensional UMAP highlighting differential gene expression in data-driven cell clusters \cite{freytag_2020}.
 

### 2023-03-17 Manuscript/manuscript.tex
revise use-cases.

@@
 \section*{Use Case 1: Spatially Differential Colocalization in DLPFC}
-To demonstrate the new visualization, we apply \coloc to visualize the colocalization of two genes, (FYN) and (EFNA5) in the human dorsolateral prefrontal cortex.
-The two genes are previously identified that relates to SCZ expression via cell-cell communication analysis.
-The purpose of the visualization is to see if the density of colocalization are different across different cortex layers.
-Using \coloc, we can simultaneously display the colocalization and cortex layer in the same spatial plot, instead of having two saprate plots, one for cortex layer and one for colocalization, making the verification of the visual within the spaital context more straightforward and easier.
-
-To create a colocalization plots, \coloc follows a similar principle as \texttt{ggplot2}'s layering strategy.
-In the first step, we will use the funciton \texttt{spCOCOON} to create a base plot that converts a spatial transcroptimcs data structure to a plot object.
-In a second step, we use the function \texttt{add\_fill} to color code each spots with colcalization status, which includes three levels, spots that expressed both genes and spots that expressed either of the genes.
-Then, we will apply the function \texttt{add\_COCOON} to show the spatial domain (cortex layer) information.
-Optionally, one can apply the funciton \texttt{add\_symbol} to display another layer of information.
-Here, we use  the symbol to highlight the colocalization spots, i.e.
-the spots that expresses both genes.
-Because we followed the layering strategy that \texttt{ggplot2}.
-One can seamsely to integrate \coloc to any \texttt{ggplot2}-based spatial visiulization by piping a \texttt{ggplot2} object to \texttt{add\_COCOON} and \texttt{add\_symbol} and customize the plots using the \texttt{ggplot2} syntex and themes, greatly reduce the techinical burden to customize these new spatial plot designs.
+In a recent study investigating the molecular organization of human dorsolateral prefrontal cortex (DLPFC) \cite{huukimyers_2023}, two schizophrenia risk genes, membrane-bound ligand ephrin A5 (\textit{EFNA5}) and ephrin type-A receptor 5 (\textit{EPHA5}), are identified to colocalize via the cell-cell communication analysis.
+In addition, data suggest Layer 6 is the most highly co-localized layer compared to other cortex layers.
+To visually examine the inference, we apply \coloc to create a multi-dimensional \textit{in-situ} spatial map (Figure \label{fig:visual}) that simultaneously exhibits the cortex layers (displayed with color-coded spot outlines) and the categorized colocalization status of \textit{FYN} and \textit{EFNA5} (displayed with color-coded spot fill).
+Compared to the traditional visualization where the cortex layers and the colocalization status are ploted in two side-by-side figures, our proposed visualization enables directly mapping colocalization status to the spatial domain, simplifying the perception of two sources of information and allowing cognitive comparison across cortex layers.
 
 \section*{Use Case 2: Multi-dimensional UMAP}
+The application of the proposed framework doesn't limit to in-situ visualizations of spatially resolved data.
+It can be broadly applied to data mapped to any 2-dimensional coordinate system to simultaneously display multiple variables.
+Such systems include euclidean space (including spatial coordinate as a special case) and data-driven embedding space, for example, UMAP and t-SNE.
+To demonstrate, we apply the proposed visualization to address the challenge of simultaneously displaying cluster membership and gene expression in a single-cell UMAP plot.
+To address the overplotting problem, \texttt{schex} proposed to apply hexagonal binning strategy to unbiasedly display the gene expression.\cite{freytag_2020} In addition, color-coded convex hulls were used to annotate different clusters of cells (Figure \label{fig:visual}).
+Nevertheless, the convex hulls create substantial overlapping areas, creating confusion when interpreting cluster memberships of hexagons in the overlapping areas.
+To improve the interpretability of the visualization, we integrate the proposed visualization to the \texttt{schex} by replacing convex hulls with color-coded hexagons boundaries to avoid possible membership confusion.
+The visual improvement is easily implemented without any modification of \texttt{schex} thanks to the Grammar of Graphics \cite{wilkinson_2012} standard.
 

### 2023-03-17 Manuscript/manuscript.tex
Edit Abstract

@@
 \subsection*{Summary}
-Colocalization is highly interested research questions in the analysis of spatial omics data.
-While many sophisticated computational models are proposed to infer the colocalization relationship between genes or proteins in the spatial context, there completely lack of intuitive spatial visualization, making interpretation extremely challenging.
-The biggest challenge for colocalization visualization is two sources of orthogonal information, spot colocalization and spot membership (such as spatial domain) the same spatial plot.
-To address this issue, we developed \coloc to display the colocalization while maintaining an intuitive representation of the surrounding spatial domain.
-\coloc leverages color contrast optical illusion to display spatial domains to create shadows representing different spatial domain while using different spots to denote colocalization status.
+Visualization is an indispensable component of data analysis and receives growing recognition as essential. Nevertheless, current methods and
+tools are inadequate to visualize the novel spatially-resolved data, limiting the ability to derive new insights and discoveries.
+Specifically, there are no multi-dimensional \textit{in-situ} visualization that simultaneously displays multiple variables in a 2D spatial map.
+To address this problem, we introduce gestalt principles to the design of spatial maps, layering aesthetics to display multiple variables.
+The proposed visualization can be broadly applied to spatially resolved data and 2D embedding methods.
+We provide an open-source R package \coloc, which seamlessly integrates into the state-of-the-art spatial omics toolboxes.
 
 \subsection*{Availability and implementation}
-The R package \coloc is freely available at Github (\url{https://github.com/boyiguo1/VisColoc}).
+The R package \coloc is freely available at Github (\url{https://github.com/boyiguo1/escheR}).
 % and Zenodo (\fixme{add link}).
@@
 Current \textit{in situ} visualizations, as opposed to embedding visualizations (e.g.
-t-SNE and UMAP) that project information to some mathematical space, visualize molecular information in the original cellular location, creating spatial maps that color-code a single variable (e.g., molecule expression, cell type, or spatial domain).
+t-SNE \cite{hinton_2002} and UMAP \cite{becht_2019}) that project information to some mathematical space, visualize molecular information in the original cellular location, creating spatial maps that color-code a single variable (e.g., molecule expression, cell type, or spatial domain).
 This approach lacks the ability to simultaneously exhibit multiple variables, possibly from disparate data domains (such as expression domain and spatial domain) or modalities (such as transcriptomics and proteomics), creating cognitive gaps to interpret findings regarding their (micro-) environment.
@@
 Nevertheless, the convex hulls create substantial overlapping areas, creating confusion when interpreting cluster memberships of hexagons in the overlapping areas.
-To improve the interpretability of the visualization, we integrate the proposed visualization to the \texttt{schex} by replacing convex hulls with color-coded hexagons boundaries to avoid possible membership confusion.
+To improve the interpretability of the visualization, we integrate the proposed visualization to \texttt{schex} by replacing convex hulls with color-coded hexagons boundaries  (Figure \label{fig:visual}) to avoid possible membership confusion.
 The visual improvement is easily implemented without any modification of \texttt{schex} thanks to the Grammar of Graphics \cite{wilkinson_2012} standard.
 
-\section*{Summary}
+\vspace{0.45in}
+In summary, we propose an innovative multi-dimensional spatial visualization to simultaneously display multiple variables in a 2D coordinate system.
+Specifically, our design leverages the gestalt principle from visual perception to introduce other dimensions of a spatial map by iterative layering aesthetics.
+Developed upon \texttt{ggplot2}, we provide an open-source R package \coloc that is seamlessly compatible with popular spatial resolved data analysis toolboxes.
+The proposed visualization has broad applications in visualizing both spot-based and image-based spatial resolved data, spatial multi-omics as well as other spatial-based data.
+The proposed visualization innovation highlights the importance and potential of translating theories in vision science to data science, addressing long-standing challenges in bioinformatics data visualization.
 
-While \coloc is to design to display colocalization in their spatial context, the underlying principle can be broadly applied to address visualizaiton challenges in informaiton rich visualizations.
+% While \coloc is to design to display colocalization in their spatial context, the underlying principle can be broadly applied to address visualizaiton challenges in informaiton rich visualizations.
 The contrast color optical illusion provides allows to add another dimension of information in any spatial visualization, which can be further used to display multi-omics presence, greatly improve the interpretability.
 
-In this manuscript, we demonstrate the utility of the proposed spatial visualization using spot-based spatial resolved data, nevertheless, this method can be applied to image-based spatial resolved data by applying \texttt{schex} to avoid overplotting commonly for image-based spatial resolved data, similar to our second UMAP example where the first step is to reduce done the dimensiona
-
-our programming provides flexibility by following the grammar of graphs, especially ggplot2
-
-translating the findings in vision science to dat science to address many of the current challenges in visualization.
+% our programming provides flexibility by following the grammar of graphs, especially ggplot2
 

### 2023-03-17 Manuscript/manuscript.tex
Replace escheR

@@
 %\setmainfont{Helvetica}
-\title{\coloc: Optimize Spatial Visualization Following Gestalt Principles}
+\title{\texttt{escheR}: Optimize Spatial Visualization Following Gestalt Principles}
 \author[1]{Boyi Guo\thanks{Corresponding author}}
@@
 The proposed visualization can be broadly applied to spatially resolved data and 2D embedding methods.
-We provide an open-source R package \coloc, which seamlessly integrates into the state-of-the-art spatial omics toolboxes.
+We provide an open-source R package \texttt{escheR}, which seamlessly integrates into the state-of-the-art spatial omics toolboxes.
 
 \subsection*{Availability and implementation}
-The R package \coloc is freely available at Github (\url{https://github.com/boyiguo1/escheR}).
+The R package \texttt{escheR} is freely available at Github (\url{https://github.com/boyiguo1/escheR}).
 % and Zenodo (\fixme{add link}).
@@
 When necessary to display an additional layer for a third variable, \texttt{shape} can be used to add symbols such as cross (+) and asterisk (*) to highlight in the spatial map.
-We provide an open-source package called \coloc (named after the graphic artist M.C.
+We provide an open-source package called \texttt{escheR} (named after the graphic artist M.C.
 Escher) in the R programming environment\cite{R}, providing a simplified interface to navigate the implementation of the multi-dimensional \textit{in situ} visualization.
-By adapting \texttt{ggplot2} standard, \coloc can be seamlessly integrated into many popular spatial resolved toolboxes, such as \texttt{SpatialLIBD}\cite{pardo_2022}, \texttt{Seurat}\cite{hao_2021}, \texttt{Giotto}\cite{dries_2021} to name a few, and allow further theme customization with ease.
+By adapting \texttt{ggplot2} standard, \texttt{escheR} can be seamlessly integrated into many popular spatial resolved toolboxes, such as \texttt{SpatialLIBD}\cite{pardo_2022}, \texttt{Seurat}\cite{hao_2021}, \texttt{Giotto}\cite{dries_2021} to name a few, and allow further theme customization with ease.
 We provide two use cases to exemplify some utility of the proposed spatial visualization: 1) the spatially differential gene colocalization in the human dorsolateral prefrontal cortex (DLPFC) using spatial transcriptomics data\cite{huukimyers_2023}; 2) multi-dimensional UMAP highlighting differential gene expression in data-driven cell clusters \cite{freytag_2020}.
@@
 In addition, data suggest Layer 6 is the most highly co-localized layer compared to other cortex layers.
-To visually examine the inference, we apply \coloc to create a multi-dimensional \textit{in-situ} spatial map (Figure \label{fig:visual}) that simultaneously exhibits the cortex layers (displayed with color-coded spot outlines) and the categorized colocalization status of \textit{FYN} and \textit{EFNA5} (displayed with color-coded spot fill).
+To visually examine the inference, we apply \texttt{escheR} to create a multi-dimensional \textit{in-situ} spatial map (Figure \label{fig:visual}) that simultaneously exhibits the cortex layers (displayed with color-coded spot outlines) and the categorized colocalization status of \textit{FYN} and \textit{EFNA5} (displayed with color-coded spot fill).
 Compared to the traditional visualization where the cortex layers and the colocalization status are ploted in two side-by-side figures, our proposed visualization enables directly mapping colocalization status to the spatial domain, simplifying the perception of two sources of information and allowing cognitive comparison across cortex layers.
@@
 Specifically, our design leverages the gestalt principle from visual perception to introduce other dimensions of a spatial map by iterative layering aesthetics.
-Developed upon \texttt{ggplot2}, we provide an open-source R package \coloc that is seamlessly compatible with popular spatial resolved data analysis toolboxes.
+Developed upon \texttt{ggplot2}, we provide an open-source R package \texttt{escheR} that is seamlessly compatible with popular spatial resolved data analysis toolboxes.
 The proposed visualization has broad applications in visualizing both spot-based and image-based spatial resolved data, spatial multi-omics as well as other spatial-based data.
@@
 
-% While \coloc is to design to display colocalization in their spatial context, the underlying principle can be broadly applied to address visualizaiton challenges in informaiton rich visualizations.
+% While \texttt{escheR} is to design to display colocalization in their spatial context, the underlying principle can be broadly applied to address visualizaiton challenges in informaiton rich visualizations.
 The contrast color optical illusion provides allows to add another dimension of information in any spatial visualization, which can be further used to display multi-omics presence, greatly improve the interpretability.
@@
 \section*{Acknowledgements}
-B.G. would like to acknowledge Dr. Leonardo Collado Torres, Louise Huuki-Myers and Leon Di Stefano for their helpful feedback.
+B.G. would like to acknowledge Dr. Leonardo Collado Torres, Louise Huuki-Myers, Dr. Lukas W Weber, and Leon Di Stefano for their helpful feedback.
 

### 2023-03-17 Manuscript/manuscript.tex
add plots, add captions, set up cross-references

@@
 For example, when studying differential expression in different spatial domains (commonly defined by tissue architecture, e.g.
-cancerous regions, cortical layers), two spatial maps are displayed side-by-side, one for gene expression and one for spatial domains, creating significant challenges to associate these two information together.
-While software-based interactive visualizations\cite{sriworarat_2023} or 3-D visualizations have the potential to mitigate this challenge, they are infeasible for scientific communications in static media, such as prints.
+cancerous regions, cortical layers), two spatial maps are displayed side-by-side (Figure \textbf{\ref{fig:visual}A-B}), one for gene expression and one for spatial domains, creating significant challenges to associate these two information together.
+While software-based interactive visualizations\cite{sriworarat_2023} or 3D visualizations have the potential to mitigate this challenge, they are infeasible for scientific communications in static media, such as prints.
 Developing a static spatial visualization that enables the simultaneous display of multiple dimensions of information is crucial for spatially-resolved data research.
@@
 spatial domains) can be plotted as the backgrounds of the circles, creating a \textit{ground} for the figure.
-For adjacent circles with limited space between them to display the background color, we use an economic implementation, colored outlines for these circles, inspired by watercolor effect\cite{pinna_1987, pinna_2001}.
+For adjacent circles with limited space between them to display the background color, we use an economic implementation, colored outlines for these circles (Figure \textbf{\ref{fig:visual}C}), inspired by watercolor effect\cite{pinna_1987, pinna_2001}.
 Watercolor effect describes the phenomenon in visual perception that surface color arises from thin boundaries and hence is applied here to perceive the background color in tight space.
@@
 In addition, data suggest Layer 6 is the most highly co-localized layer compared to other cortex layers.
-To visually examine the inference, we apply \texttt{escheR} to create a multi-dimensional \textit{in-situ} spatial map (Figure \label{fig:visual}) that simultaneously exhibits the cortex layers (displayed with color-coded spot outlines) and the categorized colocalization status of \textit{FYN} and \textit{EFNA5} (displayed with color-coded spot fill).
-Compared to the traditional visualization where the cortex layers and the colocalization status are ploted in two side-by-side figures, our proposed visualization enables directly mapping colocalization status to the spatial domain, simplifying the perception of two sources of information and allowing cognitive comparison across cortex layers.
+To visually examine the inference, we apply \texttt{escheR} to create a multi-dimensional \textit{in-situ} spatial map (Figure \textbf{\ref{fig:visual}A-B}) that simultaneously exhibits the cortex layers (displayed with color-coded spot outlines) and the categorized colocalization status of \textit{FYN} and \textit{EFNA5} (displayed with color-coded spot fill).
+Compared to the traditional visualization where the cortex layers and the colocalization status are ploted in two side-by-side figures, our proposed visualization (Figure \textbf{\ref{fig:visual}D}) enables directly mapping colocalization status to the spatial domain, simplifying the perception of two sources of information and allowing cognitive comparison across cortex layers.
 
@@
 To demonstrate, we apply the proposed visualization to address the challenge of simultaneously displaying cluster membership and gene expression in a single-cell UMAP plot.
-To address the overplotting problem, \texttt{schex} proposed to apply hexagonal binning strategy to unbiasedly display the gene expression.\cite{freytag_2020} In addition, color-coded convex hulls were used to annotate different clusters of cells (Figure \label{fig:visual}).
+To address the overplotting problem, \texttt{schex} proposed to apply hexagonal binning strategy to unbiasedly display the gene expression.\cite{freytag_2020} In addition, color-coded convex hulls were used to annotate different clusters of cells (Figure \textbf{\ref{fig:visual}E}).
 Nevertheless, the convex hulls create substantial overlapping areas, creating confusion when interpreting cluster memberships of hexagons in the overlapping areas.
-To improve the interpretability of the visualization, we integrate the proposed visualization to \texttt{schex} by replacing convex hulls with color-coded hexagons boundaries  (Figure \label{fig:visual}) to avoid possible membership confusion.
+To improve the interpretability of the visualization, we integrate the proposed visualization to \texttt{schex} by replacing convex hulls with color-coded hexagons boundaries  (Figure \textbf{\ref{fig:visual}F}) to avoid possible membership confusion.
 The visual improvement is easily implemented without any modification of \texttt{schex} thanks to the Grammar of Graphics \cite{wilkinson_2012} standard.
@@
 The UMAP example follows the tutorial from \texttt{schex} (\url{https://www.bioconductor.org/packages/release/bioc/vignettes/schex/inst/doc/using_schex.html}).
-The code that generates these figures are deposited at \fixme{github manuscript repo}.
+The code that generates these figures is deposited at \url{https://github.com/boyiguo1/Manuscript_escheR}.
 
 \clearpage
-% \vspace{-0.1in}
-\begin{wrapfigure}[13]{r}{0.8\textwidth}
-\vspace{-0.3in}
+
+\begin{figure}[13]
 \begin{center}
-\includegraphics[width=0.78\textwidth]{figure/new_visiual.png}
+\includegraphics[width=0.78\textwidth]{Manuscript/figure/fig_grid.png}
 \end{center}
-\vspace{-0.35in}
-\caption{\footnotesize \textbf{Optimized visualization interpreting co-localization in the spatial context}.
-(\textbf{A}) Current visualization of co-localization requires two separate plots, one to display spatial domain information and one to display co-localization, difficulty to associate the two sources of information for interpretation.
-(\textbf{B}) The proposed visualization is optimized to display both spatial domain and co-localization information in one plot.}
+\caption{\footnotesize \textbf{\texttt{escheR} improves spatial visualizations following the gestalt principles}.
+(\textbf{A-B}) The traditional visualization displays the cortex layers (spatial domains) (\textbf{A}) and the colocalization status (\textbf{B}) in a side-by-side figure, creating challenges to connecting colocalization status to spatial domains.
+(\textbf{C}) The watercolor effect enables displaying spatial domains by color-coding only outlines of circles.
+(\textbf{D}) \texttt{escheR} enables the multi-dimensional \texttt{in-situ} visualization that simultaneously displays the cortex layers and the colocalization status, substantially improving interpretability.
+(\textbf{E}) \texttt{schex} creates overlapped convex hulls, creating confusion when interpreting cluster memberships of hexagons in the overlapping areas.
+(\textbf{F}) \texttt{escheR} plots hexagon-specific membership to improve interpretability.}
 \label{fig:visual} 
-\end{wrapfigure}
+\end{figure}
 

### 2023-03-18 Manuscript/manuscript.tex
Make Figure larger
Stephanie's Revision

@@
-\documentclass[11pt]{article}
+\documentclass[10pt,twocolumn]{article}
 \usepackage[margin=0.6in]{geometry}
@@
 
+\usepackage{hyperref}
+\definecolor{darkblue}{rgb}{0.0,0.0,0.75}
+\hypersetup{colorlinks,breaklinks,
+            linkcolor=darkblue,urlcolor=darkblue,
+            anchorcolor=darkblue,citecolor=darkblue}
+
 \newcommand{\fixme}[1]{{\color{red} (#1)}}
@@
 %\setmainfont{Helvetica}
-\title{\texttt{escheR}: Optimize Spatial Visualization Following Gestalt Principles}
-\author[1]{Boyi Guo\thanks{Corresponding author}}
-\author[1]{Stephanie C. Hicks}
+\title{\texttt{escheR}: Unified multi-dimensional visualizations with Gestalt principles}
+\author[1]{Boyi Guo}
+\author[1,2]{Stephanie C. Hicks\thanks{Corresponding author}}
 \affil[1]{Department of Biostatistics, Johns Hopkins Bloomberg School of Public Health, MD, USA}
+\affil[2]{Malone Center for Engineering in Healthcare, Johns Hopkins University, MD, USA}
 \date{\today}
@@
 \section*{Abstract}
-\subsection*{Summary}
-Visualization is an indispensable component of data analysis and receives growing recognition as essential. Nevertheless, current methods and
+The creation of effective visualizations is a fundamental component of data analysis. Nevertheless, current methods and
 tools are inadequate to visualize the novel spatially-resolved data, limiting the ability to derive new insights and discoveries.
@@
 The proposed visualization can be broadly applied to spatially resolved data and 2D embedding methods.
-We provide an open-source R package \texttt{escheR}, which seamlessly integrates into the state-of-the-art spatial omics toolboxes.
+We provide an open source R package \texttt{escheR}, which seamlessly integrates into the state-of-the-art spatial omics toolboxes.
 
 \subsection*{Availability and implementation}
-The R package \texttt{escheR} is freely available at Github (\url{https://github.com/boyiguo1/escheR}).
+The open source R package \texttt{escheR} is freely available on GitHub and is being submitted to Bioconductor (\href{https://github.com/boyiguo1/escheR}{https://github.com/boyiguo1/escheR}).
 % and Zenodo (\fixme{add link}).
@@
 %% INTRODUCTION ============================================================================================
-\section*{Introduction}
-Visualization is an indispensable component of data analysis, providing clarity that connects quantitative evidence to key conclusions.\cite{dagostinomcgowan_2022} In bioinformatics research, visualization receives growing recognition as essential: many scientists rely on visualization to complete their cognitive process from analysis to insight, including analytic validation of automated pipelines and scientific communication.\cite{odonoghue_2021} Nevertheless, current methods and tools are inadequate to visualize the growingly complex biological data and computational models, limiting the ability to derive new insights and discoveries.\cite{odonoghue_2010}
+\section{Introduction}
+Visualization is an indispensable component of data analysis, providing clarity that connects quantitative evidence to key conclusions \cite{dagostinomcgowan_2022}.
+In biomedical research, visualization receives growing recognition as essential: many scientists rely on visualization to complete their cognitive process from analysis to insight, including analytic validation of automated pipelines and scientific communication \cite{odonoghue_2021}.
+However, an important challenge in biomedical resesearch is how to visualize increasingly complex, multi-dimensional data \cite{odonoghue_2010}.
 
-The recent technology breakthrough in spatial profiling enables characterizing the diversity of molecules, e.g.
-DNA, RNA, and proteins, in the spatial context\cite{moffitt_2022}, generating novel spatially-resolved data that requires highly tailored \textit{in situ} visualizations\cite{dries_2021, Lewis_2021, odonoghue_2021}.
-Current \textit{in situ} visualizations, as opposed to embedding visualizations (e.g.
-t-SNE \cite{hinton_2002} and UMAP \cite{becht_2019}) that project information to some mathematical space, visualize molecular information in the original cellular location, creating spatial maps that color-code a single variable (e.g., molecule expression, cell type, or spatial domain).
-This approach lacks the ability to simultaneously exhibit multiple variables, possibly from disparate data domains (such as expression domain and spatial domain) or modalities (such as transcriptomics and proteomics), creating cognitive gaps to interpret findings regarding their (micro-) environment.
-For example, when studying differential expression in different spatial domains (commonly defined by tissue architecture, e.g.
-cancerous regions, cortical layers), two spatial maps are displayed side-by-side (Figure \textbf{\ref{fig:visual}A-B}), one for gene expression and one for spatial domains, creating significant challenges to associate these two information together.
-While software-based interactive visualizations\cite{sriworarat_2023} or 3D visualizations have the potential to mitigate this challenge, they are infeasible for scientific communications in static media, such as prints.
-Developing a static spatial visualization that enables the simultaneous display of multiple dimensions of information is crucial for spatially-resolved data research.
+Here, we focus on two types of visualizations in biomedical research, but note that the proposed ideas could be extended beyond these applications: (i) embedding visualizations, which project data into some low-dimensional embedding or mathematical space (e.g.
+Principal Components Analysis (PCA) \fixme{add citation}, $t$-SNE \cite{hinton_2002} \fixme{spell out}, or UMAP \cite{becht_2019} \fixme{spell out}) and (ii) \textit{in situ} visualizations \cite{dries_2021, Lewis_2021, odonoghue_2021}, which aim to visualize molecules captured from \textit{in situ} imaging or sequencing technologies where \textit{in situ} refers to `in its original place'.
+Both of these visualizations represent data in a 2D space and are motivated by recent advances in experimental technologies that profile molecules, including DNA, RNA, and proteins, at a single-cell or spatial resolution \fixme{add citations} \cite{moffitt_2022}.
+These technologies are commonly referred to as single-cell/single-nucleus RNA-sequencing (sc/snRNA-seq) \fixme{add citations} and \textit{in situ} spatially-resolved transcriptomics (SRT) \fixme{add citations}.
 
-% Previous data visualization of omics largely relies on projecting the omics space to an embedding space, such as t-SNE and UMAP, which is infeasible for spatial omics technologies.
+A common and fundamental challenge with both of these visualizations is how to visualize multi-dimensional information in a 2D space.
+For example, in \textit{in situ} visualizations, we often want to create a spatial map to visualize a continuous (e.g.
+gene expression) or discrete (e.g.
+cell type or spatial domain) variable representing molecular information in the original spatial location.
+However, it is challenging to simultaneously visualize multi-dimensional data, such as information from disparate data domains (such as expression domain and spatial domain) or disparate data modalities (such as transcriptomics and proteomics) in the same plot.
+Currently, best practices for this include making two different plots displayed side-by-side (\textbf{Figure \textbf{\ref{fig:visual}A-B}}), one for gene expression and one for spatial domains.
+This creates cognitive gaps on how to associate the disparate information or how to interpret the biological findings of this multi-dimensional information regarding their (micro-)environment or colocalization.
+While interactive visualizations \cite{sriworarat_2023} \fixme{add more citations} or 3D visualizations \fixme{add citations} have the potential to mitigate this challenge, they are infeasible for scientific communications in static media, such as printed work.
+Developing a static and unified visualization that enables the simultaneous display of multiple dimensions of information is crucial for biomedical research.
 
- To optimize \textit{in situ} visualization, we introduce gestalt principles to the design of spatial maps.
+\begin{figure*}[!t]
+\begin{center}
+\includegraphics[width=\textwidth]{Manuscript/figure/fig_grid.png}
+\caption{\small \textbf{\texttt{escheR} improves spatial visualizations following the Gestalt principles}.
+(\textbf{A-B}) The traditional visualization displays the cortex layers (spatial domains) (\textbf{A}) and the colocalization status (\textbf{B}) in a side-by-side figure, creating challenges to connecting colocalization status to spatial domains.
+(\textbf{C}) The watercolor effect enables displaying spatial domains by color-coding only outlines of circles.
+(\textbf{D}) \texttt{escheR} enables the multi-dimensional \texttt{in-situ} visualization that simultaneously displays the cortex layers and the colocalization status, substantially improving interpretability.
+(\textbf{E}) \texttt{schex} creates overlapped convex hulls, creating confusion when interpreting cluster memberships of hexagons in the overlapping areas.
+(\textbf{F}) \texttt{escheR} plots hexagon-specific membership to improve interpretability.}
+\label{fig:visual} 
+\end{center}
+\end{figure*}
+
+\section{Results}
+
+To address these challenges, here we leverage the Gestalt (German for “unified whole”) principles for design \cite{todorovic_2008, palmer_1999} as a way to visualize multi-dimensional data in 2D visualizations.
+We focus on the two types of data visualizations previously introduced that are widely used in biomedical research: (i) embedding visualizations and (ii) \textit{in situ} visualizations.
+We provide an R package, \texttt{escheR}, implementing these ideas, which is built on the state-of-the-art data visualization framework \texttt{ggplot2} in the R programming language.
+Finally, we comment on how these ideas could be extended to other types of visualization in biomedical research.
+
+\subsection{Multi-dimensional 2D visualizations with \texttt{ggplot2} and Gestalt principles}
+
 Gestalt principles\cite{todorovic_2008, palmer_1999} refer to a set of rules describing how humans perceive and interpret visual information and are commonly applied in art and designs.
-Following gestalt principles, we add additional visual dimensions to spatial maps, leveraging different aesthetics, such as color, fill, and symbols, to simultaneously display disparate variables.
-Specifically, we apply figure-ground articulation to display two variables: one variable (e.g.
+Developed in the 1920s by German psychologists Max Wertheimer, Kurt Koffka and Wolfgang Kohler, these principles helps humans perceive a set of individual elements as a whole.
+
+Here, we leverage the principles to be able to visualize multi-dimensional data in a unified 2D plot.
+Our approach is to use the state-of-art data visualization framework \texttt{ggplot2} \cite{ggplot2} following the grammar of graphics \cite{wilkinson_2012} and map individual variables to different aesthetics to simultaneously display disparate variables.
+Specifically, we apply a figure-ground articulation to display two variables: one variable (e.g.
 expression) can be plotted as color-filled circles, serving as the \textit{figure}; one variable (e.g.
 spatial domains) can be plotted as the backgrounds of the circles, creating a \textit{ground} for the figure.
-For adjacent circles with limited space between them to display the background color, we use an economic implementation, colored outlines for these circles (Figure \textbf{\ref{fig:visual}C}), inspired by watercolor effect\cite{pinna_1987, pinna_2001}.
+In practice, we use the combination of \texttt{color} and \texttt{fill="transparent"} to create the background layer and \texttt{fill} to create the figure layer.
+When necessary to display an additional layer for a third variable, \texttt{shape} can be used to add symbols such as cross (+) and asterisk (*) to highlight in the spatial map.
+
+For adjacent circles with limited space between them to display the background color, we use an economic implementation, colored outlines for these circles (\textbf{Figure~\ref{fig:visual}C}), inspired by watercolor effect \cite{pinna_1987, pinna_2001}.
 Watercolor effect describes the phenomenon in visual perception that surface color arises from thin boundaries and hence is applied here to perceive the background color in tight space.
 Overall, the figure-ground articulation creates two isolated layers in visual perception to display the two variables while maintaining the relative spatial relationship serving as a reference between the two.
-In addition, other fundamental principles\cite{todorovic_2008}, such as proximity, similarity, continuity, and closure, incentivize the brain to group elements and dimensions in the visualization, guaranteeing an integrative perception of the complex multi-dimensional spatial map.
+In addition, other fundamental principles \cite{todorovic_2008}, such as proximity, similarity, continuity, and closure, incentivize the brain to group elements and dimensions in the visualization, guaranteeing an integrative perception of the complex multi-dimensional spatial map.
 
-To implement the proposed visualization, we leverage the state-of-art data visualization framework \texttt{ggplot2}\cite{ggplot2} following the grammar of graphics\cite{wilkinson_2012}.
-Specifically, we use an interactive process to layer each variable in spatial maps using different aesthetics.
-For example, we use the combination of \texttt{color} and \texttt{fill="transparent"} to create the background layer and \texttt{fill} to create the figure layer.
-When necessary to display an additional layer for a third variable, \texttt{shape} can be used to add symbols such as cross (+) and asterisk (*) to highlight in the spatial map.
-We provide an open-source package called \texttt{escheR} (named after the graphic artist M.C.
-Escher) in the R programming environment\cite{R}, providing a simplified interface to navigate the implementation of the multi-dimensional \textit{in situ} visualization.
+Here, we provide an open-source package called \texttt{escheR} (named after the graphic artist M.C.
+Escher) in the R programming language \cite{R}, leading to a simplified interface to navigate the implementation of the multi-dimensional visualization in 2D space.
 By adapting \texttt{ggplot2} standard, \texttt{escheR} can be seamlessly integrated into many popular spatial resolved toolboxes, such as \texttt{SpatialLIBD}\cite{pardo_2022}, \texttt{Seurat}\cite{hao_2021}, \texttt{Giotto}\cite{dries_2021} to name a few, and allow further theme customization with ease.
-We provide two use cases to exemplify some utility of the proposed spatial visualization: 1) the spatially differential gene colocalization in the human dorsolateral prefrontal cortex (DLPFC) using spatial transcriptomics data\cite{huukimyers_2023}; 2) multi-dimensional UMAP highlighting differential gene expression in data-driven cell clusters \cite{freytag_2020}.
 
-\section*{Use Case 1: Spatially Differential Colocalization in DLPFC}
-In a recent study investigating the molecular organization of human dorsolateral prefrontal cortex (DLPFC) \cite{huukimyers_2023}, two schizophrenia risk genes, membrane-bound ligand ephrin A5 (\textit{EFNA5}) and ephrin type-A receptor 5 (\textit{EPHA5}), are identified to colocalize via the cell-cell communication analysis.
-In addition, data suggest Layer 6 is the most highly co-localized layer compared to other cortex layers.
-To visually examine the inference, we apply \texttt{escheR} to create a multi-dimensional \textit{in-situ} spatial map (Figure \textbf{\ref{fig:visual}A-B}) that simultaneously exhibits the cortex layers (displayed with color-coded spot outlines) and the categorized colocalization status of \textit{FYN} and \textit{EFNA5} (displayed with color-coded spot fill).
-Compared to the traditional visualization where the cortex layers and the colocalization status are ploted in two side-by-side figures, our proposed visualization (Figure \textbf{\ref{fig:visual}D}) enables directly mapping colocalization status to the spatial domain, simplifying the perception of two sources of information and allowing cognitive comparison across cortex layers.
+Next, we give two use cases to exemplify some utility of the proposed spatial visualization: (i) the spatially differential gene colocalization in the human dorsolateral prefrontal cortex (DLPFC) using spatial transcriptomics data \cite{huukimyers_2023}; (ii) multi-dimensional UMAP highlighting differential gene expression in data-driven cell clusters \cite{freytag_2020}.
 
-\section*{Use Case 2: Multi-dimensional UMAP}
-The application of the proposed framework doesn't limit to in-situ visualizations of spatially resolved data.
-It can be broadly applied to data mapped to any 2-dimensional coordinate system to simultaneously display multiple variables.
-Such systems include euclidean space (including spatial coordinate as a special case) and data-driven embedding space, for example, UMAP and t-SNE.
-To demonstrate, we apply the proposed visualization to address the challenge of simultaneously displaying cluster membership and gene expression in a single-cell UMAP plot.
-To address the overplotting problem, \texttt{schex} proposed to apply hexagonal binning strategy to unbiasedly display the gene expression.\cite{freytag_2020} In addition, color-coded convex hulls were used to annotate different clusters of cells (Figure \textbf{\ref{fig:visual}E}).
-Nevertheless, the convex hulls create substantial overlapping areas, creating confusion when interpreting cluster memberships of hexagons in the overlapping areas.
-To improve the interpretability of the visualization, we integrate the proposed visualization to \texttt{schex} by replacing convex hulls with color-coded hexagons boundaries  (Figure \textbf{\ref{fig:visual}F}) to avoid possible membership confusion.
-The visual improvement is easily implemented without any modification of \texttt{schex} thanks to the Grammar of Graphics \cite{wilkinson_2012} standard.
+\subsection{Multi-dimensional \textit{in situ} visualization}
+In a recent study investigating the molecular organization of human dorsolateral prefrontal cortex (DLPFC) \cite{huukimyers_2023}, two schizophrenia risk genes, membrane-bound ligand ephrin A5 (\textit{EFNA5}) and ephrin type-A receptor 5 (\textit{EPHA5}), were identified to colocalize via the cell-cell communication analysis.
+In addition, data suggested Layer 6 was the most highly co-localized layer compared to other cortex layers.
+To visually examine the inference, we applied \texttt{escheR} to create a multi-dimensional \textit{in situ} spatial map (\textbf{Figure \ref{fig:visual}A-B}) that simultaneously exhibits the cortex layers (displayed with color-coded spot outlines) and the categorized colocalization status of genes \textit{FYN} and \textit{EFNA5} (displayed with color-coded spot fill).
+Compared to the traditional visualization where the cortex layers and the colocalization status are visualized in two side-by-side figures, our proposed visualization (\textbf{Figure \ref{fig:visual}D}) enables directly mapping colocalization status to the spatial domain, simplifying the perception of two sources of information and allowing cognitive comparison across cortex layers.
 
-\vspace{0.45in}
+\subsection{Multi-dimensional embedding visualizations}
+
+The application of the proposed framework is not limited to \textit{in situ} visualizations of spatially-resolved transcriptomics data.
+It is broadly applicable to data mapped to any 2-dimensional coordinate system to simultaneously display multiple variables.
+Such systems include euclidean space (including spatial coordinate as a special case) and data-driven embedding space, for example, UMAP and $t$-SNE.
+To demonstrate, we applied the proposed visualization to address the challenge of simultaneously displaying cluster membership and gene expression in a single-cell UMAP plot.
+To address the overplotting problem, previous work proposed to apply hexagonal binning strategy to display the gene expression \cite{freytag_2020}.
+Here, the color-coded convex hulls are used to annotate different clusters of cells (\textbf{Figure~\ref{fig:visual}E}).
+However, the convex hulls create substantial overlapping areas, creating confusion when interpreting cluster memberships of hexagons in the overlapping areas.
+To improve the interpretability of the visualization, we replace the convex hulls with color-coded hexagons boundaries  (\textbf{Figure~\ref{fig:visual}F}) to avoid possible membership confusion.
+We note that our contribution to improve the visualization is easily implemented without any modification of \texttt{schex} as both are built upon the Grammar of Graphics \cite{wilkinson_2012} standard.
+
+\section{Discussion}
 In summary, we propose an innovative multi-dimensional spatial visualization to simultaneously display multiple variables in a 2D coordinate system.
@@
 %% EPILOGUE ================================================================
-\section*{Acknowledgements}
-B.G. would like to acknowledge Dr. Leonardo Collado Torres, Louise Huuki-Myers, Dr. Lukas W Weber, and Leon Di Stefano for their helpful feedback.
+\section*{Abbreviations}
 
-\vspace{0.2in}
-\noindent \textit{Conflict of Interest:} none declared.
+\fixme{add any abbreviations here}
 
-\section*{Data availability}
-The spatial transcriptomics dataset was obtained from \texttt{spatialLIBD} (\url{http://research.libd.org/spatialLIBD/}).
-The UMAP example follows the tutorial from \texttt{schex} (\url{https://www.bioconductor.org/packages/release/bioc/vignettes/schex/inst/doc/using_schex.html}).
-The code that generates these figures is deposited at \url{https://github.com/boyiguo1/Manuscript_escheR}.
+\section*{Author contributions}
 
-\clearpage
+\begin{itemize}[nosep]
+    \item Conceptualization: BG
+    \item Data Curation: BG
+    \item \fixme{add more here} use credit system (https://www.elsevier.com/authors/policies-and-guidelines/credit-author-statement)
+\end{itemize}
 
-\begin{figure}[13]
-\begin{center}
-\includegraphics[width=0.78\textwidth]{Manuscript/figure/fig_grid.png}
-\end{center}
-\caption{\footnotesize \textbf{\texttt{escheR} improves spatial visualizations following the gestalt principles}.
-(\textbf{A-B}) The traditional visualization displays the cortex layers (spatial domains) (\textbf{A}) and the colocalization status (\textbf{B}) in a side-by-side figure, creating challenges to connecting colocalization status to spatial domains.
-(\textbf{C}) The watercolor effect enables displaying spatial domains by color-coding only outlines of circles.
-(\textbf{D}) \texttt{escheR} enables the multi-dimensional \texttt{in-situ} visualization that simultaneously displays the cortex layers and the colocalization status, substantially improving interpretability.
-(\textbf{E}) \texttt{schex} creates overlapped convex hulls, creating confusion when interpreting cluster memberships of hexagons in the overlapping areas.
-(\textbf{F}) \texttt{escheR} plots hexagon-specific membership to improve interpretability.}
-\label{fig:visual} 
-\end{figure}
+\section*{Declarations}
+
+\subsection*{Ethics approval and consent to participate}
+Not applicable.
+
+\subsection*{Competing interests}
+The authors declare that they have no competing interests.
+
+\subsection*{Availability of data and materials}
+The spatial transcriptomics dataset was obtained from \texttt{spatialLIBD} (\href{http://research.libd.org/spatialLIBD}{research.libd.org/spatialLIBD}).
+The UMAP example follows the `using\_schex' vignette in the \texttt{schex} package (\href{https://www.bioconductor.org/packages/schex}{bioconductor.org/packages/schex}).
+The code that generates these figures is deposited at \href{https://github.com/boyiguo1/Manuscript_escheR}{github.com/boyiguo1/Manuscript\_escheR} \fixme{version XXX}.
+
+\subsection*{Funding}
+This project was supported by the National Institute of Mental Health R01MH126393 (BG, SCH), and CZF2019-002443 (SCH) from the Chan Zuckerberg Initiative DAF, an advised fund of Silicon Valley Community Foundation.
+All funding bodies had no role in the design of the study and collection, analysis, and interpretation of data and in writing the manuscript.
+
+\subsection*{Acknowledgements}
+BG would like to acknowledge Dr.
+Leonardo Collado-Torres, Louise Huuki-Myers, Dr.
+Lukas M.
+Weber, and Leon Di Stefano for their helpful comments, feedback and suggestions on \texttt{escheR} functionality.
+
+\subsection*{Author’s information}
+
+\begin{itemize}[nosep]
+    \item Boyi Guo (ORCiD: \href{https://orcid.org/0000-0003-2950-2349}{0000-0003-2950-2349})
+    \item Stephanie C. Hicks (ORCiD: \href{https://orcid.org/0000-0002-7858-0231}{0000-0002-7858-0231})
+\end{itemize}
+
+\subsection*{Conflict of Interest} 
+None declared.
 
@@
 
+\clearpage
+\onecolumn
+
+{\huge Supplementary Materials}
+
+\hrule
+
+\vspace*{0.5cm}
+
+\begin{center}
+
+{\Large \texttt{escheR}: Unified multi-dimensional visualizations with Gestalt principles}
+
+\vspace*{0.75cm}
+
+{\large Boyi Guo, Stephanie C.\ Hicks$^*$}
+
+\vspace*{0.3cm}
+
+{\small $^*$Correspondence to \url{shicks19@jhu.edu}}
+
+\end{center}
+
+\renewcommand{\figurename}{Supplementary Figure}
+\renewcommand{\tablename}{Supplementary Table}
+\setcounter{figure}{0}
+\setcounter{table}{0}
+\setcounter{section}{0}
+\setcounter{page}{1}
+\makeatletter
+\renewcommand{\thefigure}{S\@arabic\c@figure}
+\renewcommand{\thetable}{S\@arabic\c@table}
+\renewcommand{\thesection}{Supplemental Note S\@arabic\c@section}
+\makeatother
+
+\vspace*{1cm}
+
+{\bf \large Contents}
+
+\begin{enumerate}
+    \item \textbf{Supplemental Figures~S1.}
+\end{enumerate}
+
+\clearpage
+
+\noindent {\LARGE Supplemental Figures}
+
+\begin{figure*}[!h]
+\begin{center}
+\includegraphics[width=\textwidth]{Manuscript/figure/fig_grid.png}
+\caption{\small \textbf{\texttt{escheR} improves spatial visualizations following the Gestalt principles}.
+(\textbf{A-B}) The traditional visualization displays the cortex layers (spatial domains) (\textbf{A}) and the colocalization status (\textbf{B}) in a side-by-side figure, creating challenges to connecting colocalization status to spatial domains.
+(\textbf{C}) The watercolor effect enables displaying spatial domains by color-coding only outlines of circles.
+(\textbf{D}) \texttt{escheR} enables the multi-dimensional \texttt{in-situ} visualization that simultaneously displays the cortex layers and the colocalization status, substantially improving interpretability.
+(\textbf{E}) \texttt{schex} creates overlapped convex hulls, creating confusion when interpreting cluster memberships of hexagons in the overlapping areas.
+(\textbf{F}) \texttt{escheR} plots hexagon-specific membership to improve interpretability.}
+\label{fig:embedding} 
+\end{center}
+\end{figure*}
+
 \end{document}

### 2023-03-18 Manuscript/manuscript.tex
Revisions

@@
 \author[1]{Boyi Guo}
-\author[1,2]{Stephanie C. Hicks\thanks{Corresponding author}}
+\author[1,2]{Stephanie C. Hicks\thanks{$^*$Correspondence to \url{shicks19@jhu.edu}}}
 \affil[1]{Department of Biostatistics, Johns Hopkins Bloomberg School of Public Health, MD, USA}
@@
 In biomedical research, visualization receives growing recognition as essential: many scientists rely on visualization to complete their cognitive process from analysis to insight, including analytic validation of automated pipelines and scientific communication \cite{odonoghue_2021}.
-However, an important challenge in biomedical resesearch is how to visualize increasingly complex, multi-dimensional data \cite{odonoghue_2010}.
+However, an important challenge in biomedical research is how to visualize increasingly complex, multi-dimensional data \cite{odonoghue_2010}.
 
 Here, we focus on two types of visualizations in biomedical research, but note that the proposed ideas could be extended beyond these applications: (i) embedding visualizations, which project data into some low-dimensional embedding or mathematical space (e.g.
-Principal Components Analysis (PCA) \fixme{add citation}, $t$-SNE \cite{hinton_2002} \fixme{spell out}, or UMAP \cite{becht_2019} \fixme{spell out}) and (ii) \textit{in situ} visualizations \cite{dries_2021, Lewis_2021, odonoghue_2021}, which aim to visualize molecules captured from \textit{in situ} imaging or sequencing technologies where \textit{in situ} refers to `in its original place'.
-Both of these visualizations represent data in a 2D space and are motivated by recent advances in experimental technologies that profile molecules, including DNA, RNA, and proteins, at a single-cell or spatial resolution \fixme{add citations} \cite{moffitt_2022}.
+Principal Components Analysis (PCA) \cite{hotelling_1933}, $t$-distributed Stochastic Neighbor Embedding ($t$-SNE) \cite{vandermaaten_2008}, or Uniform Manifold Approximation and Projection (UMAP) \cite{becht_2019}) and (ii) \textit{in situ} visualizations \cite{dries_2021, Lewis_2021, odonoghue_2021}, which aim to visualize molecules captured from \textit{in situ} imaging or sequencing technologies where \textit{in situ} refers to `in its original place'.
+Both of these visualizations represent data in a 2D space and are motivated by recent advances in experimental technologies that profile molecules, including DNA, RNA, and proteins, at a single-cell or spatial resolution \cite{kashima_2020, moffitt_2022}.
 These technologies are commonly referred to as single-cell/single-nucleus RNA-sequencing (sc/snRNA-seq) \fixme{add citations} and \textit{in situ} spatially-resolved transcriptomics (SRT) \fixme{add citations}.
@@
 However, it is challenging to simultaneously visualize multi-dimensional data, such as information from disparate data domains (such as expression domain and spatial domain) or disparate data modalities (such as transcriptomics and proteomics) in the same plot.
-Currently, best practices for this include making two different plots displayed side-by-side (\textbf{Figure \textbf{\ref{fig:visual}A-B}}), one for gene expression and one for spatial domains.
+Currently, best practices for this include making two different plots displayed side-by-side (\fixme{\textbf{Figure \textbf{\ref{fig:visual}A-B}}}), one for gene expression and one for spatial domains.
 This creates cognitive gaps on how to associate the disparate information or how to interpret the biological findings of this multi-dimensional information regarding their (micro-)environment or colocalization.
-While interactive visualizations \cite{sriworarat_2023} \fixme{add more citations} or 3D visualizations \fixme{add citations} have the potential to mitigate this challenge, they are infeasible for scientific communications in static media, such as printed work.
+While interactive visualizations \cite{keller_2021, pardo_2022, sriworarat_2023} have the potential to mitigate this challenge, they are infeasible for scientific communications in static media, such as printed work.
 Developing a static and unified visualization that enables the simultaneous display of multiple dimensions of information is crucial for biomedical research.
@@
 \begin{center}
-\includegraphics[width=\textwidth]{Manuscript/figure/fig_grid.png}
+\includegraphics[width=\textwidth]{Manuscript/figure/insitu.jpg}
 \caption{\small \textbf{\texttt{escheR} improves spatial visualizations following the Gestalt principles}.
-(\textbf{A-B}) The traditional visualization displays the cortex layers (spatial domains) (\textbf{A}) and the colocalization status (\textbf{B}) in a side-by-side figure, creating challenges to connecting colocalization status to spatial domains.
+(\textbf{A-B}) The traditional visualization displays the colocalization plot (\textbf{A}) and the cortex layers (spatial domains) plot (\textbf{B}) side-by-side, creating challenges to connecting colocalization status to spatial domains.
 (\textbf{C}) The watercolor effect enables displaying spatial domains by color-coding only outlines of circles.
-(\textbf{D}) \texttt{escheR} enables the multi-dimensional \texttt{in-situ} visualization that simultaneously displays the cortex layers and the colocalization status, substantially improving interpretability.
-(\textbf{E}) \texttt{schex} creates overlapped convex hulls, creating confusion when interpreting cluster memberships of hexagons in the overlapping areas.
-(\textbf{F}) \texttt{escheR} plots hexagon-specific membership to improve interpretability.}
+(\textbf{D}) \texttt{escheR} enables the multi-dimensional \texttt{in-situ} visualization that simultaneously displays the cortex layers and the colocalization status, substantially improving interpretability.}
 \label{fig:visual} 
@@
 Gestalt principles\cite{todorovic_2008, palmer_1999} refer to a set of rules describing how humans perceive and interpret visual information and are commonly applied in art and designs.
-Developed in the 1920s by German psychologists Max Wertheimer, Kurt Koffka and Wolfgang Kohler, these principles helps humans perceive a set of individual elements as a whole.
+Developed in the 1920s by German psychologists Max Wertheimer, Kurt Koffka and Wolfgang Kohler, these principles help humans perceive a set of individual elements as a whole.
 
@@
 Our approach is to use the state-of-art data visualization framework \texttt{ggplot2} \cite{ggplot2} following the grammar of graphics \cite{wilkinson_2012} and map individual variables to different aesthetics to simultaneously display disparate variables.
-Specifically, we apply a figure-ground articulation to display two variables: one variable (e.g.
+Specifically, we apply the figure-ground segmentation \cite{peterson_2010} in displaying two variables: one variable (e.g.
 expression) can be plotted as color-filled circles, serving as the \textit{figure}; one variable (e.g.
@@
 
-For adjacent circles with limited space between them to display the background color, we use an economic implementation, colored outlines for these circles (\textbf{Figure~\ref{fig:visual}C}), inspired by watercolor effect \cite{pinna_1987, pinna_2001}.
+For adjacent circles with limited space between them to display the background color, we use an economic implementation, colored outlines for these circles (\fixme{\textbf{Figure~\ref{fig:visual}C}}), inspired by watercolor effect \cite{pinna_1987, pinna_2001}.
 Watercolor effect describes the phenomenon in visual perception that surface color arises from thin boundaries and hence is applied here to perceive the background color in tight space.
-Overall, the figure-ground articulation creates two isolated layers in visual perception to display the two variables while maintaining the relative spatial relationship serving as a reference between the two.
+Overall, the figure-ground segmentation creates two isolated layers in visual perception to display the two variables while maintaining the relative spatial relationship serving as a reference between the two.
 In addition, other fundamental principles \cite{todorovic_2008}, such as proximity, similarity, continuity, and closure, incentivize the brain to group elements and dimensions in the visualization, guaranteeing an integrative perception of the complex multi-dimensional spatial map.
@@
 
-Next, we give two use cases to exemplify some utility of the proposed spatial visualization: (i) the spatially differential gene colocalization in the human dorsolateral prefrontal cortex (DLPFC) using spatial transcriptomics data \cite{huukimyers_2023}; (ii) multi-dimensional UMAP highlighting differential gene expression in data-driven cell clusters \cite{freytag_2020}.
+Next, we give two use cases to exemplify some utility of the proposed spatial visualization: (i) the spatially differential gene colocalization in the human dorsolateral prefrontal cortex using spatial transcriptomics data \cite{huukimyers_2023}; (ii) multi-dimensional UMAP highlighting differential gene expression in data-driven cell clusters \cite{freytag_2020}.
 
 \subsection{Multi-dimensional \textit{in situ} visualization}
-In a recent study investigating the molecular organization of human dorsolateral prefrontal cortex (DLPFC) \cite{huukimyers_2023}, two schizophrenia risk genes, membrane-bound ligand ephrin A5 (\textit{EFNA5}) and ephrin type-A receptor 5 (\textit{EPHA5}), were identified to colocalize via the cell-cell communication analysis.
+In a recent study investigating the molecular organization of human dorsolateral prefrontal cortex \cite{huukimyers_2023}, two schizophrenia risk genes, membrane-bound ligand ephrin A5 (\textit{EFNA5}) and ephrin type-A receptor 5 (\textit{EPHA5}), were identified to colocalize via the cell-cell communication analysis.
 In addition, data suggested Layer 6 was the most highly co-localized layer compared to other cortex layers.
-To visually examine the inference, we applied \texttt{escheR} to create a multi-dimensional \textit{in situ} spatial map (\textbf{Figure \ref{fig:visual}A-B}) that simultaneously exhibits the cortex layers (displayed with color-coded spot outlines) and the categorized colocalization status of genes \textit{FYN} and \textit{EFNA5} (displayed with color-coded spot fill).
-Compared to the traditional visualization where the cortex layers and the colocalization status are visualized in two side-by-side figures, our proposed visualization (\textbf{Figure \ref{fig:visual}D}) enables directly mapping colocalization status to the spatial domain, simplifying the perception of two sources of information and allowing cognitive comparison across cortex layers.
+To visually examine the inference, we applied \texttt{escheR} to create a multi-dimensional \textit{in situ} spatial map that simultaneously exhibits the cortex layers (displayed with color-coded spot outlines) and the categorized colocalization status of genes \textit{FYN} and \textit{EFNA5} (displayed with color-coded spot fill).
+Compared to the traditional visualization where the cortex layers and the colocalization status are visualized in two side-by-side figures (\textbf{Figure \ref{fig:visual}A-B}), our proposed visualization (\textbf{Figure \ref{fig:visual}D}) enables directly mapping colocalization status to the spatial domain, simplifying the perception of two sources of information and allowing cognitive comparison across cortex layers.
 
@@
 To address the overplotting problem, previous work proposed to apply hexagonal binning strategy to display the gene expression \cite{freytag_2020}.
-Here, the color-coded convex hulls are used to annotate different clusters of cells (\textbf{Figure~\ref{fig:visual}E}).
+Here, the color-coded convex hulls are used to annotate different clusters of cells (\textbf{Figure~\ref{fig:embedding}A}).
 However, the convex hulls create substantial overlapping areas, creating confusion when interpreting cluster memberships of hexagons in the overlapping areas.
-To improve the interpretability of the visualization, we replace the convex hulls with color-coded hexagons boundaries  (\textbf{Figure~\ref{fig:visual}F}) to avoid possible membership confusion.
+To improve the interpretability of the visualization, we replace the convex hulls with color-coded hexagons boundaries  (\textbf{Figure~\ref{fig:embedding}B}) to avoid possible membership confusion.
 We note that our contribution to improve the visualization is easily implemented without any modification of \texttt{schex} as both are built upon the Grammar of Graphics \cite{wilkinson_2012} standard.
@@
 
-\fixme{add any abbreviations here}
+\begin{itemize}[nosep]
+    \item \textbf{PCA}: principal component analysis
+    \item \textbf{$t$-SNE}: $t$-distributed Stochastic Neighbor Embedding
+    \item \textbf{UMAP}: Uniform Manifold Approximation and Projection
+    \item \textbf{EFNA5}: membrane-bound ligand ephrin A5
+    \item \textbf{EPHA5}: ephrin type-A receptor 5
+\end{itemize}
 
@@
 \begin{itemize}[nosep]
-    \item Conceptualization: BG
-    \item Data Curation: BG
-    \item \fixme{add more here} use credit system (https://www.elsevier.com/authors/policies-and-guidelines/credit-author-statement)
+    \item \textbf{Boyi Guo}:  Conceptualization, Methodology, Software, Validation, Formal analysis, Investigation, Data Curation, Writing, Visualization
+    \item \textbf{Stephanie C. Hicks}: Conceptualization, Resources, Writing - Review \& Editing, Visualization, Supervision, Project administration, Funding acquisition
 \end{itemize}
@@
 The UMAP example follows the `using\_schex' vignette in the \texttt{schex} package (\href{https://www.bioconductor.org/packages/schex}{bioconductor.org/packages/schex}).
-The code that generates these figures is deposited at \href{https://github.com/boyiguo1/Manuscript_escheR}{github.com/boyiguo1/Manuscript\_escheR} \fixme{version XXX}.
+The code that generates these figures is deposited at \href{https://github.com/boyiguo1/Manuscript_escheR}{github.com/boyiguo1/Manuscript\_escheR} (\textit{v1.0}).
 
@@
 \begin{center}
-\includegraphics[width=\textwidth]{Manuscript/figure/fig_grid.png}
-\caption{\small \textbf{\texttt{escheR} improves spatial visualizations following the Gestalt principles}.
-(\textbf{A-B}) The traditional visualization displays the cortex layers (spatial domains) (\textbf{A}) and the colocalization status (\textbf{B}) in a side-by-side figure, creating challenges to connecting colocalization status to spatial domains.
-(\textbf{C}) The watercolor effect enables displaying spatial domains by color-coding only outlines of circles.
-(\textbf{D}) \texttt{escheR} enables the multi-dimensional \texttt{in-situ} visualization that simultaneously displays the cortex layers and the colocalization status, substantially improving interpretability.
-(\textbf{E}) \texttt{schex} creates overlapped convex hulls, creating confusion when interpreting cluster memberships of hexagons in the overlapping areas.
-(\textbf{F}) \texttt{escheR} plots hexagon-specific membership to improve interpretability.}
+\includegraphics[width=\textwidth]{Manuscript/figure/embedding.jpg}
+\caption{\small \textbf{\texttt{escheR} improves multi-dimensional embedding visualizations}.
+UMAP representation along $x$ and $y$ axis of \fixme{what data?
+just generally need to describe and cite it, if appli}.
+(\textbf{A}) The \texttt{schex} R/Bioconductor package uses color-coded convex hulls to annotate cell memberships, creating confusion when interpreting hexagons in overlapping hulls.
+(\textbf{B}) Our package \texttt{escheR} plots hexagon-specific membership to improve interpretability.}
 \label{fig:embedding} 

### 2023-03-19 Manuscript/manuscript.tex
Finish Revision

@@
 \author[1]{Boyi Guo}
-\author[1,2]{Stephanie C. Hicks\thanks{$^*$Correspondence to \url{shicks19@jhu.edu}}}
+\author[1,2]{Stephanie C. Hicks\thanks{Correspondence to \url{shicks19@jhu.edu}}}
 \affil[1]{Department of Biostatistics, Johns Hopkins Bloomberg School of Public Health, MD, USA}
@@
 \section*{Abstract}
-The creation of effective visualizations is a fundamental component of data analysis. Nevertheless, current methods and
-tools are inadequate to visualize the novel spatially-resolved data, limiting the ability to derive new insights and discoveries.
-Specifically, there are no multi-dimensional \textit{in-situ} visualization that simultaneously displays multiple variables in a 2D spatial map.
-To address this problem, we introduce gestalt principles to the design of spatial maps, layering aesthetics to display multiple variables.
-The proposed visualization can be broadly applied to spatially resolved data and 2D embedding methods.
-We provide an open source R package \texttt{escheR}, which seamlessly integrates into the state-of-the-art spatial omics toolboxes.
+The creation of effective visualizations is a fundamental component of data analysis.
+In biomedical research, new challenges are emerging to visualize multi-dimensional data in a 2D space, but current data visualization tools have limited capabilities.
+To address this problem, we leverage Gestalt principles to improve the design and interpretability of multi-dimensional data in 2D data visualizations, layering aesthetics to display multiple variables.
+The proposed visualization can be applied to spatially-resolved transcriptomics data, but also broadly to data visualized in 2D space, such as embedding visualizations.
+We provide an open source R package \texttt{escheR}, which is built off of the state-of-the-art \texttt{ggplot2} visualization framework and can be seamlessly integrated into genomics toolboxes and workflows.
 
@@
 Both of these visualizations represent data in a 2D space and are motivated by recent advances in experimental technologies that profile molecules, including DNA, RNA, and proteins, at a single-cell or spatial resolution \cite{kashima_2020, moffitt_2022}.
-These technologies are commonly referred to as single-cell/single-nucleus RNA-sequencing (sc/snRNA-seq) \fixme{add citations} and \textit{in situ} spatially-resolved transcriptomics (SRT) \fixme{add citations}.
+Some most popular technologies include single-cell/nucleus RNA-sequencing (sc/snRNA-seq) \cite{aldridge_2020} and \textit{in situ} spatially-resolved transcriptomics \cite{larsson_2021}.
 
@@
 However, it is challenging to simultaneously visualize multi-dimensional data, such as information from disparate data domains (such as expression domain and spatial domain) or disparate data modalities (such as transcriptomics and proteomics) in the same plot.
-Currently, best practices for this include making two different plots displayed side-by-side (\fixme{\textbf{Figure \textbf{\ref{fig:visual}A-B}}}), one for gene expression and one for spatial domains.
+Currently, best practices for this include making two different plots displayed side-by-side (\textbf{Figure \textbf{\ref{fig:visual}A-B}}), one for gene expression and one for spatial domains.
 This creates cognitive gaps on how to associate the disparate information or how to interpret the biological findings of this multi-dimensional information regarding their (micro-)environment or colocalization.
@@
 \includegraphics[width=\textwidth]{Manuscript/figure/insitu.jpg}
-\caption{\small \textbf{\texttt{escheR} improves spatial visualizations following the Gestalt principles}.
-(\textbf{A-B}) The traditional visualization displays the colocalization plot (\textbf{A}) and the cortex layers (spatial domains) plot (\textbf{B}) side-by-side, creating challenges to connecting colocalization status to spatial domains.
+\caption{\small \textbf{\texttt{escheR} enables multi-dimensional spatial visualizations following the Gestalt principles of design}.
+(\textbf{A-B}) The traditional visualization displays the colocalization plot of the expression of two genes \textit{EFNA5} and \texit{EPHA5} (\textbf{A}) and the spatial domains from the dorsolateral prefrontal cortex (DLPFC) in postmorteum human brain \cite{huukimyers_2023} (\textbf{B}) side-by-side, creating challenges to cognitively connecting colocalization status to spatial domains.
 (\textbf{C}) The watercolor effect enables displaying spatial domains by color-coding only outlines of circles.
-(\textbf{D}) \texttt{escheR} enables the multi-dimensional \texttt{in-situ} visualization that simultaneously displays the cortex layers and the colocalization status, substantially improving interpretability.}
+(\textbf{D}) \texttt{escheR} enables the multi-dimensional \textit{in situ} visualization that simultaneously displays the cortex layers and the colocalization status, substantially improving interpretability.}
 \label{fig:visual} 
@@
 
-Gestalt principles\cite{todorovic_2008, palmer_1999} refer to a set of rules describing how humans perceive and interpret visual information and are commonly applied in art and designs.
+Gestalt principles \cite{todorovic_2008, palmer_1999} refer to a set of rules describing how humans perceive and interpret visual information and are commonly applied in art and designs.
 Developed in the 1920s by German psychologists Max Wertheimer, Kurt Koffka and Wolfgang Kohler, these principles help humans perceive a set of individual elements as a whole.
@@
 Here, we leverage the principles to be able to visualize multi-dimensional data in a unified 2D plot.
-Our approach is to use the state-of-art data visualization framework \texttt{ggplot2} \cite{ggplot2} following the grammar of graphics \cite{wilkinson_2012} and map individual variables to different aesthetics to simultaneously display disparate variables.
+Our approach is to use the state-of-art data visualization framework \texttt{ggplot2} \cite{ggplot2} following the Grammar of Graphics \cite{wilkinson_2012} and map individual variables to different aesthetics to simultaneously display disparate variables.
 Specifically, we apply the figure-ground segmentation \cite{peterson_2010} in displaying two variables: one variable (e.g.
@@
 
-For adjacent circles with limited space between them to display the background color, we use an economic implementation, colored outlines for these circles (\fixme{\textbf{Figure~\ref{fig:visual}C}}), inspired by watercolor effect \cite{pinna_1987, pinna_2001}.
+For adjacent circles with limited space between them to display the background color, we use an economic implementation, colored outlines for these circles (\textbf{Figure~\ref{fig:visual}C}), inspired by watercolor effect \cite{pinna_1987, pinna_2001}.
 Watercolor effect describes the phenomenon in visual perception that surface color arises from thin boundaries and hence is applied here to perceive the background color in tight space.
@@
 
-Next, we give two use cases to exemplify some utility of the proposed spatial visualization: (i) the spatially differential gene colocalization in the human dorsolateral prefrontal cortex using spatial transcriptomics data \cite{huukimyers_2023}; (ii) multi-dimensional UMAP highlighting differential gene expression in data-driven cell clusters \cite{freytag_2020}.
+Next, we give two use cases to exemplify some utility of the proposed spatial visualization: (i) the spatially differential gene colocalization in the human dorsolateral prefrontal cortex (DLPFC) using spatial transcriptomics data \cite{huukimyers_2023}; (ii) multi-dimensional UMAP highlighting differential gene expression in data-driven cell clusters \cite{freytag_2020}.
 
 \subsection{Multi-dimensional \textit{in situ} visualization}
-In a recent study investigating the molecular organization of human dorsolateral prefrontal cortex \cite{huukimyers_2023}, two schizophrenia risk genes, membrane-bound ligand ephrin A5 (\textit{EFNA5}) and ephrin type-A receptor 5 (\textit{EPHA5}), were identified to colocalize via the cell-cell communication analysis.
+In a recent study investigating the molecular organization of human dorsolateral prefrontal cortex \cite{huukimyers_2023}, two schizophrenia risk genes, membrane-bound ligand ephrin A5 (\textit{EFNA5}) and ephrin type-A receptor 5 (\textit{EPHA5}), were identified to colocalize via cell-cell communication analysis.
 In addition, data suggested Layer 6 was the most highly co-localized layer compared to other cortex layers.
-To visually examine the inference, we applied \texttt{escheR} to create a multi-dimensional \textit{in situ} spatial map that simultaneously exhibits the cortex layers (displayed with color-coded spot outlines) and the categorized colocalization status of genes \textit{FYN} and \textit{EFNA5} (displayed with color-coded spot fill).
+To visually examine the inference, we applied \texttt{escheR} to create a multi-dimensional \textit{in situ} spatial map that simultaneously exhibits the cortex layers (displayed with color-coded spot outlines) and the categorized colocalization status of genes \textit{EFNA5} and \textit{EPHA5} (displayed with color-coded spot fill).
 Compared to the traditional visualization where the cortex layers and the colocalization status are visualized in two side-by-side figures (\textbf{Figure \ref{fig:visual}A-B}), our proposed visualization (\textbf{Figure \ref{fig:visual}D}) enables directly mapping colocalization status to the spatial domain, simplifying the perception of two sources of information and allowing cognitive comparison across cortex layers.
@@
 To improve the interpretability of the visualization, we replace the convex hulls with color-coded hexagons boundaries  (\textbf{Figure~\ref{fig:embedding}B}) to avoid possible membership confusion.
-We note that our contribution to improve the visualization is easily implemented without any modification of \texttt{schex} as both are built upon the Grammar of Graphics \cite{wilkinson_2012} standard.
+We note that our contribution to improving the visualization is easily implemented without any modification of \texttt{schex} as both are built upon the Grammar of Graphics \cite{wilkinson_2012} standard.
 
 \section{Discussion}
-In summary, we propose an innovative multi-dimensional spatial visualization to simultaneously display multiple variables in a 2D coordinate system.
-Specifically, our design leverages the gestalt principle from visual perception to introduce other dimensions of a spatial map by iterative layering aesthetics.
-Developed upon \texttt{ggplot2}, we provide an open-source R package \texttt{escheR} that is seamlessly compatible with popular spatial resolved data analysis toolboxes.
-The proposed visualization has broad applications in visualizing both spot-based and image-based spatial resolved data, spatial multi-omics as well as other spatial-based data.
-The proposed visualization innovation highlights the importance and potential of translating theories in vision science to data science, addressing long-standing challenges in bioinformatics data visualization.
+Here, we propose an innovative multi-dimensional spatial visualization that simultaneously displays multiple variables in a 2D coordinate system.
+Specifically, our design leverages Gestalt principles from visual perception to create multiple visual dimensions in a spatial map by iterative layering aesthetics.
+Built upon \texttt{ggplot2}, we provide an open-source R package \texttt{escheR} that is seamlessly compatible with popular spatially-resolved transcriptomics and single-cell data analysis toolboxes.
 
-% While \texttt{escheR} is to design to display colocalization in their spatial context, the underlying principle can be broadly applied to address visualizaiton challenges in informaiton rich visualizations.
-The contrast color optical illusion provides allows to add another dimension of information in any spatial visualization, which can be further used to display multi-omics presence, greatly improve the interpretability.
+Adding a third dimension to 2D plots has been a long-standing challenge in visualization \cite{odonoghue_2010}.
+Our proposal addresses this fundamental challenge by introducing simple but effective design principles.
+Unlike computer-based interactive visualizations, the proposed visualization is free from any platform and technology restriction, creating an accessible and economic solution.
+In addition, the proposed visualization is easily scalable and hence can be applied to all types of spatially resolved data.
+Combining with a binning strategy, similar to \texttt{schex} \cite{freytag_2020}, to avoid possible overplotting, the proposed visualization can be also applied to visualized image-based spatially resolved data \cite{caicedo_2017}, in addition to aforementioned spot-based spatially resolved data (\textbf{Figure~\ref{fig:visual}}).
 
-% our programming provides flexibility by following the grammar of graphs, especially ggplot2
+Beyond the scope of biomedical research, the proposed visualization can be broadly translated to any visual analytic highlighting differentiation with respect to another measurement(s).
+To name a few, such visual analytics include examining differential tests,  explaining clustering, and visualizing subgroups.
+However, one of the most rewarding fields to apply the proposed visualization is the rapidly expanding field of biomedical multi-omics research \cite{hasin_2017}, where connecting different omics (data modalities) is the fundamental goal and hence greatly appreciating innovative multi-dimensional visualization.
+
+In summary, we propose a novel multi-dimensional visualization, implemented in an R package \texttt{escheR}, to address the simultaneous exhibition of multiple variables in 2D plots.
+The proposed visualization can be broadly applicable to the visual analytics of growingly complex biomedical data and beyond.
 
@@
     \item \textbf{PCA}: principal component analysis
-    \item \textbf{$t$-SNE}: $t$-distributed Stochastic Neighbor Embedding
-    \item \textbf{UMAP}: Uniform Manifold Approximation and Projection
+    \item \textbf{$t$-SNE}: $t$-distributed stochastic neighbor embedding
+    \item \textbf{UMAP}: uniform manifold approximation and projection
+    \item \textbf{scRNA-seq}: single-cell RNA-sequencing
+    \item \textbf{snRNA-seq}: single-nucleus RNA-sequencing
+    \item \textbf{SRT}: spatially-resolved transcriptomics 
     \item \textbf{EFNA5}: membrane-bound ligand ephrin A5
     \item \textbf{EPHA5}: ephrin type-A receptor 5
+    \item \textbf{DLPFC}: dorsolateral prefrontal cortex
 \end{itemize}
@@
 \includegraphics[width=\textwidth]{Manuscript/figure/embedding.jpg}
-\caption{\small \textbf{\texttt{escheR} improves multi-dimensional embedding visualizations}.
-UMAP representation along $x$ and $y$ axis of \fixme{what data?
-just generally need to describe and cite it, if appli}.
-(\textbf{A}) The \texttt{schex} R/Bioconductor package uses color-coded convex hulls to annotate cell memberships, creating confusion when interpreting hexagons in overlapping hulls.
-(\textbf{B}) Our package \texttt{escheR} plots hexagon-specific membership to improve interpretability.}
+\caption{\small \textbf{\texttt{escheR} enables multi-dimensional embedding visualizations}.
+The gene expression of \textit{POMGNT1} among peripheral blood mononuclear cells \cite{PBMC} under the UMAP representation.
+(\textbf{A}) The \texttt{schex} R/Bioconductor package uses color-coded convex hulls to annotate data-driven cell types, creating confusion when interpreting hexagons in overlapping hulls.
+(\textbf{B}) \texttt{escheR} plots hexagon-specific membership to improve interpretability.}
 \label{fig:embedding} 

### 2023-03-19 R/01-srt_coloc_plot.R
add R code to make plots in Figure 1

@@ -1,0 +1,167 @@
+# Packages ----------------------------------------------------------------
+library(SpatialExperiment)
+library(spatialLIBD)
+library(here)
+library(tidyverse)
+library(escheR)             # Proposed package
+
+
+
+# Helper Function ---------------------------------------------------------
+
+# Function to calculate the gene_name1, gene_name2 co-localization
+# of an SPE object
+# Return a SPE object with co-localization status as one variable in the colData
+calc_coloc <- function(spe, gene_name1, gene_name2, sample_id = NULL) {
+  if (!is.null(sample_id)) {
+    if (!sample_id %in% unique(spe$sample_id)) {
+      stop("Can't find sample_id")
+    }
+    spe <- spe[, spe$sample_id == sample_id]
+  }
+
+  if (!all(c(gene_name1, gene_name2) %in% rowData(spe)$gene_name)) {
+    stop("Can't find gene names")
+  }
+
+  gene1_idx <- which(rowData(spe)$gene_name == gene_name1)
+  gene2_idx <- which(rowData(spe)$gene_name == gene_name2)
+
+  cnt_df <- counts(spe)[c(gene1_idx, gene2_idx), ] |>
+    as.matrix() |>
+    t() |>
+    data.frame()
+
+  stopifnot(setequal(names(cnt_df), rowData(spe)[names(cnt_df), "gene_id"]))
+
+  old_names <- names(cnt_df)
+  names(cnt_df) <- paste0("gene", 1:2)
+  # Check if the correct rows are selected
+
+  coloc_cat <- cnt_df |>
+    pmap_chr(.f = function(gene1, gene2) {
+      if (gene1 != 0 && gene2 != 0) {
+        return(paste0(gene_name1, "&", gene_name2))
+      } else if (gene1 != 0) {
+        return(gene_name1)
+      } else if (gene2 != 0) {
+        return(gene_name2)
+      } else {
+        return("Neither")
+      }
+    })
+
+  colData(spe)$coloc <- coloc_cat |> factor()
+
+  return(spe)
+}
+
+
+# Code --------------------------------------------------------------------
+
+if (!exists("spe"))
+  spe <- fetch_data("spatialDLPFC_Visium")
+
+# Subset sample
+spe <- spe[, spe$sample_id == "Br8667_mid"]
+
+
[…]

### 2023-03-19 R/02_multi_d_umap.R
add multi-dimensional UMAP example

@@ -1,0 +1,96 @@
+# Packages ----------------------------------------------------------------
+library(igraph)
+library(schex)
+library(TENxPBMCData)
+library(scater)
+library(scran)
+library(ggrepel)
+
+
+
+# Code --------------------------------------------------------------------
+# This section of code was directly taken from schex vignettes.
+# https://github.com/SaskiaFreytag/schex/blob/master/vignettes/using_schex.Rmd
+
+# Data importing and preprocessing
+tenx_pbmc3k <- TENxPBMCData(dataset = "pbmc3k")
+
+rownames(tenx_pbmc3k) <- uniquifyFeatureNames(rowData(tenx_pbmc3k)$ENSEMBL_ID,
+                                              rowData(tenx_pbmc3k)$Symbol_TENx)
+rowData(tenx_pbmc3k)$Mito <- grepl("^MT-", rownames(tenx_pbmc3k))
+colData(tenx_pbmc3k) <- cbind(colData(tenx_pbmc3k),
+                              perCellQCMetrics(tenx_pbmc3k,
+                                               subsets=list(Mt=rowData(tenx_pbmc3k)$Mito)))
+rowData(tenx_pbmc3k) <- cbind(rowData(tenx_pbmc3k),
+                              perFeatureQCMetrics(tenx_pbmc3k))
+
+tenx_pbmc3k <- tenx_pbmc3k[, !colData(tenx_pbmc3k)$subsets_Mt_percent > 50]
+
+libsize_drop <- isOutlier(tenx_pbmc3k$total,
+                          nmads = 3,type = "lower", log = TRUE)
+feature_drop <- isOutlier(tenx_pbmc3k$detected,
+                          nmads = 3, type = "lower", log = TRUE)
+
+tenx_pbmc3k <- tenx_pbmc3k[, !(libsize_drop | feature_drop)]
+rm_ind <- calculateAverage(tenx_pbmc3k)<0
+tenx_pbmc3k <- tenx_pbmc3k[!rm_ind,]
+tenx_pbmc3k <- scater::logNormCounts(tenx_pbmc3k)
+tenx_pbmc3k <- runPCA(tenx_pbmc3k)
+
+# Changed the seed number from 10 to 100
+set.seed(100)
+
+tenx_pbmc3k <- runUMAP(tenx_pbmc3k, dimred = "PCA", spread = 1,
+                       min_dist = 0.4)
[…]

### 2023-05-09 Manuscript/manuscript.tex
Uncomment backmatters
Add co-authors' info

@@
 \usepackage[sort&compress,square,numbers]{natbib}
-\bibliographystyle{unsrtnat}
+\bibliographystyle{abbrv}
 
@@
 \author[1]{Boyi Guo}
-\author[1,2]{Stephanie C. Hicks\thanks{Correspondence to \url{shicks19@jhu.edu}}}
+\author[2]{Louise A. Huuki-Myers}
+\author[3,4]{Melissa Grant-Peters}
+\author[2]{Leonardo Collado-Torres}
+\author[1,5]{Stephanie C. Hicks\thanks{Correspondence to \url{shicks19@jhu.edu}}}
 \affil[1]{Department of Biostatistics, Johns Hopkins Bloomberg School of Public Health, MD, USA}
-\affil[2]{Malone Center for Engineering in Healthcare, Johns Hopkins University, MD, USA}
+\affil[2]{Lieber Institute for Brain Development, Johns Hopkins Medical Campus, Baltimore, MD, USA}
+\affil[3]{Genetics and Genomic Medicine, Great Ormond Street Institute of Child Health, University College London, London, UK}
+\affil[4]{Aligning Science Across Parkinson’s (ASAP) Collaborative Research Network, Chevy Chase, MD, USA}
+\affil[5]{Malone Center for Engineering in Healthcare, Johns Hopkins University, MD, USA}
 \date{\today}
@@
     \item \textbf{Boyi Guo}:  Conceptualization, Methodology, Software, Validation, Formal analysis, Investigation, Data Curation, Writing, Visualization
+    \item \textbf{Louise A. Huuki-Myers}: Conceptualization, Software
+    \item \textbf{Melissa Grant-Peters}: Conceptualization
+    \item \textbf{Leonardo Collado-Torres}: Conceptualization, Software
     \item \textbf{Stephanie C. Hicks}: Conceptualization, Resources, Writing - Review \& Editing, Visualization, Supervision, Project administration, Funding acquisition
@@
 \subsection*{Acknowledgements}
-BG would like to acknowledge Dr.
-Leonardo Collado-Torres, Louise Huuki-Myers, Dr.
-Lukas M.
-Weber, and Leon Di Stefano for their helpful comments, feedback and suggestions on \texttt{escheR} functionality.
+BG would like to acknowledge Kristen R.
+Maynard, Nicholas J.
+Eagles, Mina Ryten, Leon Di Stefano, and Lukas M.
+Weber for their helpful comments, feedback and suggestions on \texttt{escheR} functionality.
+Kristen R.
+Maynard and Nicholas J.
+Eagles are employed by the Lieber Institute for Brain Development (LIBD); Mina Ryten is employed by University College London; Leon Di Stefano and Lukas M.
+Weber are from the Johns Hopkins Bloomberg School of Public Health, Department of Biostatistics.
 
@@
     \item Boyi Guo (ORCiD: \href{https://orcid.org/0000-0003-2950-2349}{0000-0003-2950-2349})
+    \item Louise A. Huuki-Myers (ORCiD: \href{https://orcid.org/0000-0001-5148-3602}{0000-0001-5148-3602})
+    \item Melissa Grant-Peters (ORCiD: \href{https://orcid.org/0000-0003-0585-0971}{0000-0003-0585-0971})
+    \item Leonardo Collado-Torres (ORCiD: \href{https://orcid.org/0000-0003-2140-308X}{0000-0003-2140-308X})
     \item Stephanie C. Hicks (ORCiD: \href{https://orcid.org/0000-0002-7858-0231}{0000-0002-7858-0231})
@@
 
-\clearpage 
-% \printbibliography
+% \clearpage 
+\printbibliography
 \bibliography{refs}
@@
 
-{\large Boyi Guo, Stephanie C.\ Hicks$^*$}
+{\large Boyi Guo, Louise A. Huuki-Myers, Melissa Grant-Peters, Leonardo Collado-Torres, Stephanie C.\ Hicks$^*$}
 

### 2023-05-09 Manuscript/manuscript.tex
Adjust acknowledgement order

@@
 \subsection*{Acknowledgements}
-BG would like to acknowledge Kristen R.
-Maynard, Nicholas J.
-Eagles, Mina Ryten, Leon Di Stefano, and Lukas M.
-Weber for their helpful comments, feedback and suggestions on \texttt{escheR} functionality.
-Kristen R.
-Maynard and Nicholas J.
-Eagles are employed by the Lieber Institute for Brain Development (LIBD); Mina Ryten is employed by University College London; Leon Di Stefano and Lukas M.
+BG would like to acknowledge Nicholas J.
+Eagles, Kristen R.
+Maynard, Mina Ryten, Leon Di Stefano, and Lukas M.
+Weber (appearing in alphabetic order of last name) for their helpful comments, feedback and suggestions on \texttt{escheR} functionality.
+Nicholas J.
+Eagles and Kristen R.
+Maynard are employed by the Lieber Institute for Brain Development; Mina Ryten is employed by University College London; Leon Di Stefano and Lukas M.
 Weber are from the Johns Hopkins Bloomberg School of Public Health, Department of Biostatistics.

### 2023-05-09 Manuscript/manuscript.tex

@@
 \subsection*{Availability and implementation}
-The open source R package \texttt{escheR} is freely available on GitHub and is being submitted to Bioconductor (\href{https://github.com/boyiguo1/escheR}{https://github.com/boyiguo1/escheR}).
+The open source R package \texttt{escheR} is freely available on Bioconductor (\href{https://bioconductor.org/packages/escheR}{bioconductor.org/packages/escheR}).
 % and Zenodo (\fixme{add link}).
@@
 Escher) in the R programming language \cite{R}, leading to a simplified interface to navigate the implementation of the multi-dimensional visualization in 2D space.
-By adapting \texttt{ggplot2} standard, \texttt{escheR} can be seamlessly integrated into many popular spatial resolved toolboxes, such as \texttt{SpatialLIBD}\cite{pardo_2022}, \texttt{Seurat}\cite{hao_2021}, \texttt{Giotto}\cite{dries_2021} to name a few, and allow further theme customization with ease.
+By adapting \texttt{ggplot2} standard, \texttt{escheR} can be seamlessly integrated into many popular spatial resolved toolboxes, such as \texttt{SpatialLIBD }\cite{pardo_2022}, \texttt{Seurat} \cite{hao_2021}, \texttt{Giotto} \cite{dries_2021} to name a few, and allow further theme customization with ease.
 
@@
 The UMAP example follows the `using\_schex' vignette in the \texttt{schex} package (\href{https://www.bioconductor.org/packages/schex}{bioconductor.org/packages/schex}).
-The code that generates these figures is deposited at \href{https://github.com/boyiguo1/Manuscript_escheR}{github.com/boyiguo1/Manuscript\_escheR} (\textit{v1.0}).
+The code that generates these figures is deposited at \href{https://github.com/boyiguo1/Manuscript_escheR}{github.com/boyiguo1/Manuscript\_escheR} (Zenodo DOI: \href{https://zenodo.org/record/7915970}{10.5281/zenodo.7915970}).
+The open source software package \texttt{escheR} available in the R programming language is freely available on GitHub (\href{https://github.com/boyiguo1/escheR}{github.com/boyiguo1/escheR}) and Bioconductor (\href{https://bioconductor.org/packages/escheR}{bioconductor.org/packages/escheR}).
 
@@
 \subsection*{Acknowledgements}
-BG would like to acknowledge Nicholas J.
+We would like to acknowledge Nicholas J.
 Eagles, Kristen R.

### 2023-05-22 Manuscript/manuscript.tex
Add LCT's comment + funding information.

@@
 Our proposal addresses this fundamental challenge by introducing simple but effective design principles.
+These principles lead to visually easier-to-interpret graphics compared to geometry annotations (\textbf{Figure~\ref{fig:embedding}}).
 Unlike computer-based interactive visualizations, the proposed visualization is free from any platform and technology restriction, creating an accessible and economic solution.
@@
 Combining with a binning strategy, similar to \texttt{schex} \cite{freytag_2020}, to avoid possible overplotting, the proposed visualization can be also applied to visualized image-based spatially resolved data \cite{caicedo_2017}, in addition to aforementioned spot-based spatially resolved data (\textbf{Figure~\ref{fig:visual}}).
+\texttt{escheR} could also be applied with kernel gene-weighted density plots from \texttt{Nebulosa} \cite{alquicirahernandez_2021} and other ggplot2-based visualizations.
 
@@
 \subsection*{Funding}
-This project was supported by the National Institute of Mental Health R01MH126393 (BG, SCH), and CZF2019-002443 (SCH) from the Chan Zuckerberg Initiative DAF, an advised fund of Silicon Valley Community Foundation.
+This project was supported by the National Institute of Mental Health [R01MH126393 to BG and SCH, U01MH122849 to LAHM and LCT]; the Chan Zuckerberg Initiative DAF, an advised fund of Silicon Valley Community Foundation[CZF2019-002443 to SCH]; the Lieber Institute for Brain Development to LAHM and LCT; and Aligning Science Across Parkinson’s [ASAP-000478, ASAP-000509 to MGP] through the Michael J.
+Fox Foundation for Parkinson’s Research.
 All funding bodies had no role in the design of the study and collection, analysis, and interpretation of data and in writing the manuscript.

### 2023-10-05 Manuscript/manuscript.tex
Revision V2.0

--- 2023-05-09 Manuscript/manuscript.tex
@@
 
-\newcommand{\fixme}[1]{{\color{red} (#1)}}
+\newcommand{\rr}[1]{{\color{red} #1}}
 \newcommand{\coloc}{\texttt{escheR}}
@@
 
+\setlength{\bibsep}{0pt plus 0.3ex}
+
+\renewcommand\Affilfont{\fontsize{9}{10.8}\itshape}   % change auth affil
+
 %\setmainfont{Helvetica}
@@
 \author[1]{Boyi Guo}
-\author[2]{Louise A. Huuki-Myers}
-\author[3,4]{Melissa Grant-Peters}
-\author[2]{Leonardo Collado-Torres}
-\author[1,5]{Stephanie C. Hicks\thanks{Correspondence to \url{shicks19@jhu.edu}}}
+\author[2]{\rr{Louise A. Huuki-Myers}}
+\author[3,4]{\rr{Melissa Grant-Peters}}
+\author[2]{\rr{Leonardo Collado-Torres}}
+\author[1,5,6,7]{Stephanie C. Hicks\thanks{Correspondence to \url{shicks19@jhu.edu}}}
 \affil[1]{Department of Biostatistics, Johns Hopkins Bloomberg School of Public Health, MD, USA} 
@@
 \affil[4]{Aligning Science Across Parkinson’s (ASAP) Collaborative Research Network, Chevy Chase, MD, USA}
-\affil[5]{Malone Center for Engineering in Healthcare, Johns Hopkins University, MD, USA}
+\affil[5]{Department of Biomedical Engineering, Johns Hopkins School of Medicine, Baltimore, MD, USA}
+\affil[6]{Center for Computational Biology, Johns Hopkins University, Baltimore, MD, USA}
+\affil[7]{Malone Center for Engineering in Healthcare, Johns Hopkins University, MD, USA}
+
 \date{\today}
@@
 \subsection*{Availability and implementation}
-The open source R package \texttt{escheR} is freely available on Bioconductor (\href{https://bioconductor.org/packages/escheR}{bioconductor.org/packages/escheR}).
+The open source R package \texttt{escheR} is freely available on \rr{Bioconductor (\href{https://bioconductor.org/packages/escheR}{bioconductor.org/packages/escheR}).}
 % and Zenodo (\fixme{add link}).
@@
 Here, we focus on two types of visualizations in biomedical research, but note that the proposed ideas could be extended beyond these applications: (i) embedding visualizations, which project data into some low-dimensional embedding or mathematical space (e.g.
-Principal Components Analysis (PCA) \cite{hotelling_1933}, $t$-distributed Stochastic Neighbor Embedding ($t$-SNE) \cite{vandermaaten_2008}, or Uniform Manifold Approximation and Projection (UMAP) \cite{becht_2019}) and (ii) \textit{in situ} visualizations \cite{dries_2021, Lewis_2021, odonoghue_2021}, which aim to visualize molecules captured from \textit{in situ} imaging or sequencing technologies where \textit{in situ} refers to `in its original place'.
+Principal Components Analysis %(PCA)
+\cite{hotelling_1933}, $t$-distributed Stochastic Neighbor Embedding ($t$-SNE) \cite{vandermaaten_2008}, or Uniform Manifold Approximation and Projection (UMAP) \cite{becht_2019}) and (ii) \textit{in situ} visualizations \cite{dries_2021, Lewis_2021, odonoghue_2021}, which aim to visualize molecules captured from \textit{in situ} imaging or sequencing technologies where \textit{in situ} refers to `in its original place'.
 Both of these visualizations represent data in a 2D space and are motivated by recent advances in experimental technologies that profile molecules, including DNA, RNA, and proteins, at a single-cell or spatial resolution \cite{kashima_2020, moffitt_2022}.
-Some most popular technologies include single-cell/nucleus RNA-sequencing (sc/snRNA-seq) \cite{aldridge_2020} and \textit{in situ} spatially-resolved transcriptomics \cite{larsson_2021}.
+Some most popular technologies include single-cell/nucleus RNA-sequencing (sc/snRNA-seq)
+\cite{aldridge_2020} and \textit{in situ} spatially-resolved transcriptomics \cite{larsson_2021}. 
+% 
+% \vspace{-0.15in}
 
@@
 Currently, best practices for this include making two different plots displayed side-by-side (\textbf{Figure \textbf{\ref{fig:visual}A-B}}), one for gene expression and one for spatial domains.
-This creates cognitive gaps on how to associate the disparate information or how to interpret the biological findings of this multi-dimensional information regarding their (micro-)environment or colocalization.
+This creates cognitive gaps on how to associate the disparate information or how to interpret the biological findings of this multi-dimensional information regarding their microenvironment or colocalization.
 While interactive visualizations \cite{keller_2021, pardo_2022, sriworarat_2023} have the potential to mitigate this challenge, they are infeasible for scientific communications in static media, such as printed work.
@@
 \caption{\small \textbf{\texttt{escheR} enables multi-dimensional spatial visualizations following the Gestalt principles of design}.
-(\textbf{A-B}) The traditional visualization displays the colocalization plot of the expression of two genes \textit{EFNA5} and \texit{EPHA5} (\textbf{A}) and the spatial domains from the dorsolateral prefrontal cortex (DLPFC) in postmorteum human brain \cite{huukimyers_2023} (\textbf{B}) side-by-side, creating challenges to cognitively connecting colocalization status to spatial domains.
+(\textbf{A-B}) The traditional visualization displays the colocalization plot of the expression of two genes \textit{EFNA5} and \textit{EPHA5} (\textbf{A}) and the spatial domains from the dorsolateral prefrontal cortex in postmorteum human brain \cite{huukimyers_2023} (\textbf{B}) side-by-side, creating challenges to cognitively connecting colocalization status to spatial domains.
 (\textbf{C}) The watercolor effect enables displaying spatial domains by color-coding only outlines of circles.
@@
 
+% \vspace{-0.2in}
+
 \section{Results}
 
-To address these challenges, here we leverage the Gestalt (German for “unified whole”) principles for design \cite{todorovic_2008, palmer_1999} as a way to visualize multi-dimensional data in 2D visualizations.
+To address these challenges, here we leverage the Gestalt (German for “unified whole”) principles \cite{todorovic_2008, palmer_1999} as a path to visualize multi-dimensional data in 2D visualizations.
 We focus on the two types of data visualizations previously introduced that are widely used in biomedical research: (i) embedding visualizations and (ii) \textit{in situ} visualizations.
@@
 Escher) in the R programming language \cite{R}, leading to a simplified interface to navigate the implementation of the multi-dimensional visualization in 2D space.
-By adapting \texttt{ggplot2} standard, \texttt{escheR} can be seamlessly integrated into many popular spatial resolved toolboxes, such as \texttt{SpatialLIBD }\cite{pardo_2022}, \texttt{Seurat} \cite{hao_2021}, \texttt{Giotto} \cite{dries_2021} to name a few, and allow further theme customization with ease.
+By adapting \texttt{ggplot2} standard, \texttt{escheR} can be seamlessly integrated with \rr{popular data objects, including \texttt{SingleCellExperiment} \cite{amezquita2020} and \texttt{SpatialExperiment} \cite{righelli2022}, and easily work with popular pipelines, such as} \texttt{Seurat} \cite{hao_2021}, \texttt{Giotto} \cite{dries_2021} to name a few, and allow further theme customization with ease.
 
-Next, we give two use cases to exemplify some utility of the proposed spatial visualization: (i) the spatially differential gene colocalization in the human dorsolateral prefrontal cortex (DLPFC) using spatial transcriptomics data \cite{huukimyers_2023}; (ii) multi-dimensional UMAP highlighting differential gene expression in data-driven cell clusters \cite{freytag_2020}.
+Next, we give two use cases to exemplify some utility of the proposed spatial visualization: (i) the spatially differential gene colocalization in the human dorsolateral prefrontal cortex using spatial transcriptomics data \cite{huukimyers_2023}; (ii) multi-dimensional UMAP highlighting differential gene expression in data-driven cell clusters \cite{freytag_2020}.
 
@@
 Our proposal addresses this fundamental challenge by introducing simple but effective design principles.
-Unlike computer-based interactive visualizations, the proposed visualization is free from any platform and technology restriction, creating an accessible and economic solution.
+\rr{These principles lead to visually easier-to-interpret graphics compared to 2D-color gradients and geometry annotations (\textbf{Figure~\ref{fig:embedding}}).} Unlike computer-based interactive visualizations, the proposed visualization is free from any platform and technology restriction, creating an accessible and economical solution.
 In addition, the proposed visualization is easily scalable and hence can be applied to all types of spatially resolved data.
-Combining with a binning strategy, similar to \texttt{schex} \cite{freytag_2020}, to avoid possible overplotting, the proposed visualization can be also applied to visualized image-based spatially resolved data \cite{caicedo_2017}, in addition to aforementioned spot-based spatially resolved data (\textbf{Figure~\ref{fig:visual}}).
 
@@
 However, one of the most rewarding fields to apply the proposed visualization is the rapidly expanding field of biomedical multi-omics research \cite{hasin_2017}, where connecting different omics (data modalities) is the fundamental goal and hence greatly appreciating innovative multi-dimensional visualization.
+
+\rr{
+While the proposed design principle set-up a solid foundation, the proposed package \texttt{escheR} can be further optimized for image-based spatially resolved data \cite{caicedo_2017}.
+For example, to address possible overplotting of densely arranged single cells in tight space, strategies such as binning or hatching plots \cite{patrick2023spatial} can be incorporated.
+In addition, the flexibility to represent data points as polygons will be provided to allow visualization of cell morphology.
+% Combining with a binning strategy, similar to \texttt{schex} \cite{freytag_2020} to avoid possible overplotting \rr{or hatching plot proposed in \texttt{lisaClust} \cite{patrick2023spatial}}, the proposed visualization can be also applied to visualized image-based spatially resolved data , in addition to aforementioned spot-based spatially resolved data (\textbf{Figure~\ref{fig:visual}}).
+\rr{\texttt{escheR} could also be applied with kernel gene-weighted density plots from \texttt{Nebulosa} \cite{alquicirahernandez_2021} and other ggplot2-based visualizations.}
+% \rr{Independently, \texttt{lisaClust} \cite{patrick2023spatial} visualize cell types within spatial domains by generating hatching plots for immunofluorescence and mass cytometry imaging data.} 
+}
 
@@
     \item \textbf{Boyi Guo}:  Conceptualization, Methodology, Software, Validation, Formal analysis, Investigation, Data Curation, Writing, Visualization
+    \item \textbf{Erik D. Nelson}: Conceptualization, Software, Writing, Visualization
     \item \textbf{Louise A. Huuki-Myers}: Conceptualization, Software
@@
 \subsection*{Funding}
-This project was supported by the National Institute of Mental Health R01MH126393 (BG, SCH), and CZF2019-002443 (SCH) from the Chan Zuckerberg Initiative DAF, an advised fund of Silicon Valley Community Foundation.
+This project was supported by the National Institute of Mental Health [R01MH126393 to BG and SCH, U01MH122849 to LAHM and LCT]; the Chan Zuckerberg Initiative DAF, an advised fund of Silicon Valley Community Foundation [CZF2019-002443 to SCH]; the Lieber Institute for Brain Development to LAHM and LCT; and Aligning Science Across Parkinson’s [ASAP-000478, ASAP-000509 to MGP] through the Michael J.
+Fox Foundation for Parkinson’s Research.
 All funding bodies had no role in the design of the study and collection, analysis, and interpretation of data and in writing the manuscript.
 
-\subsection*{Acknowledgements}
+\subsection*{Acknowledgments}
 We would like to acknowledge Nicholas J.
@@
     \item Boyi Guo (ORCiD: \href{https://orcid.org/0000-0003-2950-2349}{0000-0003-2950-2349})
+    \item Erik D. Nelson (ORCiD: \href{https://orcid.org/0000-0001-8477-0982}{0000-0001-8477-0982})
     \item Louise A. Huuki-Myers (ORCiD: \href{https://orcid.org/0000-0001-5148-3602}{0000-0001-5148-3602})
@@
 
-% \clearpage 
-\printbibliography
+\clearpage 
+% \printbibliography
 \bibliography{refs}
@@
 \includegraphics[width=\textwidth]{Manuscript/figure/embedding.jpg}
-\caption{\small \textbf{\texttt{escheR} enables multi-dimensional embedding visualizations}.
+\caption{\small \textbf{\texttt{escheR} \rr{improves} multi-dimensional embedding visualizations}.
 The gene expression of \textit{POMGNT1} among peripheral blood mononuclear cells \cite{PBMC} under the UMAP representation.
 (\textbf{A}) The \texttt{schex} R/Bioconductor package uses color-coded convex hulls to annotate data-driven cell types, creating confusion when interpreting hexagons in overlapping hulls.
-(\textbf{B}) \texttt{escheR} plots hexagon-specific membership to improve interpretability.}
+(\textbf{B}) \texttt{escheR} plots hexagon-specific membership to avoid substantial overlapping of cluster membership.
+\rr{Due to the binning strategy, cautions in interpretation are needed due to possible cluster intermixing within each hexagon.} }
 \label{fig:embedding} 

### 2023-11-08 Manuscript/manuscript.tex
Sync overleaf and Github

@@
 \author[1]{Boyi Guo}
-\author[2]{\rr{Louise A. Huuki-Myers}}
-\author[3,4]{\rr{Melissa Grant-Peters}}
-\author[2]{\rr{Leonardo Collado-Torres}}
+\author[2]{Louise A. Huuki-Myers}
+\author[3,4]{Melissa Grant-Peters}
+\author[2]{Leonardo Collado-Torres}
 \author[1,5,6,7]{Stephanie C. Hicks\thanks{Correspondence to \url{shicks19@jhu.edu}}}
@@
 \subsection*{Availability and implementation}
-The open source R package \texttt{escheR} is freely available on \rr{Bioconductor (\href{https://bioconductor.org/packages/escheR}{bioconductor.org/packages/escheR}).}
+The open source R package \texttt{escheR} is freely available on Bioconductor (\href{https://bioconductor.org/packages/escheR}{bioconductor.org/packages/escheR}).
 % and Zenodo (\fixme{add link}).
@@
 \end{center}
+\vspace{-0.3in}
 \end{figure*}
@@
 
+\begin{figure*}[t]
+\begin{center}
+\includegraphics[width=\textwidth]{Manuscript/figure/embedding.jpg}
+\caption{\small \textbf{\texttt{escheR} improves multi-dimensional embedding visualizations}.
+The gene expression of \textit{POMGNT1} among peripheral blood mononuclear cells \cite{PBMC} under the UMAP representation.
+(\textbf{A}) The \texttt{schex} R/Bioconductor package uses color-coded convex hulls to annotate data-driven cell types, creating confusion when interpreting hexagons in overlapping hulls.
+(\textbf{B}) \texttt{escheR} plots hexagon-specific membership to avoid substantial overlapping of cluster membership.
+Due to the binning strategy, cautions in interpretation are needed due to possible cluster intermixing within each hexagon.}
+\label{fig:embedding} 
+\end{center}
+\vspace{-0.3in}
+\end{figure*}
+
 Here, we provide an open-source package called \texttt{escheR} (named after the graphic artist M.C.
 Escher) in the R programming language \cite{R}, leading to a simplified interface to navigate the implementation of the multi-dimensional visualization in 2D space.
-By adapting \texttt{ggplot2} standard, \texttt{escheR} can be seamlessly integrated with \rr{popular data objects, including \texttt{SingleCellExperiment} \cite{amezquita2020} and \texttt{SpatialExperiment} \cite{righelli2022}, and easily work with popular pipelines, such as} \texttt{Seurat} \cite{hao_2021}, \texttt{Giotto} \cite{dries_2021} to name a few, and allow further theme customization with ease.
+By adapting \texttt{ggplot2} standard, \texttt{escheR} can be seamlessly integrated with popular data objects, including \texttt{SingleCellExperiment} \cite{amezquita2020} and \texttt{SpatialExperiment} \cite{righelli2022}, and easily work with popular pipelines, such as \texttt{Seurat} \cite{hao_2021}, \texttt{Giotto} \cite{dries_2021} to name a few, and allow further theme customization with ease.
 
@@
 Our proposal addresses this fundamental challenge by introducing simple but effective design principles.
-\rr{These principles lead to visually easier-to-interpret graphics compared to 2D-color gradients and geometry annotations (\textbf{Figure~\ref{fig:embedding}}).} Unlike computer-based interactive visualizations, the proposed visualization is free from any platform and technology restriction, creating an accessible and economical solution.
+These principles lead to visually easier-to-interpret graphics compared to 2D-color gradients and geometry annotations (\textbf{Figure~\ref{fig:embedding}}).
+Unlike computer-based interactive visualizations, the proposed visualization is free from any platform and technology restriction, creating an accessible and economical solution.
 In addition, the proposed visualization is easily scalable and hence can be applied to all types of spatially resolved data.
@@
 
-\rr{
 While the proposed design principle set-up a solid foundation, the proposed package \texttt{escheR} can be further optimized for image-based spatially resolved data \cite{caicedo_2017}.
@@
 % \rr{Independently, \texttt{lisaClust} \cite{patrick2023spatial} visualize cell types within spatial domains by generating hatching plots for immunofluorescence and mass cytometry imaging data.}
-}
 
@@
 
+%% BIBLIOGRAPHY ==========================================================
+
+% \printbibliography
+\bibliography{refs}
+
+\clearpage
 %% EPILOGUE ================================================================
@@
     \item \textbf{Boyi Guo}:  Conceptualization, Methodology, Software, Validation, Formal analysis, Investigation, Data Curation, Writing, Visualization
-    \item \textbf{Erik D. Nelson}: Conceptualization, Software, Writing, Visualization
     \item \textbf{Louise A. Huuki-Myers}: Conceptualization, Software
@@
 None declared.
+% \onecolumn
 
-%% BIBLIOGRAPHY ==========================================================
+% {\huge Supplementary Materials}
 
-\clearpage 
-% \printbibliography
-\bibliography{refs}
+% \hrule
 
-\clearpage
-\onecolumn
+% \vspace*{0.5cm}
 
-{\huge Supplementary Materials}
+% \begin{center}
 
-\hrule
+% {\Large \texttt{escheR}: Unified multi-dimensional visualizations with Gestalt principles}
 
-\vspace*{0.5cm}
+% \vspace*{0.75cm}
 
-\begin{center}
+% {\large Boyi Guo, Louise A. Huuki-Myers, Melissa Grant-Peters, Leonardo Collado-Torres, Stephanie C.\ Hicks$^*$}
 
-{\Large \texttt{escheR}: Unified multi-dimensional visualizations with Gestalt principles}
+% \vspace*{0.3cm}
 
-\vspace*{0.75cm}
+% {\small $^*$Correspondence to \url{shicks19@jhu.edu}}
 
-{\large Boyi Guo, Louise A. Huuki-Myers, Melissa Grant-Peters, Leonardo Collado-Torres, Stephanie C.\ Hicks$^*$}
+% \end{center}
 
-\vspace*{0.3cm}
+% \renewcommand{\figurename}{Supplementary Figure}
+% \renewcommand{\tablename}{Supplementary Table}
+% \setcounter{figure}{0}
+% \setcounter{table}{0}
+% \setcounter{section}{0}
+% \setcounter{page}{1}
+% \makeatletter
+% \renewcommand{\thefigure}{S\@arabic\c@figure}
+% \renewcommand{\thetable}{S\@arabic\c@table}
+% \renewcommand{\thesection}{Supplemental Note S\@arabic\c@section}
+% \makeatother
 
-{\small $^*$Correspondence to \url{shicks19@jhu.edu}}
+% \vspace*{1cm}
 
-\end{center}
+% {\bf \large Contents}
 
-\renewcommand{\figurename}{Supplementary Figure}
-\renewcommand{\tablename}{Supplementary Table}
-\setcounter{figure}{0}
-\setcounter{table}{0}
-\setcounter{section}{0}
-\setcounter{page}{1}
-\makeatletter
-\renewcommand{\thefigure}{S\@arabic\c@figure}
-\renewcommand{\thetable}{S\@arabic\c@table}
-\renewcommand{\thesection}{Supplemental Note S\@arabic\c@section}
-\makeatother
+% % \begin{enumerate}
+%     \item \textbf{Supplemental Figures~S1.}
+% \end{enumerate}
 
-\vspace*{1cm}
+% \clearpage
 
-{\bf \large Contents}
-
-\begin{enumerate}
-    \item \textbf{Supplemental Figures~S1.}
-\end{enumerate}
-
-\clearpage
-
-\noindent {\LARGE Supplemental Figures}
-
-\begin{figure*}[!h]
-\begin{center}
-\includegraphics[width=\textwidth]{Manuscript/figure/embedding.jpg}
-\caption{\small \textbf{\texttt{escheR} \rr{improves} multi-dimensional embedding visualizations}.
-The gene expression of \textit{POMGNT1} among peripheral blood mononuclear cells \cite{PBMC} under the UMAP representation.
-(\textbf{A}) The \texttt{schex} R/Bioconductor package uses color-coded convex hulls to annotate data-driven cell types, creating confusion when interpreting hexagons in overlapping hulls.
-(\textbf{B}) \texttt{escheR} plots hexagon-specific membership to avoid substantial overlapping of cluster membership.
-\rr{Due to the binning strategy, cautions in interpretation are needed due to possible cluster intermixing within each hexagon.} }
-\label{fig:embedding} 
-\end{center}
-\end{figure*}
+% \noindent {\LARGE Supplemental Figures}
 

### 2023-11-13 R/02_multi_d_umap.R
Update Code to create Figure 2 Embbeding Vis

@@ -4,19 +4,27 @@ library(schex)
 library(TENxPBMCData)
 library(scater)
 library(scran)
+library(scuttle)
 library(ggrepel)
+library(BiocSingular)
+library(escheR)
+stopifnot(package.version("escheR") >= "1.3.1")
+
+if(Biobase::package.version("Matrix") > "1.6-1")
+  # Bug report on Matrix 1.6-2: https://github.com/bwlewis/irlba/issues/70
+  stop("Matrix 1.6-2 may not function properly to calculate PCA.")
 
 
 
-# Code --------------------------------------------------------------------
+# Data Reprocessing Code --------------------------------------------------------------------
 # This section of code was directly taken from schex vignettes.
 # https://github.com/SaskiaFreytag/schex/blob/master/vignettes/using_schex.Rmd
-
+set.seed(3)
 # Data importing and preprocessing
 tenx_pbmc3k <- TENxPBMCData(dataset = "pbmc3k")
 
-rownames(tenx_pbmc3k) <- uniquifyFeatureNames(rowData(tenx_pbmc3k)$ENSEMBL_ID,
-                                              rowData(tenx_pbmc3k)$Symbol_TENx)
+rownames(tenx_pbmc3k) <- scuttle::uniquifyFeatureNames(rowData(tenx_pbmc3k)$ENSEMBL_ID,
+                                                       rowData(tenx_pbmc3k)$Symbol_TENx)
 rowData(tenx_pbmc3k)$Mito <- grepl("^MT-", rownames(tenx_pbmc3k))
 colData(tenx_pbmc3k) <- cbind(colData(tenx_pbmc3k),
                               perCellQCMetrics(tenx_pbmc3k,
@@ -35,62 +43,106 @@ tenx_pbmc3k <- tenx_pbmc3k[, !(libsize_drop | feature_drop)]
 rm_ind <- calculateAverage(tenx_pbmc3k)<0
 tenx_pbmc3k <- tenx_pbmc3k[!rm_ind,]
 tenx_pbmc3k <- scater::logNormCounts(tenx_pbmc3k)
-tenx_pbmc3k <- runPCA(tenx_pbmc3k)
+tenx_pbmc3k <- scater::runPCA(tenx_pbmc3k)
 
 # Changed the seed number from 10 to 100
-set.seed(100)
 
+#
 tenx_pbmc3k <- runUMAP(tenx_pbmc3k, dimred = "PCA", spread = 1,
                        min_dist = 0.4)
[…]

### 2023-11-13 R/02_multi_d_umap.R
Print figure to manuscript folder

@@ -142,7 +142,7 @@ p_fig2 <- ggpubr::ggarrange(
 )
 
 
-ggsave("figs/escher_UMAP.tiff", p_fig2,
+ggsave("Manuscript/figure/embedding.tiff", p_fig2,
        height = 8,
        width = 12)
 

### 2023-11-13 Manuscript/manuscript_BioAdv_format.tex

%%
%% Copyright 2022 OXFORD UNIVERSITY PRESS
%%
%% This file is part of the 'oup-authoring-template Bundle'.
%% ---------------------------------------------
%%
%% It may be distributed under the conditions of the LaTeX Project Public
%% License, either version 1.2 of this license or (at your option) any
%% later version.  The latest version of this license is in
%%    http://www.latex-project.org/lppl.txt
%% and version 1.2 or later is part of all distributions of LaTeX
%% version 1999/12/01 or later.
%%
%% The list of all files belonging to the 'oup-authoring-template Bundle' is
%% given in the file `manifest.txt'.
%%
%% Template article for OXFORD UNIVERSITY PRESS's document class `oup-authoring-template'
%% with bibliographic references
%%

%%%CONTEMPORARY%%%
\documentclass[unnumsec,webpdf,contemporary,large]{oup-authoring-template}%
%\documentclass[unnumsec,webpdf,contemporary,large,namedate]{oup-authoring-template}% uncomment this line for author year citations and comment the above
%\documentclass[unnumsec,webpdf,contemporary,medium]{oup-authoring-template}
%\documentclass[unnumsec,webpdf,contemporary,small]{oup-authoring-template}

%%%MODERN%%%
%\documentclass[unnumsec,webpdf,modern,large]{oup-authoring-template}
%\documentclass[unnumsec,webpdf,modern,large,namedate]{oup-authoring-template}% uncomment this line for author year citations and comment the above
%\documentclass[unnumsec,webpdf,modern,medium]{oup-authoring-template}
%\documentclass[unnumsec,webpdf,modern,small]{oup-authoring-template}

%%%TRADITIONAL%%%
%\documentclass[unnumsec,webpdf,traditional,large]{oup-authoring-template}
%\documentclass[unnumsec,webpdf,traditional,large,namedate]{oup-authoring-template}% uncomment this line for author year citations and comment the above
%\documentclass[unnumsec,namedate,webpdf,traditional,medium]{oup-authoring-template}
%\documentclass[namedate,webpdf,traditional,small]{oup-authoring-template}

%\onecolumn % for one column layouts

%\usepackage{showframe}

\graphicspath{{figure/}}

% line numbers
\usepackage[mathlines, switch]{lineno}
\usepackage[right]{lineno}

\theoremstyle{thmstyleone}%
\newtheorem{theorem}{Theorem}%  meant for continuous numbers
%%\newtheorem{theorem}{Theorem}[section]% meant for sectionwise numbers
%% optional argument [theorem] produces theorem numbering sequence instead of independent numbers for Proposition
\newtheorem{proposition}[theorem]{Proposition}%
%%\newtheorem{proposition}{Proposition}% to get separate numbers for theorem and proposition etc.
\theoremstyle{thmstyletwo}%
\newtheorem{example}{Example}%
\newtheorem{remark}{Remark}%
\theoremstyle{thmstylethree}%
\newtheorem{definition}{Definition}

\begin{document}

\journaltitle{Journal Title Here}
\DOI{DOI HERE}
\copyrightyear{2022}
\pubyear{2019}
\access{Advance Access Publication Date: Day Month Year}
\appnotes{Applicaiton Notes}

\firstpage{1}

%\subtitle{Subject Section}

\title[Short Article Title]{Article Title}

\author[1,$\ast$]{First Author}
\author[2]{Second Author}
\author[3]{Third Author}
\author[3]{Fourth Author}
\author[4]{Fifth Author\ORCID{0000-0000-0000-0000}}

\authormark{Author Name et al.}

\address[1]{\orgdiv{Department}, \orgname{Organization}, \orgaddress{\street{Street}, \postcode{Postcode}, \state{State}, \country{Country}}}
\address[2]{\orgdiv{Department}, \orgname{Organization}, \orgaddress{\street{Street}, \postcode{Postcode}, \state{State}, \country{Country}}}
\address[3]{\orgdiv{Department}, \orgname{Organization}, \orgaddress{\street{Street}, \postcode{Postcode}, \state{State}, \country{Country}}}
\address[4]{\orgdiv{Department}, \orgname{Organization}, \orgaddress{\street{Street}, \postcode{Postcode}, \state{State}, \country{Country}}}

\corresp[$\ast$]{Corresponding author. \href{email:email-id.com}{email-id.com}}

\received{Date}{0}{Year}
\revised{Date}{0}{Year}
\accepted{Date}{0}{Year}

%\editor{Associate Editor: Name}

%\abstract{
%\textbf{Motivation:} .\\
%\textbf{Results:} .\\
%\textbf{Availability:} .\\
%\textbf{Contact:} \href{name@email.com}{name@email.com}\\
%\textbf{Supplementary information:} Supplementary data are available at \textit{Journal Name}
%online.}

\abstract{Abstracts must be able to stand alone and so cannot contain citations to
the paper's references, equations, etc. An abstract must consist of a single
paragraph and be concise. Because of online formatting, abstracts must appear
as plain as possible.}
\keywords{keyword1, Keyword2, Keyword3, Keyword4}

% \boxedtext{
% \begin{itemize}
% \item Key boxed text here.
% \item Key boxed text here.
% \item Key boxed text here.
% \end{itemize}}

\maketitle

\section{Introduction}
The introduction introduces the context and summarizes the manuscript. It is importantly to clearly state the contributions
of this piece of work. Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.

This is an example of a new parapgraph with a numbered footnote\footnote{\url{https://data.gov.uk/}} and a second footnote marker.\footnote{Example of footnote text.}

\section{This is an example for first level head - section head}\label{sec2}

Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum (refer Section~\ref{sec5}).

\subsection{This is an example for second level head - subsection head}\label{subsec1}

Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.

\subsubsection{This is an example for third level head - subsubsection head}\label{subsubsec1}

Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do
eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. %Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.

\paragraph{This is an example for fourth level head - paragraph head}

Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.

\section{This is an example for first level head}\label{sec3}

\subsection{This is an example for second level head - subsection head}\label{subsec2}

\subsubsection{This is an example for third level head - subsubsection head}\label{subsubsec2}

Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.

\paragraph{This is an example for fourth level head - paragraph head}

Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.

\section{Equations}\label{sec4}

Equations in \LaTeX{} can either be inline or set as display equations. For
inline equations use the \verb+$...$+ commands. Eg: the equation
$H\psi = E \psi$ is written via the command \verb+$H \psi = E \psi$+.

For display equations (with auto generated equation numbers)
one can use the equation or eqnarray environments:
\begin{equation}
\|\tilde{X}(k)\|^2 \leq\frac{\sum\limits_{i=1}^{p}\left\|\tilde{Y}_i(k)\right\|^2+\sum\limits_{j=1}^{q}\left\|\tilde{Z}_j(k)\right\|^2 }{p+q},\label{eq1}
\end{equation}
where,
\begin{align}
D_\mu &=  \partial_\mu - ig \frac{\lambda^a}{2} A^a_\mu \nonumber \\
F^a_{\mu\nu} &= \partial_\mu A^a_\nu - \partial_\nu A^a_\mu + g f^{abc} A^b_\mu A^a_\nu.\label{eq2}
\end{align}
Notice the use of \verb+\nonumber+ in the align environment at the end
of each line, except the last, so as not to produce equation numbers on
lines where no equation numbers are required. The \verb+\label{}+ command
should only be used at the last line of an align environment where
\verb+\nonumber+ is not used.
\begin{equation}
Y_\infty = \left( \frac{m}{\textrm{GeV}} \right)^{-3}
    \left[ 1 + \frac{3 \ln(m/\textrm{GeV})}{15}
    + \frac{\ln(c_2/5)}{15} \right].
\end{equation}
The class file also supports the use of \verb+\mathbb{}+, \verb+\mathscr{}+ and
\verb+\mathcal{}+ commands. As such \verb+\mathbb{R}+, \verb+\mathscr{R}+
and \verb+\mathcal{R}+ produces $\mathbb{R}$, $\mathscr{R}$ and $\mathcal{R}$
respectively (refer Subsubsection~\ref{subsubsec3}).

Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do
eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum. Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do
eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum. 
Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do
eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.

\section{Tables}\label{sec5}

Tables can be inserted via the normal table and tabular environment. To put
footnotes inside tables one has to Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum. use the additional ``tablenotes" environment
enclosing the tabular environment. The footnote appears just below the table
itself (refer Tables~\ref{tab1} and \ref{tab2}).

\begin{verbatim}
\begin{table}[t]
\begin{center}
\begin{minipage}{<width>}
\caption{<table-caption>\label{<table-label>}}%
\begin{tabular}{@{}llll@{}}
\toprule
column 1 & column 2 & column 3 & column 4\\
\midrule
row 1 & data 1 & data 2          & data 3 \\
row 2 & data 4 & data 5$^{1}$ & data 6 \\
row 3 & data 7 & data 8      & data 9$^{2}$\\
\botrule
\end{tabular}
\begin{tablenotes}%
\item Source: Example for source.
\item[$^{1}$] Example for a 1st table footnote.
\item[$^{2}$] Example for a 2nd table footnote.
\end{tablenotes}
\end{minipage}
\end{center}
\end{table}
\end{verbatim}

Lengthy tables which do not fit within textwidth should be set as rotated tables. For this, we need to use \verb+\begin{sidewaystable}...+ \verb+\end{sidewaystable}+ instead of\break \verb+\begin{table}...+ \verb+\end{table}+ environment.

\begin{table}[!t]
\caption{Caption text\label{tab1}}%
\begin{tabular*}{\columnwidth}{@{\extracolsep\fill}llll@{\extracolsep\fill}}
\toprule
column 1 & column 2  & column 3 & column 4\\
\midrule
row 1    & data 1   & data 2  & data 3  \\
row 2    & data 4   & data 5$^{1}$  & data 6  \\
row 3    & data 7   & data 8  & data 9$^{2}$  \\
\botrule
\end{tabular*}
\begin{tablenotes}%
\item Source: This is an example of table footnote this is an example of table footnote this is an example of table footnote this is an example of~table footnote this is an example of table footnote
\item[$^{1}$] Example for a first table footnote.
\item[$^{2}$] Example for a second table footnote.
\end{tablenotes}
\end{table}

\begin{table*}[t]
\caption{Example of a lengthy table which is set to full textwidth.\label{tab2}}
\tabcolsep=0pt%%
\begin{tabular*}{\textwidth}{@{\extracolsep{\fill}}lcccccc@{\extracolsep{\fill}}}
\toprule%
& \multicolumn{3}{@{}c@{}}{Element 1$^{1}$} & \multicolumn{3}{@{}c@{}}{Element 2$^{2}$} \\
\cline{2-4}\cline{5-7}%
Project & Energy & $\sigma_{calc}$ & $\sigma_{expt}$ & Energy & $\sigma_{calc}$ & $\sigma_{expt}$ \\
\midrule
Element 3  & 990 A & 1168 & $1547\pm12$ & 780 A & 1166 & $1239\pm100$\\
Element 4  & 500 A & 961  & $922\pm10$  & 900 A & 1268 & $1092\pm40$\\
\botrule
\end{tabular*}
\begin{tablenotes}%
\item Note: This is an example of table footnote this is an example of table footnote this is an example of table footnote this is an example of~table footnote this is an example of table footnote
\item[$^{1}$] Example for a first table footnote.
\item[$^{2}$] Example for a second table footnote.\vspace*{6pt}
\end{tablenotes}
\end{table*}

\begin{sidewaystable}%[!p]
\caption{Tables which are too long to fit, should be written using the ``sidewaystable" environment as shown here\label{tab3}}
\tabcolsep=0pt%
\begin{tabular*}{\textwidth}{@{\extracolsep{\fill}}lcccccc@{\extracolsep{\fill}}}
\toprule%
& \multicolumn{3}{@{}c@{}}{Element 1$^{1}$}& \multicolumn{3}{@{}c@{}}{Element$^{2}$} \\
\cline{2-4}\cline{5-7}%
Projectile & Energy     & $\sigma_{calc}$ & $\sigma_{expt}$ & Energy & $\sigma_{calc}$ & $\sigma_{expt}$ \\
\midrule
Element 3 & 990 A & 1168 & $1547\pm12$ & 780 A & 1166 & $1239\pm100$ \\
Element 4 & 500 A & 961  & $922\pm10$  & 900 A & 1268 & $1092\pm40$ \\
\botrule
\end{tabular*}
\begin{tablenotes}%
\item Note: This is an example of a table footnote this is an example of a table footnote this is an example of a table footnote this is an example of a table footnote this is an example of a table footnote
\item[$^{1}$] This is an example of a table footnote
\end{tablenotes}
\end{sidewaystable}

\section{Figures}\label{sec6}

As per display \LaTeX\ standards one has to use eps images for \verb+latex+ compilation and \verb+pdf/jpg/png+ images for
\verb+pdflatex+ compilation. This is one of the major differences between \verb+latex+
and \verb+pdflatex+. The images should be single-page documents. The command for inserting images
for \verb+latex+ and \verb+pdflatex+ can be generalized. The package used to insert images in \verb+latex/pdflatex+ is the
graphicx package. Figures can be inserted via the normal figure environment as shown in the below example:

\begin{figure}[!t]%
\centering
{\color{black!20}\rule{213pt}{37pt}}
\caption{This is a widefig. This is an example of a long caption this is an example of a long caption  this is an example of a long caption this is an example of a long caption}\label{fig1}
\end{figure}

\begin{figure*}[!t]%
\centering
{\color{black!20}\rule{438pt}{74pt}}
\caption{This is a widefig. This is an example of a long caption this is an example of a long caption  this is an example of a long caption this is an example of a long caption}\label{fig2}
\end{figure*}

\begin{verbatim}
\begin{figure}[t]
        \centering\includegraphics{<eps-file>}
        \caption{<figure-caption>}
        \label{<figure-label>}
\end{figure}
\end{verbatim}

Test text here.

For sample purposes, we have included the width of images in the
optional argument of \verb+\includegraphics+ tag. Please ignore this.
Lengthy figures which do not fit within textwidth should be set in rotated mode. For rotated figures, we need to use \verb+\begin{sidewaysfigure}+ \verb+...+ \verb+\end{sidewaysfigure}+ instead of the \verb+\begin{figure}+ \verb+...+ \verb+\end{figure}+ environment.

\begin{sidewaysfigure}%
\centering
{\color{black!20}\rule{610pt}{102pt}}
\caption{This is an example for a sideways figure. This is an example of a long caption this is an example of a long caption  this is an example of a long caption this is an example of a long caption}\label{fig3}
\end{sidewaysfigure}

\section{Algorithms, Program codes and Listings}\label{sec7}

Packages \verb+algorithm+, \verb+algorithmicx+ and \verb+algpseudocode+ are used for setting algorithms in latex.
For this, one has to use the below format:

\begin{verbatim}
\begin{algorithm}
\caption{<alg-caption>}\label{<alg-label>}
\begin{algorithmic}[1]
. . .
\end{algorithmic}
\end{algorithm}
\end{verbatim}

You may need to refer to the above-listed package documentations for more details before setting an \verb+algorithm+ environment.
To set program codes, one has to use the \verb+program+ package. We need to use the \verb+\begin{program}+ \verb+...+
\verb+\end{program}+ environment to set program codes.

\begin{algorithm}[!t]
\caption{Calculate $y = x^n$}\label{algo1}
\begin{algorithmic}[1]
\Require $n \geq 0 \vee x \neq 0$
\Ensure $y = x^n$
\State $y \Leftarrow 1$
\If{$n < 0$}
        \State $X \Leftarrow 1 / x$
        \State $N \Leftarrow -n$
\Else
        \State $X \Leftarrow x$
        \State $N \Leftarrow n$
\EndIf
\While{$N \neq 0$}
        \If{$N$ is even}
            \State $X \Leftarrow X \times X$
            \State $N \Leftarrow N / 2$
        \Else[$N$ is odd]
            \State $y \Leftarrow y \times X$
            \State $N \Leftarrow N - 1$
        \EndIf
\EndWhile
\end{algorithmic}
\end{algorithm}

Similarly, for \verb+listings+, one has to use the \verb+listings+ package. The \verb+\begin{lstlisting}+ \verb+...+ \verb+\end{lstlisting}+ environment is used to set environments similar to the \verb+verbatim+ environment. Refer to the \verb+lstlisting+ package documentation for more details on this.

\begin{minipage}{\hsize}%
\lstset{language=Pascal}% Set your language (you can change the language for each code-block optionally)
\begin{lstlisting}[frame=single,framexleftmargin=-1pt,framexrightmargin=-17pt,framesep=12pt,linewidth=0.98\textwidth]
for i:=maxint to 0 do
begin
{ do nothing }
end;
Write('Case insensitive ');
Write('Pascal keywords.');
\end{lstlisting}
\end{minipage}

\section{Cross referencing}\label{sec8}

Environments such as figure, table, equation, and align can have a label
declared via the \verb+\label{#label}+ command. For figures and table
environments one should use the \verb+\label{}+ command inside or just
below the \verb+\caption{}+ command.  One can then use the
\verb+\ref{#label}+ command to cross-reference them. As an example, consider
the label declared for Figure \ref{fig1} which is
\verb+\label{fig1}+. To cross-reference it, use the command
\verb+ Figure \ref{fig1}+, for which it comes up as
``Figure~\ref{fig1}".

\subsection{Details on reference citations}\label{subsec3}

With standard numerical .bst files, only numerical citations are possible.
With an author-year .bst file, both numerical and author-year citations are possible.

If author-year citations are selected, \verb+\bibitem+ must have one of the following forms:

{\footnotesize%
\begin{verbatim}
\bibitem[Jones et al.(1990)]{key}...
\bibitem[Jones et al.(1990)Jones,
                Baker, and Williams]{key}...
\bibitem[Jones et al., 1990]{key}...
\bibitem[\protect\citeauthoryear{Jones,
                Baker, and Williams}
                {Jones et al.}{1990}]{key}...
\bibitem[\protect\citeauthoryear{Jones et al.}
                {1990}]{key}...
\bibitem[\protect\astroncite{Jones et al.}
                {1990}]{key}...
\bibitem[\protect\citename{Jones et al., }
                1990]{key}...
\harvarditem[Jones et al.]{Jones, Baker, and
                Williams}{1990}{key}...
\end{verbatim}}

This is either to be made up manually, or to be generated by an
appropriate .bst file with BibTeX. Then,

{%
\begin{verbatim}
                    Author-year mode
                        || Numerical mode
\citet{key} ==>>  Jones et al. (1990)
                        || Jones et al. [21]
\citep{key} ==>> (Jones et al., 1990) || [21]
\end{verbatim}}

\noindent
Multiple citations as normal:

{%
\begin{verbatim}
\citep{key1,key2} ==> (Jones et al., 1990;
                         Smith, 1989)||[21,24]
        or (Jones et al., 1990, 1991)||[21,24]
        or (Jones et al., 1990a,b)   ||[21,24]
\end{verbatim}}

\noindent
\verb+\cite{key}+ is the equivalent of \verb+\citet{key}+ in author-year mode
and  of \verb+\citep{key}+ in numerical mode. Full author lists may be forced with
\verb+\citet*+ or \verb+\citep*+, e.g.

{%
\begin{verbatim}
\citep*{key} ==>> (Jones, Baker, and Mark, 1990)
\end{verbatim}}

\noindent
Optional notes as:

{%
\begin{verbatim}
\citep[chap. 2]{key}     ==>>
        (Jones et al., 1990, chap. 2)
\citep[e.g.,][]{key}     ==>>
        (e.g., Jones et al., 1990)
\citep[see][pg. 34]{key} ==>>
        (see Jones et al., 1990, pg. 34)
\end{verbatim}}

\noindent
(Note: in standard LaTeX, only one note is allowed, after the ref.
Here, one note is like the standard, two make pre- and post-notes.)

{%
\begin{verbatim}
\citealt{key}   ==>> Jones et al. 1990
\citealt*{key}  ==>> Jones, Baker, and
                        Williams 1990
\citealp{key}   ==>> Jones et al., 1990
\citealp*{key}  ==>> Jones, Baker, and
                        Williams, 1990
\end{verbatim}}

\noindent
Additional citation possibilities (both author-year and numerical modes):

{%
\begin{verbatim}
\citeauthor{key}       ==>> Jones et al.
\citeauthor*{key}      ==>> Jones, Baker, and
                                Williams
\citeyear{key}         ==>> 1990
\citeyearpar{key}      ==>> (1990)
\citetext{priv. comm.} ==>> (priv. comm.)
\citenum{key}          ==>> 11 [non-superscripted]
\end{verbatim}}

\noindent
Note: full author lists depend on whether the bib style supports them;
if not, the abbreviated list is printed even when full is requested.

\noindent
For names like della Robbia at the start of a sentence, use

{%
\begin{verbatim}
\Citet{dRob98}      ==>> Della Robbia (1998)
\Citep{dRob98}      ==>> (Della Robbia, 1998)
\Citeauthor{dRob98} ==>> Della Robbia
\end{verbatim}}

\noindent
The following is an example for \verb+\cite{...}+: \cite{rahman2019centroidb}. Another example for \verb+\citep{...}+: \citep{bahdanau2014neural,imboden2018cardiorespiratory,motiian2017unified,murphy2012machine,ji20123d}.
Sample cites here \cite{krizhevsky2012imagenet,horvath2018dna} and \cite{pyrkov2018quantitative}, \cite{wang2018face}, \cite{lecun2015deep,zhang2018fine,ravi2016deep}.

\section{Lists}\label{sec9}

List in \LaTeX{} can be of three types: numbered, bulleted and unnumbered. The ``enumerate'' environment produces a numbered list, the 
``itemize'' environment produces a bulleted list and the ``unlist''
environment produces an unnumbered list.
In each environment, a new entry is added via the \verb+\item+ command.
\begin{enumerate}[1.]
\item This is the 1st item

\item Enumerate creates numbered lists, itemize creates bulleted lists and
unnumerate creates unnumbered lists.
\begin{enumerate}[(a)]
\item Second level numbered list. Enumerate creates numbered lists, itemize creates bulleted lists and
description creates unnumbered lists.

\item Second level numbered list. Enumerate creates numbered lists, itemize creates bulleted lists and
description creates unnumbered lists.
\begin{enumerate}[(ii)]
\item Third level numbered list. Enumerate creates numbered lists, itemize creates bulleted lists and
description creates unnumbered lists.

\item Third level numbered list. Enumerate creates numbered lists, itemize creates bulleted lists and
description creates unnumbered lists.
\end{enumerate}

\item Second level numbered list. Enumerate creates numbered lists, itemize creates bulleted lists and
description creates unnumbered lists.

\item Second level numbered list. Enumerate creates numbered lists, itemize creates bulleted lists and
description creates unnumbered lists.
\end{enumerate}

\item Enumerate creates numbered lists, itemize creates bulleted lists and
description creates unnumbered lists.

\item Numbered lists continue.
\end{enumerate}
Lists in \LaTeX{} can be of three types: enumerate, itemize and description.
In each environment, a new entry is added via the \verb+\item+ command.
\begin{itemize}
\item First level bulleted list. This is the 1st item

\item First level bulleted list. Itemize creates bulleted lists and description creates unnumbered lists.
\begin{itemize}
\item Second level dashed list. Itemize creates bulleted lists and description creates unnumbered lists.

\item Second level dashed list. Itemize creates bulleted lists and description creates unnumbered lists.

\item Second level dashed list. Itemize creates bulleted lists and description creates unnumbered lists.
\end{itemize}

\item First level bulleted list. Itemize creates bulleted lists and description creates unnumbered lists.

\item First level bulleted list. Bullet lists continue.
\end{itemize}

\noindent
Example for unnumbered list items:

\begin{unlist}
\item Sample unnumberd list text. Sample unnumberd list text. Sample unnumberd list text. Sample unnumberd list text. Sample unnumberd list text.

\item Sample unnumberd list text. Sample unnumberd list text. Sample unnumberd list text.

\item sample unnumberd list text. Sample unnumberd list text. Sample unnumberd list text. Sample unnumberd list text. Sample unnumberd list text. Sample unnumberd list text. Sample unnumberd list text.
\end{unlist}

\section{Examples for theorem-like environments}\label{sec10}

For theorem-like environments, we require the \verb+amsthm+ package. There are three types of predefined theorem styles - \verb+thmstyleone+, \verb+thmstyletwo+ and \verb+thmstylethree+   (check your journal's instructions page in case a specific style is required).

\medskip
\noindent\begin{tabular}{|l|p{13pc}|}
\hline
\verb+thmstyleone+ & Numbered, theorem head in bold font and theorem text in italic style \\\hline
\verb+thmstyletwo+ & Numbered, theorem head in roman font and theorem text in italic style \\\hline
\verb+thmstylethree+ & Numbered, theorem head in bold font and theorem text in roman style \\\hline
\end{tabular}

\begin{theorem}[Theorem subhead]\label{thm1}
Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text.
Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text.
Example theorem text.
\end{theorem}

Quisque ullamcorper placerat ipsum. Cras nibh. Morbi vel justo vitae lacus tincidunt ultrices. Lorem ipsum dolor sit
amet, consectetuer adipiscing elit. In hac habitasse platea dictumst. Integer tempus convallis augue.

\begin{proposition}
Example proposition text. Example proposition text. Example proposition text. Example proposition text. Example proposition text.
Example proposition text. Example proposition text. Example proposition text. Example proposition text. Example proposition text.
\end{proposition}

Nulla malesuada porttitor diam. Donec felis erat, congue non, volutpat at, tincidunt tristique, libero. Vivamus
viverra fermentum felis. Donec nonummy pellentesque ante.

\begin{example}
Phasellus adipiscing semper elit. Proin fermentum massa
ac quam. Sed diam turpis, molestie vitae, placerat a, molestie nec, leo. Maecenas lacinia. Nam ipsum ligula, eleifend
at, accumsan nec, suscipit a, ipsum. Morbi blandit ligula feugiat magna. Nunc eleifend consequat lorem.
\end{example}

Nulla malesuada porttitor diam. Donec felis erat, congue non, volutpat at, tincidunt tristique, libero. Vivamus
viverra fermentum felis. Donec nonummy pellentesque ante.

\begin{remark}
Phasellus adipiscing semper elit. Proin fermentum massa
ac quam. Sed diam turpis, molestie vitae, placerat a, molestie nec, leo. Maecenas lacinia. Nam ipsum ligula, eleifend
at, accumsan nec, suscipit a, ipsum. Morbi blandit ligula feugiat magna. Nunc eleifend consequat lorem.
\end{remark}

Quisque ullamcorper placerat ipsum. Cras nibh. Morbi vel justo vitae lacus tincidunt ultrices. Lorem ipsum dolor sit
amet, consectetuer adipiscing elit. In hac habitasse platea dictumst.

\begin{definition}[Definition sub head]
Example definition text. Example definition text. Example definition text. Example definition text. Example definition text. Example definition text. Example definition text. Example definition text.
\end{definition}

Apart from the above styles, we have the \verb+\begin{proof}+ \verb+...+ \verb+\end{proof}+ environment - with the proof head in italic style and the body text in roman font with an open square at the end of each proof environment.

\begin{proof}Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text.
\end{proof}

Nam dui ligula, fringilla a, euismod sodales, sollicitudin vel, wisi. Morbi auctor lorem non justo. Nam lacus libero,
pretium at, lobortis vitae, ultricies et, tellus. Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna,
vitae ornare odio metus a mi.

\begin{proof}[Proof of Theorem~{\upshape\ref{thm1}}]
Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text. Example for proof text.
\end{proof}

\noindent
For a quote environment, one has to use\newline \verb+\begin{quote}...\end{quote}+
\begin{quote}
Quoted text example. Aliquam porttitor quam a lacus. Praesent vel arcu ut tortor cursus volutpat. In vitae pede quis diam bibendum placerat. Fusce elementum
convallis neque. Sed dolor orci, scelerisque ac, dapibus nec, ultricies ut, mi. Duis nec dui quis leo sagittis commodo.
\end{quote}
Donec congue. Maecenas urna mi, suscipit in, placerat ut, vestibulum ut, massa. Fusce ultrices nulla et nisl (refer Figure~\ref{fig3}). Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas. Etiam ligula arcu,
elementum a, venenatis quis, sollicitudin sed, metus. Donec nunc pede, tincidunt in, venenatis vitae, faucibus vel (refer Table~\ref{tab3}).

\section{Conclusion}

Some Conclusions here.

%%%%%%%%%%%%%%

\begin{appendices}

\section{Section title of first appendix}\label{sec11}

Nam dui ligula, fringilla a, euismod sodales, sollicitudin vel, wisi. Morbi auctor lorem non justo. Nam lacus libero,
pretium at, lobortis vitae, ultricies et, tellus. Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna,
vitae ornare odio metus a mi. Morbi ac orci et nisl hendrerit mollis. Suspendisse ut massa. Cras nec ante. Pellentesque
a nulla. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Aliquam tincidunt
urna. Nulla ullamcorper vestibulum turpis. Pellentesque cursus luctus mauris.

\subsection{Subsection title of first appendix}\label{subsec4}

Nam dui ligula, fringilla a, euismod sodales, sollicitudin vel, wisi. Morbi auctor lorem non justo. Nam lacus libero,
pretium at, lobortis vitae, ultricies et, tellus. Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna,
vitae ornare odio metus a mi. Morbi ac orci et nisl hendrerit mollis. Suspendisse ut massa. Cras nec ante. Pellentesque
a nulla. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Aliquam tincidunt
urna. Nulla ullamcorper vestibulum turpis. Pellentesque cursus luctus mauris.

\subsubsection{Subsubsection title of first appendix}\label{subsubsec3}

Example for an unnumbered figure:

\begin{figure}[!h]
\centering
{\color{black!20}\rule{85pt}{92pt}}
\end{figure}

Fusce mauris. Vestibulum luctus nibh at lectus. Sed bibendum, nulla a faucibus semper, leo velit ultricies tellus, ac
venenatis arcu wisi vel nisl. Vestibulum diam. Aliquam pellentesque, augue quis sagittis posuere, turpis lacus congue
quam, in hendrerit risus eros eget felis.

\section{Section title of second appendix}\label{sec12}%

Fusce mauris. Vestibulum luctus nibh at lectus. Sed bibendum, nulla a faucibus semper, leo velit ultricies tellus, ac
venenatis arcu wisi vel nisl. Vestibulum diam. Aliquam pellentesque, augue quis sagittis posuere, turpis lacus congue
quam, in hendrerit risus eros eget felis. Maecenas eget erat in sapien mattis porttitor. Vestibulum porttitor. Nulla
facilisi. Sed a turpis eu lacus commodo facilisis. Morbi fringilla, wisi in dignissim interdum, justo lectus sagittis dui, et
vehicula libero dui cursus dui. Mauris tempor ligula sed lacus. Duis cursus enim ut augue. Cras ac magna. Cras nulla.

\begin{figure}[b]
\centering
{\color{black!20}\rule{217pt}{120pt}}
\caption{This is an example for appendix figure\label{fig4}}
\end{figure}

\begin{table}[t]%
\begin{center}
\begin{minipage}{.52\columnwidth}
\caption{This is an example of Appendix table showing food requirements of army, navy and airforce\label{tab4}}%
\begin{tabular}{@{}lcc@{}}%
\toprule
col1 head & col2 head & col3 head \\
\midrule
col1 text & col2 text & col3 text \\
col1 text & col2 text & col3 text \\
col1 text & col2 text & col3 text \\
\botrule
\end{tabular}
\end{minipage}
\end{center}
\end{table}

\subsection{Subsection title of second appendix}\label{subsec5}

Sed commodo posuere pede. Mauris ut est. Ut quis purus. Sed ac odio. Sed vehicula hendrerit sem. Duis non odio.
Morbi ut dui. Sed accumsan risus eget odio. In hac habitasse platea dictumst. Pellentesque non elit. Fusce sed justo
eu urna porta tincidunt. Mauris felis odio, sollicitudin sed, volutpat a, ornare ac, erat. Morbi quis dolor. Donec
pellentesque, erat ac sagittis semper, nunc dui lobortis purus, quis congue purus metus ultricies tellus. Proin et quam.
Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos hymenaeos. Praesent sapien turpis,
fermentum vel, eleifend faucibus, vehicula eu, lacus.

Sed commodo posuere pede. Mauris ut est. Ut quis purus. Sed ac odio. Sed vehicula hendrerit sem. Duis non odio.
Morbi ut dui. Sed accumsan risus eget odio. In hac habitasse platea dictumst. Pellentesque non elit. Fusce sed justo
eu urna porta tincidunt. Mauris felis odio, sollicitudin sed, volutpat a, ornare ac, erat. Morbi quis dolor. Donec
pellentesque, erat ac sagittis semper, nunc dui lobortis purus, quis congue purus metus ultricies tellus. Proin et quam.
Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos hymenaeos. Praesent sapien turpis,
fermentum vel, eleifend faucibus, vehicula eu, lacus.

\subsubsection{Subsubsection title of second appendix}\label{subsubsec4}

Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae,
felis. Curabitur dictum gravida mauris. Nam arcu libero, nonummy eget, consectetuer id, vulputate a, magna. Donec
vehicula augue eu neque.

Example for an equation inside the appendix:
\begin{align}
 p &= \frac{\gamma^{2} - (n_{C} -1)H}{(n_{C} - 1) + H - 2\gamma}, \label{1eq:hybobo:pfromgH} \\
 \theta &= \frac{(\gamma - H)^{2}(\gamma - n_{C} -1)^{2}}{(n_{C} - 1 + H - 2\gamma)^{2}} \label{2eq:hybobo:tfromgH}\; .
\end{align}

\section{Example of another appendix section}\label{sec13}%

Nam dui ligula, fringilla a, euismod sodales, sollicitudin vel, wisi. Morbi auctor lorem non justo. Nam lacus libero,
pretium at, lobortis vitae, ultricies et, tellus. Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna,
vitae ornare odio metus a mi. Morbi ac orci et nisl hendrerit mollis. Suspendisse ut massa. Cras nec ante. Pellentesque
a nulla. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Aliquam tincidunt
urna. Nulla ullamcorper vestibulum turpis. Pellentesque cursus luctus mauris
\begin{equation}
\mathcal{L} = i \bar{\psi} \gamma^\mu D_\mu \psi
    - \frac{1}{4} F_{\mu\nu}^a F^{a\mu\nu} - m \bar{\psi} \psi.
\label{eq26}
\end{equation}

Nulla malesuada porttitor diam. Donec felis erat, congue non, volutpat at, tincidunt tristique, libero. Vivamus
viverra fermentum felis. Donec nonummy pellentesque ante. Phasellus adipiscing semper elit. Proin fermentum massa
ac quam. Sed diam turpis, molestie vitae, placerat a, molestie nec, leo. Maecenas lacinia. Nam ipsum ligula, eleifend
at, accumsan nec, suscipit a, ipsum. Morbi blandit ligula feugiat magna. Nunc eleifend consequat lorem. Sed lacinia
nulla vitae enim. Pellentesque tincidunt purus vel magna. Integer non enim. Praesent euismod nunc eu purus. Donec
bibendum quam in tellus. Nullam cursus pulvinar lectus. Donec et mi. Nam vulputate metus eu enim. Vestibulum
pellentesque felis eu massa.

Nulla malesuada porttitor diam. Donec felis erat, congue non, volutpat at, tincidunt tristique, libero. Vivamus
viverra fermentum felis. Donec nonummy pellentesque ante. Phasellus adipiscing semper elit. Proin fermentum massa
ac quam. Sed diam turpis, molestie vitae, placerat a, molestie nec, leo. Maecenas lacinia. Nam ipsum ligula, eleifend
at, accumsan nec, suscipit a, ipsum. Morbi blandit ligula feugiat magna. Nunc eleifend consequat lorem. Sed lacinia
nulla vitae enim. Pellentesque tincidunt purus vel magna. Integer non enim. Praesent euismod nunc eu purus. Donec
bibendum quam in tellus. Nullam cursus pulvinar lectus. Donec et mi. Nam vulputate metus eu enim. Vestibulum
pellentesque felis eu massa.

Nulla malesuada porttitor diam. Donec felis erat, congue non, volutpat at, tincidunt tristique, libero. Vivamus
viverra fermentum felis. Donec nonummy pellentesque ante. Phasellus adipiscing semper elit. Proin fermentum massa
ac quam. Sed diam turpis, molestie vitae, placerat a, molestie nec, leo. Maecenas lacinia. Nam ipsum ligula, eleifend
at, accumsan nec, suscipit a, ipsum. Morbi blandit ligula feugiat magna. Nunc eleifend consequat lorem. Sed lacinia
nulla vitae enim. Pellentesque tincidunt purus vel magna. Integer non enim. Praesent euismod nunc eu purus. Donec
bibendum quam in tellus. Nullam cursus pulvinar lectus. Donec et mi. Nam vulputate metus eu enim. Vestibulum
pellentesque felis eu massa.

Nulla malesuada porttitor diam. Donec felis erat, congue non, volutpat at, tincidunt tristique, libero. Vivamus
viverra fermentum felis. Donec nonummy pellentesque ante. Phasellus adipiscing semper elit. Proin fermentum massa
ac quam. Sed diam turpis, molestie vitae, placerat a, molestie nec, leo. Maecenas lacinia. Nam ipsum ligula, eleifend
at, accumsan nec, suscipit a, ipsum. Morbi blandit ligula feugiat magna. Nunc eleifend consequat lorem. Sed lacinia
nulla vitae enim. Pellentesque tincidunt purus vel magna. Integer non enim. Praesent euismod nunc eu purus. Donec
bibendum quam in tellus. Nullam cursus pulvinar lectus. Donec et mi. Nam vulputate metus eu enim. Vestibulum
pellentesque felis eu massa. Donec
bibendum quam in tellus. Nullam cursus pulvinar lectus. Donec et mi. Nam vulputate metus eu enim. Vestibulum
pellentesque felis eu massa.

%% Example for unnumbered table inside appendix
\begin{table}
\begin{center}
\begin{minipage}{.52\columnwidth}
\caption{}{%
\begin{tabular}{lcc}%
\toprule
col1 head & col2 head & col3 head \\
\midrule
col1 text & col2 text & col3 text \\
col1 text & col2 text & col3 text \\
col1 text & col2 text & col3 text \\
\botrule
\end{tabular}}{}
\end{minipage}
\end{center}
\end{table}

\end{appendices}

\section{Competing interests}
No competing interest is declared.

\section{Author contributions statement}

Must include all authors, identified by initials, for example:
S.R. and D.A. conceived the experiment(s),  S.R. conducted the experiment(s), S.R. and D.A. analysed the results.  S.R. and D.A. wrote and reviewed the manuscript.

\section{Acknowledgments}
The authors thank the anonymous reviewers for their valuable suggestions. This work is supported in part by funds from the National Science Foundation (NSF: \# 1636933 and \# 1920920).

%\bibliographystyle{plain}
%\bibliography{reference}

\begin{thebibliography}{10}

\bibitem{bahdanau2014neural}
Dzmitry Bahdanau, Kyunghyun Cho, and Yoshua Bengio.
\newblock Neural machine translation by jointly learning to align and
  translate.
\newblock {\em arXiv preprint arXiv:1409.0473}, 2014.

\bibitem{horvath2018dna}
Steve Horvath and Kenneth Raj.
\newblock Dna methylation-based biomarkers and the epigenetic clock theory of
  ageing.
\newblock {\em Nature Reviews Genetics}, 19(6):371, 2018.

\bibitem{imboden2018cardiorespiratory}
Mary~T Imboden, Matthew~P Harber, Mitchell~H Whaley, W~Holmes Finch, Derron~L
  Bishop, and Leonard~A Kaminsky.
\newblock Cardiorespiratory fitness and mortality in healthy men and women.
\newblock {\em Journal of the American College of Cardiology},
  72(19):2283--2292, 2018.

\bibitem{ji20123d}
Shuiwang Ji, Wei Xu, Ming Yang, and Kai Yu.
\newblock 3d convolutional neural networks for human action recognition.
\newblock {\em IEEE Transactions on Pattern Analysis and Machine Intelligence},
  35(1):221--231, 2012.

\bibitem{krizhevsky2012imagenet}
Alex Krizhevsky, Ilya Sutskever, and Geoffrey~E Hinton.
\newblock Imagenet classification with deep convolutional neural networks.
\newblock In {\em Advances in Neural Information Processing Systems}, pages
  1097--1105, 2012.

\bibitem{lecun2015deep}
Yann LeCun, Yoshua Bengio, and Geoffrey Hinton.
\newblock Deep learning.
\newblock {\em Nature}, 521(7553):436, 2015.

\bibitem{motiian2017unified}
Saeid Motiian, Marco Piccirilli, Donald~A Adjeroh, and Gianfranco Doretto.
\newblock Unified deep supervised domain adaptation and generalization.
\newblock In {\em Proceedings of the IEEE International Conference on Computer
  Vision}, pages 5715--5725, 2017.

\bibitem{murphy2012machine}
Kevin~P Murphy.
\newblock {\em Machine learning: A probabilistic perspective}.
\newblock MIT press, 2012.

\bibitem{american2013acsm}
American~College of~Sports~Medicine et~al.
\newblock {\em ACSM's guidelines for exercise testing and prescription}.
\newblock Lippincott Williams \& Wilkins, 2013.

\bibitem{pyrkov2018quantitative}
Timothy~V Pyrkov, Evgeny Getmantsev, Boris Zhurov, Konstantin Avchaciov,
  Mikhail Pyatnitskiy, Leonid Menshikov, Kristina Khodova, Andrei~V Gudkov, and
  Peter~O Fedichev.
\newblock Quantitative characterization of biological age and frailty based on
  locomotor activity records.
\newblock {\em Aging (Albany NY)}, 10(10):2973, 2018.

\bibitem{rahman2019centroidb}
Syed~Ashiqur Rahman and Donald Adjeroh.
\newblock Centroid of age neighborhoods: A generalized approach to estimate
  biological age.
\newblock In {\em 2019 IEEE EMBS International Conference on Biomedical \&
  Health Informatics (BHI)}, pages 1--4. IEEE, 2019.

\bibitem{ravi2016deep}
Daniele Rav{\`\i}, Charence Wong, Fani Deligianni, Melissa Berthelot, Javier
  Andreu-Perez, Benny Lo, and Guang-Zhong Yang.
\newblock Deep learning for health informatics.
\newblock {\em IEEE {J}ournal of {B}iomedical and {H}ealth {I}nformatics},
  21(1):4--21, 2016.

\bibitem{wang2018face}
Zongwei Wang, Xu~Tang, Weixin Luo, and Shenghua Gao.
\newblock Face aging with identity-preserved conditional generative adversarial
  networks.
\newblock In {\em Proceedings of the IEEE Conference on Computer Vision and
  Pattern Recognition}, pages 7939--7947, 2018.

\bibitem{zhang2018fine}
Ke~Zhang, Na~Liu, Xingfang Yuan, Xinyao Guo, Ce~Gao, and Zhenbing Zhao.
\newblock Fine-grained age estimation in the wild with attention {LSTM}
  networks.
\newblock {\em arXiv preprint arXiv:1805.10445}, 2018.

\end{thebibliography}

%USE THE BELOW OPTIONS IN CASE YOU NEED AUTHOR YEAR FORMAT.
%\bibliographystyle{abbrvnat}
%\bibliography{reference}

%% sample for biography with author's image
\begin{biography}{{\color{black!20}\rule{77pt}{77pt}}}{\author{Author Name.} This is sample author biography text. The values provided in the optional argument are meant for sample purposes. There is no need to include the width and height of an image in the optional argument for live articles. This is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text.}
\end{biography}

%% sample for biography without author's image
\begin{biography}{}{\author{Author Name.} This is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text.}
\end{biography}

\end{document}

### 2023-11-14 Manuscript/manuscript.tex
Format manuscript according Bio Adv template

@@
 \author[3,4]{Melissa Grant-Peters}
-\author[2]{Leonardo Collado-Torres}
+\author[1,2]{Leonardo Collado-Torres}
 \author[1,5,6,7]{Stephanie C. Hicks\thanks{Correspondence to \url{shicks19@jhu.edu}}}
@@
 Here, we focus on two types of visualizations in biomedical research, but note that the proposed ideas could be extended beyond these applications: (i) embedding visualizations, which project data into some low-dimensional embedding or mathematical space (e.g.
-Principal Components Analysis %(PCA)
-\cite{hotelling_1933}, $t$-distributed Stochastic Neighbor Embedding ($t$-SNE) \cite{vandermaaten_2008}, or Uniform Manifold Approximation and Projection (UMAP) \cite{becht_2019}) and (ii) \textit{in situ} visualizations \cite{dries_2021, Lewis_2021, odonoghue_2021}, which aim to visualize molecules captured from \textit{in situ} imaging or sequencing technologies where \textit{in situ} refers to `in its original place'.
-Both of these visualizations represent data in a 2D space and are motivated by recent advances in experimental technologies that profile molecules, including DNA, RNA, and proteins, at a single-cell or spatial resolution \cite{kashima_2020, moffitt_2022}.
-Some most popular technologies include single-cell/nucleus RNA-sequencing (sc/snRNA-seq)
-\cite{aldridge_2020} and \textit{in situ} spatially-resolved transcriptomics \cite{larsson_2021}. 
+Principal Components Analysis , $t$-distributed Stochastic Neighbor Embedding ($t$-SNE), and Uniform Manifold Approximation and Projection (UMAP)) and (ii) \textit{in situ} visualizations \cite{dries_2021}, which aim to visualize molecules captured from \textit{in situ} imaging or sequencing technologies where \textit{in situ} refers to `in its original place'.
+Both of these visualizations represent data in a 2D space and are motivated by recent advances in experimental technologies that profile molecules, including DNA, RNA, and proteins, at a single-cell or spatial resolution% \cite{kashima_2020, moffitt_2022}
+. Some most popular technologies include single-cell/nucleus RNA-sequencing (sc/snRNA-seq)
+and \textit{in situ} spatially-resolved transcriptomics. \cite{vandereyken2023methods} 
 % 
@@
 
-To address these challenges, here we leverage the Gestalt (German for “unified whole”) principles \cite{todorovic_2008, palmer_1999} as a path to visualize multi-dimensional data in 2D visualizations.
+To address these challenges, here we leverage the Gestalt (German for “unified whole”) principles \cite{todorovic_2008} as a path to visualize multi-dimensional data in 2D visualizations.
 We focus on the two types of data visualizations previously introduced that are widely used in biomedical research: (i) embedding visualizations and (ii) \textit{in situ} visualizations.
@@
 
-Gestalt principles \cite{todorovic_2008, palmer_1999} refer to a set of rules describing how humans perceive and interpret visual information and are commonly applied in art and designs.
+Gestalt principles \cite{todorovic_2008} refer to a set of rules describing how humans perceive and interpret visual information and are commonly applied in art and designs.
 Developed in the 1920s by German psychologists Max Wertheimer, Kurt Koffka and Wolfgang Kohler, these principles help humans perceive a set of individual elements as a whole.
@@
 Our approach is to use the state-of-art data visualization framework \texttt{ggplot2} \cite{ggplot2} following the Grammar of Graphics \cite{wilkinson_2012} and map individual variables to different aesthetics to simultaneously display disparate variables.
-Specifically, we apply the figure-ground segmentation \cite{peterson_2010} in displaying two variables: one variable (e.g.
+Specifically, we apply the figure-ground segmentation \cite{todorovic_2008} in displaying two variables: one variable (e.g.
 expression) can be plotted as color-filled circles, serving as the \textit{figure}; one variable (e.g.
@@
 
-For adjacent circles with limited space between them to display the background color, we use an economic implementation, colored outlines for these circles (\textbf{Figure~\ref{fig:visual}C}), inspired by watercolor effect \cite{pinna_1987, pinna_2001}.
+For adjacent circles with limited space between them to display the background color, we use an economic implementation, colored outlines for these circles (\textbf{Figure~\ref{fig:visual}C}), inspired by watercolor effect \cite{pinna_2001}.
 Watercolor effect describes the phenomenon in visual perception that surface color arises from thin boundaries and hence is applied here to perceive the background color in tight space.
@@
 To address the overplotting problem, previous work proposed to apply hexagonal binning strategy to display the gene expression \cite{freytag_2020}.
-Here, the color-coded convex hulls are used to annotate different clusters of cells (\textbf{Figure~\ref{fig:embedding}A}).
+Here, the color-coded convex hulls are used to annotate different clusters of \rr{hexagons, where the cluster label of each hexagon is determined by the most popular label within each hexagon bin} (\textbf{Figure~\ref{fig:embedding}A}).
 However, the convex hulls create substantial overlapping areas, creating confusion when interpreting cluster memberships of hexagons in the overlapping areas.
@@
 
-While the proposed design principle set-up a solid foundation, the proposed package \texttt{escheR} can be further optimized for image-based spatially resolved data \cite{caicedo_2017}.
-For example, to address possible overplotting of densely arranged single cells in tight space, strategies such as binning or hatching plots \cite{patrick2023spatial} can be incorporated.
+While the proposed design principle set-up a solid foundation, the proposed package \texttt{escheR} can be further optimized for image-based spatially resolved data.
+For example, to address possible overplotting of densely arranged single cells in tight space, strategies such as hatching plots \cite{patrick2023spatial} can be incorporated.
 In addition, the flexibility to represent data points as polygons will be provided to allow visualization of cell morphology.
@@
 Eagles and Kristen R.
-Maynard are employed by the Lieber Institute for Brain Development; Mina Ryten is employed by University College London; Leon Di Stefano and Lukas M.
-Weber are from the Johns Hopkins Bloomberg School of Public Health, Department of Biostatistics.
+Maynard are employed by the Lieber Institute for Brain Development; Mina Ryten is employed by University College London; Leon Di Stefano is from the Johns Hopkins Bloomberg School of Public Health, Department of Biostatistics; Lukas M.
+Weber is in the Department of Biostatistics at Boston University School of Public Health.
 
@@
     \item Boyi Guo (ORCiD: \href{https://orcid.org/0000-0003-2950-2349}{0000-0003-2950-2349})
-    \item Erik D. Nelson (ORCiD: \href{https://orcid.org/0000-0001-8477-0982}{0000-0001-8477-0982})
     \item Louise A. Huuki-Myers (ORCiD: \href{https://orcid.org/0000-0001-5148-3602}{0000-0001-5148-3602})

### 2023-11-14 Manuscript/manuscript_BioAdv_format.tex
Format manuscript according Bio Adv template

@@
 %%%CONTEMPORARY%%%
-\documentclass[unnumsec,webpdf,contemporary,large]{oup-authoring-template}%
-%\documentclass[unnumsec,webpdf,contemporary,large,namedate]{oup-authoring-template}% uncomment this line for author year citations and comment the above
-%\documentclass[unnumsec,webpdf,contemporary,medium]{oup-authoring-template}
+% \documentclass[unnumsec,webpdf,contemporary,large]{oup-authoring-template}%
+% \documentclass[unnumsec,webpdf,contemporary,large,namedate]{oup-authoring-template}% uncomment this line for author year citations and comment the above
+% \documentclass[unnumsec,webpdf,contemporary,medium]{oup-authoring-template}
 %\documentclass[unnumsec,webpdf,contemporary,small]{oup-authoring-template}
@@
 %\documentclass[unnumsec,webpdf,modern,large]{oup-authoring-template}
-%\documentclass[unnumsec,webpdf,modern,large,namedate]{oup-authoring-template}% uncomment this line for author year citations and comment the above
+\documentclass[unnumsec,webpdf,modern,large,namedate]{oup-authoring-template}% uncomment this line for author year citations and comment the above
 %\documentclass[unnumsec,webpdf,modern,medium]{oup-authoring-template}
@@
 %\documentclass[unnumsec,webpdf,traditional,large]{oup-authoring-template}
-%\documentclass[unnumsec,webpdf,traditional,large,namedate]{oup-authoring-template}% uncomment this line for author year citations and comment the above
+% \documentclass[unnumsec,webpdf,traditional,large,namedate]{oup-authoring-template}% uncomment this line for author year citations and comment the above
 %\documentclass[unnumsec,namedate,webpdf,traditional,medium]{oup-authoring-template}
@@
 
-%\onecolumn % for one column layouts
+\onecolumn % for one column layouts
 
@@
 % line numbers
-\usepackage[mathlines, switch]{lineno}
+% \usepackage[mathlines, switch]{lineno}
 \usepackage[right]{lineno}
+\linenumbers
 
@@
 \newtheorem{definition}{Definition}
+
+\newcommand{\rr}[1]{{\color{red} #1}}
+\usepackage{setspace}
+\doublespacing
 
@@
 
-\title[Short Article Title]{Article Title}
+\title[escheR]{\texttt{escheR}: Unified multi-dimensional visualizations with Gestalt principles}
 
-\author[1,$\ast$]{First Author}
-\author[2]{Second Author}
-\author[3]{Third Author}
-\author[3]{Fourth Author}
-\author[4]{Fifth Author\ORCID{0000-0000-0000-0000}}
+% \author[1,$\ast$]{First Author}
+% \author[2]{Second Author}
+% \author[3]{Third Author}
+% \author[3]{Fourth Author}
+% \author[4]{Fifth Author\ORCID{0000-0000-0000-0000}}
 
-\authormark{Author Name et al.}
+\author[1]{Boyi Guo\ORCID{0000-0003-2950-2349}}
+\author[2]{Louise A. Huuki-Myers\ORCID{0000-0001-5148-3602}}
+\author[3,4]{Melissa Grant-Peters\ORCID{0000-0003-0585-0971}}
+\author[1,2]{Leonardo Collado-Torres\ORCID{0000-0003-2140-308X}}
+\author[1,5,6,7, $\ast$]{Stephanie C. Hicks\ORCID{0000-0002-7858-0231}}
 
-\address[1]{\orgdiv{Department}, \orgname{Organization}, \orgaddress{\street{Street}, \postcode{Postcode}, \state{State}, \country{Country}}}
-\address[2]{\orgdiv{Department}, \orgname{Organization}, \orgaddress{\street{Street}, \postcode{Postcode}, \state{State}, \country{Country}}}
-\address[3]{\orgdiv{Department}, \orgname{Organization}, \orgaddress{\street{Street}, \postcode{Postcode}, \state{State}, \country{Country}}}
-\address[4]{\orgdiv{Department}, \orgname{Organization}, \orgaddress{\street{Street}, \postcode{Postcode}, \state{State}, \country{Country}}}
+\authormark{B.Guo et al.}
 
-\corresp[$\ast$]{Corresponding author. \href{email:email-id.com}{email-id.com}}
+\address[1]{Department of Biostatistics, Johns Hopkins Bloomberg School of Public Health, MD, USA} 
+\address[2]{Lieber Institute for Brain Development, Johns Hopkins Medical Campus, Baltimore, MD, USA}
+\address[3]{Genetics and Genomic Medicine, Great Ormond Street Institute of Child Health, University College London, London, UK}
+\address[4]{Aligning Science Across Parkinson’s (ASAP) Collaborative Research Network, Chevy Chase, MD, USA}
+\address[5]{Department of Biomedical Engineering, Johns Hopkins School of Medicine, Baltimore, MD, USA}
+\address[6]{Center for Computational Biology, Johns Hopkins University, Baltimore, MD, USA}
+\address[7]{Malone Center for Engineering in Healthcare, Johns Hopkins University, MD, USA}
+
+% \address[1]{\orgdiv{Department}, \orgname{Organization}, \orgaddress{\street{Street}, \postcode{Postcode}, \state{State}, \country{Country}}}
+% \address[2]{\orgdiv{Department}, \orgname{Organization}, \orgaddress{\street{Street}, \postcode{Postcode}, \state{State}, \country{Country}}}
+% \address[3]{\orgdiv{Department}, \orgname{Organization}, \orgaddress{\street{Street}, \postcode{Postcode}, \state{State}, \country{Country}}}
+% \address[4]{\orgdiv{Department}, \orgname{Organization}, \orgaddress{\street{Street}, \postcode{Postcode}, \state{State}, \country{Country}}}
+
+\corresp[$\ast$]{Corresponding author. \textbf{Contact}: \href{email:shicks19@jhu.edu}{shicks19@jhu.edu}}
 
@@
 
-\abstract{Abstracts must be able to stand alone and so cannot contain citations to
-the paper's references, equations, etc. An abstract must consist of a single
-paragraph and be concise. Because of online formatting, abstracts must appear
-as plain as possible.}
-\keywords{keyword1, Keyword2, Keyword3, Keyword4}
+\abstract{The creation of effective visualizations is a fundamental component of data analysis.
+In biomedical research, new challenges are emerging to visualize multi-dimensional data in a 2D space, but current data visualization tools have limited capabilities.
+To address this problem, we leverage Gestalt principles to improve the design and interpretability of multi-dimensional data in 2D data visualizations, layering aesthetics to display multiple variables.
+The proposed visualization can be applied to spatially-resolved transcriptomics data, but also broadly to data visualized in 2D space, such as embedding visualizations.
+We provide an open source R package \texttt{escheR}, which is built off of the state-of-the-art \texttt{ggplot2} visualization framework and can be seamlessly integrated into genomics toolboxes and workflows.\\
+\textbf{Availability and implementation}: The open source R package \texttt{escheR} is freely available on Bioconductor (\href{https://bioconductor.org/packages/escheR}{https://bioconductor.org/packages/escheR}).}
+\keywords{Visualization, Software, Single Cell, Multi-omics, Transcriptomics}
 
@@
 \section{Introduction}
-The introduction introduces the context and summarizes the manuscript. It is importantly to clearly state the contributions
-of this piece of work.
-Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
-Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
-Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
-Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
+Visualization is an indispensable component of data analysis, providing clarity that connects quantitative evidence to key conclusions \citep{dagostinomcgowan_2022}.
+In biomedical research, visualization receives growing recognition as essential: many scientists rely on visualization to complete their cognitive process from analysis to insight, including analytic validation of automated pipelines and scientific communication \citep{odonoghue_2021}.
+However, an important challenge in biomedical research is how to visualize increasingly complex, multi-dimensional data \citep{odonoghue_2010}.
 
-This is an example of a new parapgraph with a numbered footnote\footnote{\url{https://data.gov.uk/}} and a second footnote marker.\footnote{Example of footnote text.}
+Here, we focus on two types of visualizations in biomedical research, but note that the proposed ideas could be extended beyond these applications: (i) embedding visualizations, which project data into some low-dimensional embedding or mathematical space (e.g.
+Principal Components Analysis , $t$-distributed Stochastic Neighbor Embedding ($t$-SNE), and Uniform Manifold Approximation and Projection (UMAP)) and (ii) \textit{in situ} visualizations \citep{dries_2021}, which aim to visualize molecules captured from \textit{in situ} imaging or sequencing technologies where \textit{in situ} refers to `in its original place'.
+Both of these visualizations represent data in a 2D space and are motivated by recent advances in experimental technologies that profile molecules, including DNA, RNA, and proteins, at a single-cell or spatial resolution% \citep{kashima_2020, moffitt_2022}
+. Some most popular technologies include single-cell/nucleus RNA-sequencing (sc/snRNA-seq)
+and \textit{in situ} spatially-resolved transcriptomics. \citep{vandereyken2023methods} 
+% 
+% \vspace{-0.15in}
 
-\section{This is an example for first level head - section head}\label{sec2}
+A common and fundamental challenge with both of these visualizations is how to visualize multi-dimensional information in a 2D space.
+For example, in \textit{in situ} visualizations, we often want to create a spatial map to visualize a continuous (e.g.
+gene expression) or discrete (e.g.
+cell type or spatial domain) variable representing molecular information in the original spatial location.
+However, it is challenging to simultaneously visualize multi-dimensional data, such as information from disparate data domains (such as expression domain and spatial domain) or disparate data modalities (such as transcriptomics and proteomics) in the same plot.
+Currently, best practices for this include making two different plots displayed side-by-side (\textbf{Figure \textbf{\ref{fig:visual}A-B}}), one for gene expression and one for spatial domains.
+This creates cognitive gaps on how to associate the disparate information or how to interpret the biological findings of this multi-dimensional information regarding their microenvironment or colocalization.
+While interactive visualizations \citep{keller_2021, pardo_2022, sriworarat_2023} have the potential to mitigate this challenge, they are infeasible for scientific communications in static media, such as printed work.
+Developing a static and unified visualization that enables the simultaneous display of multiple dimensions of information is crucial for biomedical research.
 
-Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
-Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
-Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
-Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum (refer Section~\ref{sec5}).
-
-\subsection{This is an example for second level head - subsection head}\label{subsec1}
-
-Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
-Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
-Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
-Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
-
-\subsubsection{This is an example for third level head - subsubsection head}\label{subsubsec1}
-
-Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do
-eiusmod tempor incididunt ut labore et dolore magna aliqua.
-Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
-%Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
-Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
-
-\paragraph{This is an example for fourth level head - paragraph head}
-
-Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
-Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
-Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
-Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
-
-\section{This is an example for first level head}\label{sec3}
-
-\subsection{This is an example for second level head - subsection head}\label{subsec2}
-
-\subsubsection{This is an example for third level head - subsubsection head}\label{subsubsec2}
-
-Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
-Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
-Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
-Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
-
-\paragraph{This is an example for fourth level head - paragraph head}
-
-Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
-Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
-Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
-Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
-
-\section{Equations}\label{sec4}
-
-Equations in \LaTeX{} can either be inline or set as display equations. For
-inline equations use the \verb+$...$+ commands. Eg: the equation
-$H\psi = E \psi$ is written via the command \verb+$H \psi = E \psi$+.
-
-For display equations (with auto generated equation numbers)
-one can use the equation or eqnarray environments:
-\begin{equation}
-\|\tilde{X}(k)\|^2 \leq\frac{\sum\limits_{i=1}^{p}\left\|\tilde{Y}_i(k)\right\|^2+\sum\limits_{j=1}^{q}\left\|\tilde{Z}_j(k)\right\|^2 }{p+q},\label{eq1}
-\end{equation}
-where,
-\begin{align}
-D_\mu &=  \partial_\mu - ig \frac{\lambda^a}{2} A^a_\mu \nonumber \\
-F^a_{\mu\nu} &= \partial_\mu A^a_\nu - \partial_\nu A^a_\mu + g f^{abc} A^b_\mu A^a_\nu.\label{eq2}
-\end{align}
-Notice the use of \verb+\nonumber+ in the align environment at the end
-of each line, except the last, so as not to produce equation numbers on
-lines where no equation numbers are required. The \verb+\label{}+ command
-should only be used at the last line of an align environment where
-\verb+\nonumber+ is not used.
-\begin{equation}
-Y_\infty = \left( \frac{m}{\textrm{GeV}} \right)^{-3}
-    \left[ 1 + \frac{3 \ln(m/\textrm{GeV})}{15}
-    + \frac{\ln(c_2/5)}{15} \right].
-\end{equation}
-The class file also supports the use of \verb+\mathbb{}+, \verb+\mathscr{}+ and
-\verb+\mathcal{}+ commands. As such \verb+\mathbb{R}+, \verb+\mathscr{R}+
-and \verb+\mathcal{R}+ produces $\mathbb{R}$, $\mathscr{R}$ and $\mathcal{R}$
-respectively (refer Subsubsection~\ref{subsubsec3}).
-
-Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do
-eiusmod tempor incididunt ut labore et dolore magna aliqua.
-Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
-Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
-Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
-Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do
-eiusmod tempor incididunt ut labore et dolore magna aliqua.
-Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
-Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
-Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
-Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do
-eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
-
-\section{Tables}\label{sec5}
-
-Tables can be inserted via the normal table and tabular environment. To put
-footnotes inside tables one has to Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
-Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
-Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
-Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
-use the additional ``tablenotes" environment
-enclosing the tabular environment. The footnote appears just below the table
-itself (refer Tables~\ref{tab1} and \ref{tab2}).
-
-\begin{verbatim}
-\begin{table}[t]
+\begin{figure*}[!t]
 \begin{center}
-\begin{minipage}{<width>}
-\caption{<table-caption>\label{<table-label>}}%
-\begin{tabular}{@{}llll@{}}
-\toprule
-column 1 & column 2 & column 3 & column 4\\
-\midrule
-row 1 & data 1 & data 2          & data 3 \\
-row 2 & data 4 & data 5$^{1}$ & data 6 \\
-row 3 & data 7 & data 8      & data 9$^{2}$\\
-\botrule
-\end{tabular}
-\begin{tablenotes}%
-\item Source: Example for source.
-\item[$^{1}$] Example for a 1st table footnote.
-\item[$^{2}$] Example for a 2nd table footnote.
-\end{tablenotes}
-\end{minipage}
+\includegraphics[width=\textwidth]{Manuscript/figure/insitu.jpg}
+\caption{\small \textbf{\texttt{escheR} enables multi-dimensional spatial visualizations following the Gestalt principles of design}.
+(\textbf{A-B}) The traditional visualization displays the colocalization plot of the expression of two genes \textit{EFNA5} and \textit{EPHA5} (\textbf{A}) and the spatial domains from the dorsolateral prefrontal cortex in postmorteum human brain \citep{huukimyers_2023} (\textbf{B}) side-by-side, creating challenges to cognitively connecting colocalization status to spatial domains.
+(\textbf{C}) The watercolor effect enables displaying spatial domains by color-coding only outlines of circles.
+(\textbf{D}) \texttt{escheR} enables the multi-dimensional \textit{in situ} visualization that simultaneously displays the cortex layers and the colocalization status, substantially improving interpretability.}
+\label{fig:visual} 
 \end{center}
-\end{table}
-\end{verbatim}
-
-Lengthy tables which do not fit within textwidth should be set as rotated tables.
-For this, we need to use \verb+\begin{sidewaystable}...+ \verb+\end{sidewaystable}+ instead of\break \verb+\begin{table}...+ \verb+\end{table}+ environment.
-
-\begin{table}[!t]
-\caption{Caption text\label{tab1}}%
-\begin{tabular*}{\columnwidth}{@{\extracolsep\fill}llll@{\extracolsep\fill}}
-\toprule
-column 1 & column 2  & column 3 & column 4\\
-\midrule
-row 1    & data 1   & data 2  & data 3  \\
-row 2    & data 4   & data 5$^{1}$  & data 6  \\
-row 3    & data 7   & data 8  & data 9$^{2}$  \\
-\botrule
-\end{tabular*}
-\begin{tablenotes}%
-\item Source: This is an example of table footnote this is an example of table footnote this is an example of table footnote this is an example of~table footnote this is an example of table footnote
-\item[$^{1}$] Example for a first table footnote.
-\item[$^{2}$] Example for a second table footnote.
-\end{tablenotes}
-\end{table}
-
-\begin{table*}[t]
-\caption{Example of a lengthy table which is set to full textwidth.\label{tab2}}
-\tabcolsep=0pt%%
-\begin{tabular*}{\textwidth}{@{\extracolsep{\fill}}lcccccc@{\extracolsep{\fill}}}
-\toprule%
-& \multicolumn{3}{@{}c@{}}{Element 1$^{1}$} & \multicolumn{3}{@{}c@{}}{Element 2$^{2}$} \\
-\cline{2-4}\cline{5-7}%
-Project & Energy & $\sigma_{calc}$ & $\sigma_{expt}$ & Energy & $\sigma_{calc}$ & $\sigma_{expt}$ \\
-\midrule
-Element 3  & 990 A & 1168 & $1547\pm12$ & 780 A & 1166 & $1239\pm100$\\
-Element 4  & 500 A & 961  & $922\pm10$  & 900 A & 1268 & $1092\pm40$\\
-\botrule
-\end{tabular*}
-\begin{tablenotes}%
-\item Note: This is an example of table footnote this is an example of table footnote this is an example of table footnote this is an example of~table footnote this is an example of table footnote
-\item[$^{1}$] Example for a first table footnote.
-\item[$^{2}$] Example for a second table footnote.\vspace*{6pt}
-\end{tablenotes}
-\end{table*}
-
-\begin{sidewaystable}%[!p]
-\caption{Tables which are too long to fit, should be written using the ``sidewaystable" environment as shown here\label{tab3}}
-\tabcolsep=0pt%
-\begin{tabular*}{\textwidth}{@{\extracolsep{\fill}}lcccccc@{\extracolsep{\fill}}}
-\toprule%
-& \multicolumn{3}{@{}c@{}}{Element 1$^{1}$}& \multicolumn{3}{@{}c@{}}{Element$^{2}$} \\
-\cline{2-4}\cline{5-7}%
-Projectile & Energy     & $\sigma_{calc}$ & $\sigma_{expt}$ & Energy & $\sigma_{calc}$ & $\sigma_{expt}$ \\
-\midrule
-Element 3 & 990 A & 1168 & $1547\pm12$ & 780 A & 1166 & $1239\pm100$ \\
-Element 4 & 500 A & 961  & $922\pm10$  & 900 A & 1268 & $1092\pm40$ \\
-\botrule
-\end{tabular*}
-\begin{tablenotes}%
-\item Note: This is an example of a table footnote this is an example of a table footnote this is an example of a table footnote this is an example of a table footnote this is an example of a table footnote
-\item[$^{1}$] This is an example of a table footnote
-\end{tablenotes}
-\end{sidewaystable}
-
-\section{Figures}\label{sec6}
-
-As per display \LaTeX\ standards one has to use eps images for \verb+latex+ compilation and \verb+pdf/jpg/png+ images for
-\verb+pdflatex+ compilation. This is one of the major differences between \verb+latex+
-and \verb+pdflatex+. The images should be single-page documents. The command for inserting images
-for \verb+latex+ and \verb+pdflatex+ can be generalized. The package used to insert images in \verb+latex/pdflatex+ is the
-graphicx package. Figures can be inserted via the normal figure environment as shown in the below example:
-
-\begin{figure}[!t]%
-\centering
-{\color{black!20}\rule{213pt}{37pt}}
-\caption{This is a widefig. This is an example of a long caption this is an example of a long caption  this is an example of a long caption this is an example of a long caption}\label{fig1}
-\end{figure}
-
-\begin{figure*}[!t]%
-\centering
-{\color{black!20}\rule{438pt}{74pt}}
-\caption{This is a widefig. This is an example of a long caption this is an example of a long caption  this is an example of a long caption this is an example of a long caption}\label{fig2}
+\vspace{-0.3in}
 \end{figure*}
 
-\begin{verbatim}
-\begin{figure}[t]
-        \centering\includegraphics{<eps-file>}
-        \caption{<figure-caption>}
-        \label{<figure-label>}
-\end{figure}
-\end{verbatim}
+\section{Results}
 
-Test text here.
+To address these challenges, here we leverage the Gestalt (German for “unified whole”) principles \citep{todorovic_2008} as a path to visualize multi-dimensional data in 2D visualizations.
+We focus on the two types of data visualizations previously introduced that are widely used in biomedical research: (i) embedding visualizations and (ii) \textit{in situ} visualizations.
+We provide an R package, \texttt{escheR}, implementing these ideas, which is built on the state-of-the-art data visualization framework \texttt{ggplot2} in the R programming language.
+Finally, we comment on how these ideas could be extended to other types of visualization in biomedical research.
 
-For sample purposes, we have included the width of images in the
-optional argument of \verb+\includegraphics+ tag. Please ignore this.
-Lengthy figures which do not fit within textwidth should be set in rotated mode.
-For rotated figures, we need to use \verb+\begin{sidewaysfigure}+ \verb+...+ \verb+\end{sidewaysfigure}+ instead of the \verb+\begin{figure}+ \verb+...+ \verb+\end{figure}+ environment.
+\subsection{Multi-dimensional 2D visualizations with \texttt{ggplot2} and Gestalt principles}
 
-\begin{sidewaysfigure}%
-\centering
-{\color{black!20}\rule{610pt}{102pt}}
-\caption{This is an example for a sideways figure.
-This is an example of a long caption this is an example of a long caption  this is an example of a long caption this is an example of a long caption}\label{fig3}
-\end{sidewaysfigure}
+Gestalt principles \citep{todorovic_2008} refer to a set of rules describing how humans perceive and interpret visual information and are commonly applied in art and designs.
+Developed in the 1920s by German psychologists Max Wertheimer, Kurt Koffka and Wolfgang Kohler, these principles help humans perceive a set of individual elements as a whole.
 
-\section{Algorithms, Program codes and Listings}\label{sec7}
+Here, we leverage the principles to be able to visualize multi-dimensional data in a unified 2D plot.
+Our approach is to use the state-of-art data visualization framework \texttt{ggplot2} \citep{ggplot2} following the Grammar of Graphics \citep{wilkinson_2012} and map individual variables to different aesthetics to simultaneously display disparate variables.
+Specifically, we apply the figure-ground segmentation \citep{todorovic_2008} in displaying two variables: one variable (e.g.
+expression) can be plotted as color-filled circles, serving as the \textit{figure}; one variable (e.g.
+spatial domains) can be plotted as the backgrounds of the circles, creating a \textit{ground} for the figure.
+In practice, we use the combination of \texttt{color} and \texttt{fill="transparent"} to create the background layer and \texttt{fill} to create the figure layer.
+When necessary to display an additional layer for a third variable, \texttt{shape} can be used to add symbols such as cross (+) and asterisk (*) to highlight in the spatial map.
 
-Packages \verb+algorithm+, \verb+algorithmicx+ and \verb+algpseudocode+ are used for setting algorithms in latex.
-For this, one has to use the below format:
+For adjacent circles with limited space between them to display the background color, we use an economic implementation, colored outlines for these circles (\textbf{Figure~\ref{fig:visual}C}), inspired by watercolor effect \citep{pinna_2001}.
+Watercolor effect describes the phenomenon in visual perception that surface color arises from thin boundaries and hence is applied here to perceive the background color in tight space.
+Overall, the figure-ground segmentation creates two isolated layers in visual perception to display the two variables while maintaining the relative spatial relationship serving as a reference between the two.
+In addition, other fundamental principles \citep{todorovic_2008}, such as proximity, similarity, continuity, and closure, incentivize the brain to group elements and dimensions in the visualization, guaranteeing an integrative perception of the complex multi-dimensional spatial map.
 
-\begin{verbatim}
-\begin{algorithm}
-\caption{<alg-caption>}\label{<alg-label>}
-\begin{algorithmic}[1]
-. . .
-\end{algorithmic}
-\end{algorithm}
-\end{verbatim}
+\begin{figure*}[t]
+\begin{center}
+\includegraphics[width=\textwidth]{Manuscript/figure/embedding.jpg}
+\caption{\small \textbf{\texttt{escheR} improves multi-dimensional embedding visualizations}.
+The gene expression of \textit{POMGNT1} among peripheral blood mononuclear cells \citep{PBMC} under the UMAP representation.
+(\textbf{A}) The \texttt{schex} R/Bioconductor package uses color-coded convex hulls to annotate data-driven cell types, creating confusion when interpreting hexagons in overlapping hulls.
+(\textbf{B}) \texttt{escheR} plots hexagon-specific membership to avoid substantial overlapping of cluster membership.
+Due to the binning strategy, cautions in interpretation are needed due to possible cluster intermixing within each hexagon.}
+\label{fig:embedding} 
+\end{center}
+\vspace{-0.3in}
+\end{figure*}
 
-You may need to refer to the above-listed package documentations for more details before setting an \verb+algorithm+ environment.
-To set program codes, one has to use the \verb+program+ package. We need to use the \verb+\begin{program}+ \verb+...+
-\verb+\end{program}+ environment to set program codes.
+Here, we provide an open-source package called \texttt{escheR} (named after the graphic artist M.C.
+Escher) in the R programming language \citep{R}, leading to a simplified interface to navigate the implementation of the multi-dimensional visualization in 2D space.
+By adapting \texttt{ggplot2} standard, \texttt{escheR} can be seamlessly integrated with popular data objects, including \texttt{SingleCellExperiment} \citep{amezquita2020} and \texttt{SpatialExperiment} \citep{righelli2022}, and easily work with popular pipelines, such as \texttt{Seurat} \citep{hao_2021}, \texttt{Giotto} \citep{dries_2021} to name a few, and allow further theme customization with ease.
 
-\begin{algorithm}[!t]
-\caption{Calculate $y = x^n$}\label{algo1}
-\begin{algorithmic}[1]
-\Require $n \geq 0 \vee x \neq 0$
-\Ensure $y = x^n$
-\State $y \Leftarrow 1$
-\If{$n < 0$}
-        \State $X \Leftarrow 1 / x$
-        \State $N \Leftarrow -n$
-\Else
-        \State $X \Leftarrow x$
-        \State $N \Leftarrow n$
-\EndIf
-\While{$N \neq 0$}
-        \If{$N$ is even}
-            \State $X \Leftarrow X \times X$
-            \State $N \Leftarrow N / 2$
-        \Else[$N$ is odd]
-            \State $y \Leftarrow y \times X$
-            \State $N \Leftarrow N - 1$
-        \EndIf
-\EndWhile
-\end{algorithmic}
-\end{algorithm}
+Next, we give two use cases to exemplify some utility of the proposed spatial visualization: (i) the spatially differential gene colocalization in the human dorsolateral prefrontal cortex using spatial transcriptomics data \citep{huukimyers_2023}; (ii) multi-dimensional UMAP highlighting differential gene expression in data-driven cell clusters \citep{freytag_2020}.
 
-Similarly, for \verb+listings+, one has to use the \verb+listings+ package.
-The \verb+\begin{lstlisting}+ \verb+...+ \verb+\end{lstlisting}+ environment is used to set environments similar to the \verb+verbatim+ environment.
-Refer to the \verb+lstlisting+ package documentation for more details on this.
+\subsection{Multi-dimensional \textit{in situ} visualization}
+In a recent study investigating the molecular organization of human dorsolateral prefrontal cortex \citep{huukimyers_2023}, two schizophrenia risk genes, membrane-bound ligand ephrin A5 (\textit{EFNA5}) and ephrin type-A receptor 5 (\textit{EPHA5}), were identified to colocalize via cell-cell communication analysis.
+In addition, data suggested Layer 6 was the most highly co-localized layer compared to other cortex layers.
+To visually examine the inference, we applied \texttt{escheR} to create a multi-dimensional \textit{in situ} spatial map that simultaneously exhibits the cortex layers (displayed with color-coded spot outlines) and the categorized colocalization status of genes \textit{EFNA5} and \textit{EPHA5} (displayed with color-coded spot fill).
+Compared to the traditional visualization where the cortex layers and the colocalization status are visualized in two side-by-side figures (\textbf{Figure \ref{fig:visual}A-B}), our proposed visualization (\textbf{Figure \ref{fig:visual}D}) enables directly mapping colocalization status to the spatial domain, simplifying the perception of two sources of information and allowing cognitive comparison across cortex layers.
 
-\begin{minipage}{\hsize}%
-\lstset{language=Pascal}% Set your language (you can change the language for each code-block optionally)
-\begin{lstlisting}[frame=single,framexleftmargin=-1pt,framexrightmargin=-17pt,framesep=12pt,linewidth=0.98\textwidth]
-for i:=maxint to 0 do
-begin
-{ do nothing }
-end;
-Write('Case insensitive ');
-Write('Pascal keywords.');
-\end{lstlisting}
-\end{minipage}
+\subsection{Multi-dimensional embedding visualizations}
 
-\section{Cross referencing}\label{sec8}
+The application of the proposed framework is not limited to \textit{in situ} visualizations of spatially-resolved transcriptomics data.
+It is broadly applicable to data mapped to any 2-dimensional coordinate system to simultaneously display multiple variables.
+Such systems include euclidean space (including spatial coordinate as a special case) and data-driven embedding space, for example, UMAP and $t$-SNE.
+To demonstrate, we applied the proposed visualization to address the challenge of simultaneously displaying cluster membership and gene expression in a single-cell UMAP plot.
+To address the overplotting problem, previous work proposed to apply hexagonal binning strategy to display the gene expression \citep{freytag_2020}.
+Here, the color-coded convex hulls are used to annotate different clusters of hexagons, where the cluster label of each hexagon is determined by the most popular label within each hexagon bin (\textbf{Figure~\ref{fig:embedding}A}).
+However, the convex hulls create substantial overlapping areas, creating confusion when interpreting cluster memberships of hexagons in the overlapping areas.
+To improve the interpretability of the visualization, we replace the convex hulls with color-coded hexagons boundaries (\textbf{Figure~\ref{fig:embedding}B}) to avoid possible membership confusion.
+We note that our contribution to improving the visualization is easily implemented without any modification of \texttt{schex} as both are built upon the Grammar of Graphics \citep{wilkinson_2012} standard.
 
-Environments such as figure, table, equation, and align can have a label
-declared via the \verb+\label{#label}+ command. For figures and table
-environments one should use the \verb+\label{}+ command inside or just
-below the \verb+\caption{}+ command.  One can then use the
-\verb+\ref{#label}+ command to cross-reference them. As an example, consider
-the label declared for Figure \ref{fig1} which is
-\verb+\label{fig1}+. To cross-reference it, use the command
-\verb+ Figure \ref{fig1}+, for which it comes up as
-``Figure~\ref{fig1}".
+\section{Discussion}
+Here, we propose an innovative multi-dimensional spatial visualization that simultaneously displays multiple variables in a 2D coordinate system.
+Specifically, our design leverages Gestalt principles from visual perception to create multiple visual dimensions in a spatial map by iterative layering aesthetics.
+Built upon \texttt{ggplot2}, we provide an open-source R package \texttt{escheR} that is seamlessly compatible with popular spatially-resolved transcriptomics and single-cell data analysis toolboxes.
 
-\subsection{Details on reference citations}\label{subsec3}
+Adding a third dimension to 2D plots has been a long-standing challenge in visualization \citep{odonoghue_2010}.
+Our proposal addresses this fundamental challenge by introducing simple but effective design principles.
+These principles lead to visually easier-to-interpret graphics compared to 2D-color gradients and geometry annotations (\textbf{Figure~\ref{fig:embedding}}).
+Unlike computer-based interactive visualizations, the proposed visualization is free from any platform and technology restriction, creating an accessible and economical solution.
+In addition, the proposed visualization is easily scalable and hence can be applied to all types of spatially resolved data.
 
-With standard numerical .bst files, only numerical citations are possible.
-With an author-year .bst file, both numerical and author-year citations are possible.
+Beyond the scope of biomedical research, the proposed visualization can be broadly translated to any visual analytic highlighting differentiation with respect to other measurements.
+To name a few, such visual analytics include examining differential tests,  explaining clustering, and visualizing subgroups.
+However, one of the most rewarding fields to apply the proposed visualization is the rapidly expanding field of biomedical multi-omics research \citep{hasin_2017}, where connecting different omics (data modalities) is the fundamental goal and hence greatly appreciating innovative multi-dimensional visualization.
 
-If author-year citations are selected, \verb+\bibitem+ must have one of the following forms:
-
-{\footnotesize%
-\begin{verbatim}
-\bibitem[Jones et al.(1990)]{key}...
-\bibitem[Jones et al.(1990)Jones,
-                Baker, and Williams]{key}...
-\bibitem[Jones et al., 1990]{key}...
-\bibitem[\protect\citeauthoryear{Jones,
-                Baker, and Williams}
-                {Jones et al.}{1990}]{key}...
-\bibitem[\protect\citeauthoryear{Jones et al.}
-                {1990}]{key}...
-\bibitem[\protect\astroncite{Jones et al.}
-                {1990}]{key}...
-\bibitem[\protect\citename{Jones et al., }
-                1990]{key}...
-\harvarditem[Jones et al.]{Jones, Baker, and
-                Williams}{1990}{key}...
-\end{verbatim}}
-
-This is either to be made up manually, or to be generated by an
-appropriate .bst file with BibTeX. Then,
-
-{%
-\begin{verbatim}
-                    Author-year mode
-                        || Numerical mode
-\citet{key} ==>>  Jones et al. (1990)
-                        || Jones et al. [21]
-\citep{key} ==>> (Jones et al., 1990) || [21]
-\end{verbatim}}
-
-\noindent
-Multiple citations as normal:
-
-{%
-\begin{verbatim}
-\citep{key1,key2} ==> (Jones et al., 1990;
-                         Smith, 1989)||[21,24]
-        or (Jones et al., 1990, 1991)||[21,24]
-        or (Jones et al., 1990a,b)   ||[21,24]
-\end{verbatim}}
-
-\noindent
-\verb+\cite{key}+ is the equivalent of \verb+\citet{key}+ in author-year mode
-and  of \verb+\citep{key}+ in numerical mode. Full author lists may be forced with
-\verb+\citet*+ or \verb+\citep*+, e.g.
-
-{%
-\begin{verbatim}
-\citep*{key} ==>> (Jones, Baker, and Mark, 1990)
-\end{verbatim}}
-
-\noindent
-Optional notes as:
-
-{%
-\begin{verbatim}
-\citep[chap. 2]{key}     ==>>
-        (Jones et al., 1990, chap. 2)
-\citep[e.g.,][]{key}     ==>>
-        (e.g., Jones et al., 1990)
-\citep[see][pg. 34]{key} ==>>
-        (see Jones et al., 1990, pg. 34)
-\end{verbatim}}
-
-\noindent
-(Note: in standard LaTeX, only one note is allowed, after the ref.
-Here, one note is like the standard, two make pre- and post-notes.)
-
-{%
-\begin{verbatim}
-\citealt{key}   ==>> Jones et al. 1990
-\citealt*{key}  ==>> Jones, Baker, and
-                        Williams 1990
-\citealp{key}   ==>> Jones et al., 1990
-\citealp*{key}  ==>> Jones, Baker, and
-                        Williams, 1990
-\end{verbatim}}
-
-\noindent
-Additional citation possibilities (both author-year and numerical modes):
-
-{%
-\begin{verbatim}
-\citeauthor{key}       ==>> Jones et al.
-\citeauthor*{key}      ==>> Jones, Baker, and
-                                Williams
-\citeyear{key}         ==>> 1990
-\citeyearpar{key}      ==>> (1990)
-\citetext{priv. comm.} ==>> (priv. comm.)
-\citenum{key}          ==>> 11 [non-superscripted]
-\end{verbatim}}
-
-\noindent
-Note: full author lists depend on whether the bib style supports them;
-if not, the abbreviated list is printed even when full is requested.
-
-\noindent
-For names like della Robbia at the start of a sentence, use
-
-{%
-\begin{verbatim}
-\Citet{dRob98}      ==>> Della Robbia (1998)
-\Citep{dRob98}      ==>> (Della Robbia, 1998)
-\Citeauthor{dRob98} ==>> Della Robbia
-\end{verbatim}}
-
-\noindent
-The following is an example for \verb+\cite{...}+: \cite{rahman2019centroidb}.
-Another example for \verb+\citep{...}+: \citep{bahdanau2014neural,imboden2018cardiorespiratory,motiian2017unified,murphy2012machine,ji20123d}.
-Sample cites here \cite{krizhevsky2012imagenet,horvath2018dna} and \cite{pyrkov2018quantitative}, \cite{wang2018face}, \cite{lecun2015deep,zhang2018fine,ravi2016deep}.
-
-\section{Lists}\label{sec9}
-
-List in \LaTeX{} can be of three types: numbered, bulleted and unnumbered. The ``enumerate'' environment produces a numbered list, the 
-``itemize'' environment produces a bulleted list and the ``unlist''
-environment produces an unnumbered list.
-In each environment, a new entry is added via the \verb+\item+ command.
-\begin{enumerate}[1.]
-\item This is the 1st item
-
-\item Enumerate creates numbered lists, itemize creates bulleted lists and
-unnumerate creates unnumbered lists.
-\begin{enumerate}[(a)]
-\item Second level numbered list. Enumerate creates numbered lists, itemize creates bulleted lists and
-description creates unnumbered lists.
-
-\item Second level numbered list. Enumerate creates numbered lists, itemize creates bulleted lists and
-description creates unnumbered lists.
-\begin{enumerate}[(ii)]
-\item Third level numbered list. Enumerate creates numbered lists, itemize creates bulleted lists and
-description creates unnumbered lists.
-
-\item Third level numbered list. Enumerate creates numbered lists, itemize creates bulleted lists and
-description creates unnumbered lists.
-\end{enumerate}
-
-\item Second level numbered list. Enumerate creates numbered lists, itemize creates bulleted lists and
-description creates unnumbered lists.
-
-\item Second level numbered list. Enumerate creates numbered lists, itemize creates bulleted lists and
-description creates unnumbered lists.
-\end{enumerate}
-
-\item Enumerate creates numbered lists, itemize creates bulleted lists and
-description creates unnumbered lists.
-
-\item Numbered lists continue.
-\end{enumerate}
-Lists in \LaTeX{} can be of three types: enumerate, itemize and description.
-In each environment, a new entry is added via the \verb+\item+ command.
-\begin{itemize}
-\item First level bulleted list. This is the 1st item
-
-\item First level bulleted list. Itemize creates bulleted lists and description creates unnumbered lists.
-\begin{itemize}
-\item Second level dashed list. Itemize creates bulleted lists and description creates unnumbered lists.
-
-\item Second level dashed list. Itemize creates bulleted lists and description creates unnumbered lists.
-
-\item Second level dashed list. Itemize creates bulleted lists and description creates unnumbered lists.
-\end{itemize}
-
-\item First level bulleted list. Itemize creates bulleted lists and description creates unnumbered lists.
-
-\item First level bulleted list. Bullet lists continue.
-\end{itemize}
-
-\noindent
-Example for unnumbered list items:
-
-\begin{unlist}
-\item Sample unnumberd list text. Sample unnumberd list text. Sample unnumberd list text. Sample unnumberd list text. Sample unnumberd list text.
-
-\item Sample unnumberd list text. Sample unnumberd list text. Sample unnumberd list text.
-
-\item sample unnumberd list text.
-Sample unnumberd list text.
-Sample unnumberd list text.
-Sample unnumberd list text.
-Sample unnumberd list text.
-Sample unnumberd list text.
-Sample unnumberd list text.
-\end{unlist}
-
-\section{Examples for theorem-like environments}\label{sec10}
-
-For theorem-like environments, we require the \verb+amsthm+ package.
-There are three types of predefined theorem styles - \verb+thmstyleone+, \verb+thmstyletwo+ and \verb+thmstylethree+   (check your journal's instructions page in case a specific style is required).
-
-\medskip
-\noindent\begin{tabular}{|l|p{13pc}|}
-\hline
-\verb+thmstyleone+ & Numbered, theorem head in bold font and theorem text in italic style \\\hline
-\verb+thmstyletwo+ & Numbered, theorem head in roman font and theorem text in italic style \\\hline
-\verb+thmstylethree+ & Numbered, theorem head in bold font and theorem text in roman style \\\hline
-\end{tabular}
-
-\begin{theorem}[Theorem subhead]\label{thm1}
-Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text.
-Example theorem text. Example theorem text. Example theorem text. Example theorem text. Example theorem text.
-Example theorem text.
-\end{theorem}
-
-Quisque ullamcorper placerat ipsum. Cras nibh. Morbi vel justo vitae lacus tincidunt ultrices. Lorem ipsum dolor sit
-amet, consectetuer adipiscing elit. In hac habitasse platea dictumst. Integer tempus convallis augue.
-
-\begin{proposition}
-Example proposition text. Example proposition text. Example proposition text. Example proposition text. Example proposition text.
-Example proposition text. Example proposition text. Example proposition text. Example proposition text. Example proposition text.
-\end{proposition}
-
-Nulla malesuada porttitor diam. Donec felis erat, congue non, volutpat at, tincidunt tristique, libero. Vivamus
-viverra fermentum felis. Donec nonummy pellentesque ante.
-
-\begin{example}
-Phasellus adipiscing semper elit. Proin fermentum massa
-ac quam. Sed diam turpis, molestie vitae, placerat a, molestie nec, leo. Maecenas lacinia. Nam ipsum ligula, eleifend
-at, accumsan nec, suscipit a, ipsum. Morbi blandit ligula feugiat magna. Nunc eleifend consequat lorem.
-\end{example}
-
-Nulla malesuada porttitor diam. Donec felis erat, congue non, volutpat at, tincidunt tristique, libero. Vivamus
-viverra fermentum felis. Donec nonummy pellentesque ante.
-
-\begin{remark}
-Phasellus adipiscing semper elit. Proin fermentum massa
-ac quam. Sed diam turpis, molestie vitae, placerat a, molestie nec, leo. Maecenas lacinia. Nam ipsum ligula, eleifend
-at, accumsan nec, suscipit a, ipsum. Morbi blandit ligula feugiat magna. Nunc eleifend consequat lorem.
-\end{remark}
-
-Quisque ullamcorper placerat ipsum. Cras nibh. Morbi vel justo vitae lacus tincidunt ultrices. Lorem ipsum dolor sit
-amet, consectetuer adipiscing elit. In hac habitasse platea dictumst.
-
-\begin{definition}[Definition sub head]
-Example definition text. Example definition text. Example definition text. Example definition text. Example definition text. Example definition text. Example definition text. Example definition text.
-\end{definition}
-
-Apart from the above styles, we have the \verb+\begin{proof}+ \verb+...+ \verb+\end{proof}+ environment - with the proof head in italic style and the body text in roman font with an open square at the end of each proof environment.
-
-\begin{proof}Example for proof text.
-Example for proof text.
-Example for proof text.
-Example for proof text.
-Example for proof text.
-Example for proof text.
-Example for proof text.
-Example for proof text.
-Example for proof text.
-Example for proof text.
-\end{proof}
-
-Nam dui ligula, fringilla a, euismod sodales, sollicitudin vel, wisi. Morbi auctor lorem non justo. Nam lacus libero,
-pretium at, lobortis vitae, ultricies et, tellus. Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna,
-vitae ornare odio metus a mi.
-
-\begin{proof}[Proof of Theorem~{\upshape\ref{thm1}}]
-Example for proof text.
-Example for proof text.
-Example for proof text.
-Example for proof text.
-Example for proof text.
-Example for proof text.
-Example for proof text.
-Example for proof text.
-Example for proof text.
-Example for proof text.
-\end{proof}
-
-\noindent
-For a quote environment, one has to use\newline \verb+\begin{quote}...\end{quote}+
-\begin{quote}
-Quoted text example. Aliquam porttitor quam a lacus. Praesent vel arcu ut tortor cursus volutpat. In vitae pede quis diam bibendum placerat. Fusce elementum
-convallis neque. Sed dolor orci, scelerisque ac, dapibus nec, ultricies ut, mi. Duis nec dui quis leo sagittis commodo.
-\end{quote}
-Donec congue.
-Maecenas urna mi, suscipit in, placerat ut, vestibulum ut, massa.
-Fusce ultrices nulla et nisl (refer Figure~\ref{fig3}).
-Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas.
-Etiam ligula arcu,
-elementum a, venenatis quis, sollicitudin sed, metus. Donec nunc pede, tincidunt in, venenatis vitae, faucibus vel (refer Table~\ref{tab3}).
-
-\section{Conclusion}
-
-Some Conclusions here.
-
+While the proposed design principle set-up a solid foundation, the proposed package \texttt{escheR} can be further optimized for image-based spatially resolved data.
+For example, to address possible overplotting of densely arranged single cells in tight space, strategies such as hatching plots \citep{patrick2023spatial} can be incorporated.
+In addition, the flexibility to represent data points as polygons will be provided to allow visualization of cell morphology.
+In summary, we propose a novel multi-dimensional visualization, implemented in an R package \texttt{escheR}, to address the simultaneous exhibition of multiple variables in 2D plots.
+The proposed visualization can be broadly applicable to the visual analytics of growingly complex biomedical data and beyond.
 %%%%%%%%%%%%%%
 
-\begin{appendices}
-
-\section{Section title of first appendix}\label{sec11}
-
-Nam dui ligula, fringilla a, euismod sodales, sollicitudin vel, wisi. Morbi auctor lorem non justo. Nam lacus libero,
-pretium at, lobortis vitae, ultricies et, tellus. Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna,
-vitae ornare odio metus a mi. Morbi ac orci et nisl hendrerit mollis. Suspendisse ut massa. Cras nec ante. Pellentesque
-a nulla. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Aliquam tincidunt
-urna. Nulla ullamcorper vestibulum turpis. Pellentesque cursus luctus mauris.
-
-\subsection{Subsection title of first appendix}\label{subsec4}
-
-Nam dui ligula, fringilla a, euismod sodales, sollicitudin vel, wisi. Morbi auctor lorem non justo. Nam lacus libero,
-pretium at, lobortis vitae, ultricies et, tellus. Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna,
-vitae ornare odio metus a mi. Morbi ac orci et nisl hendrerit mollis. Suspendisse ut massa. Cras nec ante. Pellentesque
-a nulla. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Aliquam tincidunt
-urna. Nulla ullamcorper vestibulum turpis. Pellentesque cursus luctus mauris.
-
-\subsubsection{Subsubsection title of first appendix}\label{subsubsec3}
-
-Example for an unnumbered figure:
-
-\begin{figure}[!h]
-\centering
-{\color{black!20}\rule{85pt}{92pt}}
-\end{figure}
-
-Fusce mauris. Vestibulum luctus nibh at lectus. Sed bibendum, nulla a faucibus semper, leo velit ultricies tellus, ac
-venenatis arcu wisi vel nisl. Vestibulum diam. Aliquam pellentesque, augue quis sagittis posuere, turpis lacus congue
-quam, in hendrerit risus eros eget felis.
-
-\section{Section title of second appendix}\label{sec12}%
-
-Fusce mauris. Vestibulum luctus nibh at lectus. Sed bibendum, nulla a faucibus semper, leo velit ultricies tellus, ac
-venenatis arcu wisi vel nisl. Vestibulum diam. Aliquam pellentesque, augue quis sagittis posuere, turpis lacus congue
-quam, in hendrerit risus eros eget felis. Maecenas eget erat in sapien mattis porttitor. Vestibulum porttitor. Nulla
-facilisi. Sed a turpis eu lacus commodo facilisis. Morbi fringilla, wisi in dignissim interdum, justo lectus sagittis dui, et
-vehicula libero dui cursus dui. Mauris tempor ligula sed lacus. Duis cursus enim ut augue. Cras ac magna. Cras nulla.
-
-\begin{figure}[b]
-\centering
-{\color{black!20}\rule{217pt}{120pt}}
-\caption{This is an example for appendix figure\label{fig4}}
-\end{figure}
-
-\begin{table}[t]%
-\begin{center}
-\begin{minipage}{.52\columnwidth}
-\caption{This is an example of Appendix table showing food requirements of army, navy and airforce\label{tab4}}%
-\begin{tabular}{@{}lcc@{}}%
-\toprule
-col1 head & col2 head & col3 head \\
-\midrule
-col1 text & col2 text & col3 text \\
-col1 text & col2 text & col3 text \\
-col1 text & col2 text & col3 text \\
-\botrule
-\end{tabular}
-\end{minipage}
-\end{center}
-\end{table}
-
-\subsection{Subsection title of second appendix}\label{subsec5}
-
-Sed commodo posuere pede. Mauris ut est. Ut quis purus. Sed ac odio. Sed vehicula hendrerit sem. Duis non odio.
-Morbi ut dui. Sed accumsan risus eget odio. In hac habitasse platea dictumst. Pellentesque non elit. Fusce sed justo
-eu urna porta tincidunt. Mauris felis odio, sollicitudin sed, volutpat a, ornare ac, erat. Morbi quis dolor. Donec
-pellentesque, erat ac sagittis semper, nunc dui lobortis purus, quis congue purus metus ultricies tellus. Proin et quam.
-Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos hymenaeos. Praesent sapien turpis,
-fermentum vel, eleifend faucibus, vehicula eu, lacus.
-
-Sed commodo posuere pede. Mauris ut est. Ut quis purus. Sed ac odio. Sed vehicula hendrerit sem. Duis non odio.
-Morbi ut dui. Sed accumsan risus eget odio. In hac habitasse platea dictumst. Pellentesque non elit. Fusce sed justo
-eu urna porta tincidunt. Mauris felis odio, sollicitudin sed, volutpat a, ornare ac, erat. Morbi quis dolor. Donec
-pellentesque, erat ac sagittis semper, nunc dui lobortis purus, quis congue purus metus ultricies tellus. Proin et quam.
-Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos hymenaeos. Praesent sapien turpis,
-fermentum vel, eleifend faucibus, vehicula eu, lacus.
-
-\subsubsection{Subsubsection title of second appendix}\label{subsubsec4}
-
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae,
-felis. Curabitur dictum gravida mauris. Nam arcu libero, nonummy eget, consectetuer id, vulputate a, magna. Donec
-vehicula augue eu neque.
-
-Example for an equation inside the appendix:
-\begin{align}
- p &= \frac{\gamma^{2} - (n_{C} -1)H}{(n_{C} - 1) + H - 2\gamma}, \label{1eq:hybobo:pfromgH} \\
- \theta &= \frac{(\gamma - H)^{2}(\gamma - n_{C} -1)^{2}}{(n_{C} - 1 + H - 2\gamma)^{2}} \label{2eq:hybobo:tfromgH}\; .
-\end{align}
-
-\section{Example of another appendix section}\label{sec13}%
-
-Nam dui ligula, fringilla a, euismod sodales, sollicitudin vel, wisi. Morbi auctor lorem non justo. Nam lacus libero,
-pretium at, lobortis vitae, ultricies et, tellus. Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna,
-vitae ornare odio metus a mi. Morbi ac orci et nisl hendrerit mollis. Suspendisse ut massa. Cras nec ante. Pellentesque
-a nulla. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Aliquam tincidunt
-urna. Nulla ullamcorper vestibulum turpis. Pellentesque cursus luctus mauris
-\begin{equation}
-\mathcal{L} = i \bar{\psi} \gamma^\mu D_\mu \psi
-    - \frac{1}{4} F_{\mu\nu}^a F^{a\mu\nu} - m \bar{\psi} \psi.
-\label{eq26}
-\end{equation}
-
-Nulla malesuada porttitor diam. Donec felis erat, congue non, volutpat at, tincidunt tristique, libero. Vivamus
-viverra fermentum felis. Donec nonummy pellentesque ante. Phasellus adipiscing semper elit. Proin fermentum massa
-ac quam. Sed diam turpis, molestie vitae, placerat a, molestie nec, leo. Maecenas lacinia. Nam ipsum ligula, eleifend
-at, accumsan nec, suscipit a, ipsum. Morbi blandit ligula feugiat magna. Nunc eleifend consequat lorem. Sed lacinia
-nulla vitae enim. Pellentesque tincidunt purus vel magna. Integer non enim. Praesent euismod nunc eu purus. Donec
-bibendum quam in tellus. Nullam cursus pulvinar lectus. Donec et mi. Nam vulputate metus eu enim. Vestibulum
-pellentesque felis eu massa.
-
-Nulla malesuada porttitor diam. Donec felis erat, congue non, volutpat at, tincidunt tristique, libero. Vivamus
-viverra fermentum felis. Donec nonummy pellentesque ante. Phasellus adipiscing semper elit. Proin fermentum massa
-ac quam. Sed diam turpis, molestie vitae, placerat a, molestie nec, leo. Maecenas lacinia. Nam ipsum ligula, eleifend
-at, accumsan nec, suscipit a, ipsum. Morbi blandit ligula feugiat magna. Nunc eleifend consequat lorem. Sed lacinia
-nulla vitae enim. Pellentesque tincidunt purus vel magna. Integer non enim. Praesent euismod nunc eu purus. Donec
-bibendum quam in tellus. Nullam cursus pulvinar lectus. Donec et mi. Nam vulputate metus eu enim. Vestibulum
-pellentesque felis eu massa.
-
-Nulla malesuada porttitor diam. Donec felis erat, congue non, volutpat at, tincidunt tristique, libero. Vivamus
-viverra fermentum felis. Donec nonummy pellentesque ante. Phasellus adipiscing semper elit. Proin fermentum massa
-ac quam. Sed diam turpis, molestie vitae, placerat a, molestie nec, leo. Maecenas lacinia. Nam ipsum ligula, eleifend
-at, accumsan nec, suscipit a, ipsum. Morbi blandit ligula feugiat magna. Nunc eleifend consequat lorem. Sed lacinia
-nulla vitae enim. Pellentesque tincidunt purus vel magna. Integer non enim. Praesent euismod nunc eu purus. Donec
-bibendum quam in tellus. Nullam cursus pulvinar lectus. Donec et mi. Nam vulputate metus eu enim. Vestibulum
-pellentesque felis eu massa.
-
-Nulla malesuada porttitor diam. Donec felis erat, congue non, volutpat at, tincidunt tristique, libero. Vivamus
-viverra fermentum felis. Donec nonummy pellentesque ante. Phasellus adipiscing semper elit. Proin fermentum massa
-ac quam. Sed diam turpis, molestie vitae, placerat a, molestie nec, leo. Maecenas lacinia. Nam ipsum ligula, eleifend
-at, accumsan nec, suscipit a, ipsum. Morbi blandit ligula feugiat magna. Nunc eleifend consequat lorem. Sed lacinia
-nulla vitae enim. Pellentesque tincidunt purus vel magna. Integer non enim. Praesent euismod nunc eu purus. Donec
-bibendum quam in tellus. Nullam cursus pulvinar lectus. Donec et mi. Nam vulputate metus eu enim. Vestibulum
-pellentesque felis eu massa. Donec
-bibendum quam in tellus. Nullam cursus pulvinar lectus. Donec et mi. Nam vulputate metus eu enim. Vestibulum
-pellentesque felis eu massa.
-
-%% Example for unnumbered table inside appendix
-\begin{table}
-\begin{center}
-\begin{minipage}{.52\columnwidth}
-\caption{}{%
-\begin{tabular}{lcc}%
-\toprule
-col1 head & col2 head & col3 head \\
-\midrule
-col1 text & col2 text & col3 text \\
-col1 text & col2 text & col3 text \\
-col1 text & col2 text & col3 text \\
-\botrule
-\end{tabular}}{}
-\end{minipage}
-\end{center}
-\end{table}
-
-\end{appendices}
+\section{Availability of data and materials}
+The spatial transcriptomics dataset was obtained from \texttt{spatialLIBD} (\href{http://research.libd.org/spatialLIBD}{http://research.libd.org/spatialLIBD}).
+The UMAP example follows the `using\_schex' vignette in the \texttt{schex} package (\href{https://www.bioconductor.org/packages/schex}{http://bioconductor.org/packages/schex}).
+The code that generates these figures is deposited at \href{https://github.com/boyiguo1/Manuscript_escheR}{http://github.com/boyiguo1/Manuscript\_escheR} (Zenodo DOI: \href{https://zenodo.org/record/7915970}{10.5281/zenodo.7915970}).
+The open source software package \texttt{escheR} available in the R programming language is freely available on GitHub (\href{https://github.com/boyiguo1/escheR}{http://github.com/boyiguo1/escheR}) and Bioconductor (\href{https://bioconductor.org/packages/escheR}{http://bioconductor.org/packages/escheR}).
 
@@
 \section{Author contributions statement}
-
-Must include all authors, identified by initials, for example:
-S.R. and D.A. conceived the experiment(s),  S.R. conducted the experiment(s), S.R. and D.A. analysed the results.  S.R. and D.A. wrote and reviewed the manuscript.
+\textbf{B. G.}:  Conceptualization, Methodology, Software, Validation, Formal analysis, Investigation, Data Curation, Writing, Visualization; 
+\textbf{L.A.H.M.}: Conceptualization, Software;
+\textbf{M.G.P.}: Conceptualization;
+\textbf{L.C.T.}: Conceptualization, Software;
+ \textbf{S.C.H.}: Conceptualization, Resources, Writing - Review \& Editing, Visualization, Supervision, Project administration, Funding acquisition.
 
 \section{Acknowledgments}
-The authors thank the anonymous reviewers for their valuable suggestions. This work is supported in part by funds from the National Science Foundation (NSF: \# 1636933 and \# 1920920).
+The authors thank the anonymous reviewers for their valuable suggestions.
+This project was supported by the National Institute of Mental Health [R01MH126393 to BG and SCH, U01MH122849 to LAHM and LCT]; the Chan Zuckerberg Initiative DAF, an advised fund of Silicon Valley Community Foundation [CZF2019-002443 to SCH]; the Lieber Institute for Brain Development to LAHM and LCT; and Aligning Science Across Parkinson’s [ASAP-000478, ASAP-000509 to MGP] through the Michael J.
+Fox Foundation for Parkinson’s Research.
+All funding bodies had no role in the design of the study and collection, analysis, and interpretation of data and in writing the manuscript.
+We also would like to acknowledge Nicholas J.
+Eagles, Kristen R.
+Maynard, Mina Ryten, Leon Di Stefano, and Lukas M.
+Weber (appearing in alphabetic order of last name) for their helpful comments, feedback and suggestions on \texttt{escheR} functionality.
+Nicholas J.
+Eagles and Kristen R.
+Maynard are employed by the Lieber Institute for Brain Development; Mina Ryten is employed by University College London; Leon Di Stefano is from the Johns Hopkins Bloomberg School of Public Health, Department of Biostatistics; Lukas M.
+Weber is in the Department of Biostatistics at Boston University School of Public Health.
 
-%\bibliographystyle{plain}
-%\bibliography{reference}
-
-\begin{thebibliography}{10}
-
-\bibitem{bahdanau2014neural}
-Dzmitry Bahdanau, Kyunghyun Cho, and Yoshua Bengio.
-\newblock Neural machine translation by jointly learning to align and
-  translate.
-\newblock {\em arXiv preprint arXiv:1409.0473}, 2014.
-
-\bibitem{horvath2018dna}
-Steve Horvath and Kenneth Raj.
-\newblock Dna methylation-based biomarkers and the epigenetic clock theory of
-  ageing.
-\newblock {\em Nature Reviews Genetics}, 19(6):371, 2018.
-
-\bibitem{imboden2018cardiorespiratory}
-Mary~T Imboden, Matthew~P Harber, Mitchell~H Whaley, W~Holmes Finch, Derron~L
-  Bishop, and Leonard~A Kaminsky.
-\newblock Cardiorespiratory fitness and mortality in healthy men and women.
-\newblock {\em Journal of the American College of Cardiology},
-  72(19):2283--2292, 2018.
-
-\bibitem{ji20123d}
-Shuiwang Ji, Wei Xu, Ming Yang, and Kai Yu.
-\newblock 3d convolutional neural networks for human action recognition.
-\newblock {\em IEEE Transactions on Pattern Analysis and Machine Intelligence},
-  35(1):221--231, 2012.
-
-\bibitem{krizhevsky2012imagenet}
-Alex Krizhevsky, Ilya Sutskever, and Geoffrey~E Hinton.
-\newblock Imagenet classification with deep convolutional neural networks.
-\newblock In {\em Advances in Neural Information Processing Systems}, pages
-  1097--1105, 2012.
-
-\bibitem{lecun2015deep}
-Yann LeCun, Yoshua Bengio, and Geoffrey Hinton.
-\newblock Deep learning.
-\newblock {\em Nature}, 521(7553):436, 2015.
-
-\bibitem{motiian2017unified}
-Saeid Motiian, Marco Piccirilli, Donald~A Adjeroh, and Gianfranco Doretto.
-\newblock Unified deep supervised domain adaptation and generalization.
-\newblock In {\em Proceedings of the IEEE International Conference on Computer
-  Vision}, pages 5715--5725, 2017.
-
-\bibitem{murphy2012machine}
-Kevin~P Murphy.
-\newblock {\em Machine learning: A probabilistic perspective}.
-\newblock MIT press, 2012.
-
-\bibitem{american2013acsm}
-American~College of~Sports~Medicine et~al.
-\newblock {\em ACSM's guidelines for exercise testing and prescription}.
-\newblock Lippincott Williams \& Wilkins, 2013.
-
-\bibitem{pyrkov2018quantitative}
-Timothy~V Pyrkov, Evgeny Getmantsev, Boris Zhurov, Konstantin Avchaciov,
-  Mikhail Pyatnitskiy, Leonid Menshikov, Kristina Khodova, Andrei~V Gudkov, and
-  Peter~O Fedichev.
-\newblock Quantitative characterization of biological age and frailty based on
-  locomotor activity records.
-\newblock {\em Aging (Albany NY)}, 10(10):2973, 2018.
-
-\bibitem{rahman2019centroidb}
-Syed~Ashiqur Rahman and Donald Adjeroh.
-\newblock Centroid of age neighborhoods: A generalized approach to estimate
-  biological age.
-\newblock In {\em 2019 IEEE EMBS International Conference on Biomedical \&
-  Health Informatics (BHI)}, pages 1--4. IEEE, 2019.
-
-\bibitem{ravi2016deep}
-Daniele Rav{\`\i}, Charence Wong, Fani Deligianni, Melissa Berthelot, Javier
-  Andreu-Perez, Benny Lo, and Guang-Zhong Yang.
-\newblock Deep learning for health informatics.
-\newblock {\em IEEE {J}ournal of {B}iomedical and {H}ealth {I}nformatics},
-  21(1):4--21, 2016.
-
-\bibitem{wang2018face}
-Zongwei Wang, Xu~Tang, Weixin Luo, and Shenghua Gao.
-\newblock Face aging with identity-preserved conditional generative adversarial
-  networks.
-\newblock In {\em Proceedings of the IEEE Conference on Computer Vision and
-  Pattern Recognition}, pages 7939--7947, 2018.
-
-\bibitem{zhang2018fine}
-Ke~Zhang, Na~Liu, Xingfang Yuan, Xinyao Guo, Ce~Gao, and Zhenbing Zhao.
-\newblock Fine-grained age estimation in the wild with attention {LSTM}
-  networks.
-\newblock {\em arXiv preprint arXiv:1805.10445}, 2018.
-
-\end{thebibliography}
+% \bibliographystyle{plain}
+% \bibliography{refs}
 
 %USE THE BELOW OPTIONS IN CASE YOU NEED AUTHOR YEAR FORMAT.
-%\bibliographystyle{abbrvnat}
-%\bibliography{reference}
-
-%% sample for biography with author's image
-\begin{biography}{{\color{black!20}\rule{77pt}{77pt}}}{\author{Author Name.} This is sample author biography text.
-The values provided in the optional argument are meant for sample purposes.
-There is no need to include the width and height of an image in the optional argument for live articles.
-This is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text.}
-\end{biography}
-
-%% sample for biography without author's image
-\begin{biography}{}{\author{Author Name.} This is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text this is sample author biography text.}
-\end{biography}
+\bibliographystyle{abbrvnat}
+\bibliography{refs}
 

### 2023-11-14 R/02_multi_d_umap.R
Add session info

@@ -146,3 +146,245 @@ ggsave("Manuscript/figure/embedding.tiff", p_fig2,
        height = 8,
        width = 12)
 
+# Session Info ------------------------------------------------------------
+# > sessionInfo()
+# R version 4.3.0 (2023-04-21)
+# Platform: aarch64-apple-darwin20 (64-bit)
+# Running under: macOS Ventura 13.6
+#
+# Matrix products: default
+# BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib
+# LAPACK: /Library/Frameworks/R.framework/Versions/4.3-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.11.0
+#
+# locale:
+#   [1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8
+#
+# time zone: America/New_York
+# tzcode source: internal
+#
+# attached base packages:
+#   [1] stats4    stats     graphics
+# [4] grDevices utils     datasets
+# [7] methods   base
+#
+# other attached packages:
+#   [1] escheR_1.3.1
+# [2] BiocSingular_1.18.0
+# [3] ggrepel_0.9.4
+# [4] scran_1.30.0
+# [5] scater_1.30.0
+# [6] scuttle_1.12.0
+# [7] TENxPBMCData_1.20.0
+# [8] HDF5Array_1.30.0
+# [9] rhdf5_2.46.0
+# [10] DelayedArray_0.28.0
+# [11] SparseArray_1.2.1
+# [12] S4Arrays_1.2.0
+# [13] abind_1.4-5
+# [14] Matrix_1.6-1
+# [15] schex_1.16.0
+# [16] shiny_1.7.5.1
+# [17] ggplot2_3.4.4
+# [18] Seurat_5.0.0
+# [19] SeuratObject_5.0.0
+# [20] sp_2.1-1
+# [21] SingleCellExperiment_1.24.0
+# [22] SummarizedExperiment_1.32.0
+# [23] Biobase_2.62.0
+# [24] GenomicRanges_1.54.1
[…]

### 2023-11-14 Manuscript/manuscript_BioAdv_format.tex
Clean up ref

@@
 To address the overplotting problem, previous work proposed to apply hexagonal binning strategy to display the gene expression \citep{freytag_2020}.
-Here, the color-coded convex hulls are used to annotate different clusters of hexagons, where the cluster label of each hexagon is determined by the most popular label within each hexagon bin (\textbf{Figure~\ref{fig:embedding}A}).
+Here, the color-coded convex hulls are used to annotate different clusters of \rr{hexagons, where the cluster label of each hexagon is determined by the most popular label within each hexagon bin} (\textbf{Figure~\ref{fig:embedding}A}).
 However, the convex hulls create substantial overlapping areas, creating confusion when interpreting cluster memberships of hexagons in the overlapping areas.

### 2023-11-15 Manuscript/escheR_BioAdv_clean.tex
Clean up ref
typo

--- 2023-11-14 Manuscript/manuscript_BioAdv_format.tex
@@
 
-\newcommand{\rr}[1]{{\color{red} #1}}
 \usepackage{setspace}
@@
 \caption{\small \textbf{\texttt{escheR} enables multi-dimensional spatial visualizations following the Gestalt principles of design}.
-(\textbf{A-B}) The traditional visualization displays the colocalization plot of the expression of two genes \textit{EFNA5} and \textit{EPHA5} (\textbf{A}) and the spatial domains from the dorsolateral prefrontal cortex in postmorteum human brain \citep{huukimyers_2023} (\textbf{B}) side-by-side, creating challenges to cognitively connecting colocalization status to spatial domains.
+(\textbf{A-B}) The traditional visualization displays the colocalization plot of the expression of two genes \textit{EFNA5} and \textit{EPHA5} (\textbf{A}) and the spatial domains from the dorsolateral prefrontal cortex in postmortem human brain \citep{huukimyers_2023} (\textbf{B}) side-by-side, creating challenges to cognitively connecting colocalization status to spatial domains.
 (\textbf{C}) The watercolor effect enables displaying spatial domains by color-coding only outlines of circles.

### 2023-11-28 Manuscript/escheR_BioAdv_final.tex
Clean up reference section.

--- 2023-11-15 Manuscript/escheR_BioAdv_clean.tex
@@
 
-\onecolumn % for one column layouts
-
 %\usepackage{showframe}
@@
 % \usepackage[mathlines, switch]{lineno}
-\usepackage[right]{lineno}
-\linenumbers
+% \usepackage[right]{lineno}
+% \linenumbers
 
@@
 \newtheorem{definition}{Definition}
-
-\usepackage{setspace}
-\doublespacing
 
@@
 This creates cognitive gaps on how to associate the disparate information or how to interpret the biological findings of this multi-dimensional information regarding their microenvironment or colocalization.
-While interactive visualizations \citep{keller_2021, pardo_2022, sriworarat_2023} have the potential to mitigate this challenge, they are infeasible for scientific communications in static media, such as printed work.
+While interactive visualizations \citep{pardo_2022, sriworarat_2023} have the potential to mitigate this challenge, they are infeasible for scientific communications in static media, such as printed work.
 Developing a static and unified visualization that enables the simultaneous display of multiple dimensions of information is crucial for biomedical research.
@@
 
-% \bibliographystyle{plain}
-% \bibliography{refs}
-
 %USE THE BELOW OPTIONS IN CASE YOU NEED AUTHOR YEAR FORMAT.
-\bibliographystyle{abbrvnat}
+% \bibliographystyle{bioinformatics}
+\bibliographystyle{apalike2}
 \bibliography{refs}
